---
title: "1. Getting started"
author: "1.7 The statistical forecasting perspective"
date: "OTexts.org/fpp3/"
classoption: aspectratio=169
titlepage: fpp3title.png
titlecolor: fpp3red
toc: false
output:
  binb::monash:
    colortheme: monashwhite
    fig_width: 7.5
    fig_height: 3.5
    keep_tex: no
    includes:
      in_header: fpp3header.tex
---

```{r setup, include=FALSE}
source("setup.R")
library(patchwork)
library(gganimate)
austa <- readr::read_csv("http://OTexts.com/fpp3/extrafiles/austa.csv") |>
  as_tsibble(index = Year)
melsyd <- tsibbledata::ansett |>
  filter(Airports == "MEL-SYD")
```

## Random futures

```{r austa, echo=FALSE}
# Grab ABS data
austa <- readxl::read_excel("data/340101.xlsx", sheet = "Data1", skip = 9) |>
  rename(date = `Series ID`, value = A85375847A) |>
  select(date, value) |>
  transmute(
    Month = yearmonth(date),
    Visitors = value / 1e3
  ) |>
  bind_rows(tibble(
    Month = yearmonth(seq(as.Date("2021-11-01"), by = "1 month", length = 2)),
    Visitors = NA_real_
  )) |>
  as_tsibble(index = Month) |>
  filter(Month >= yearmonth("2000 Jan"))
# Fit ETS model
fit <- austa |>
  filter(Month < yearmonth("2018 Jan")) |>
  model(ETS(Visitors))
# Product forecasts
fc <- forecast(fit, h = 48) |>
  mutate(Month = as.Date(Month))
# Simulate 100 future sample paths
set.seed(1967)
sim <- fit |>
  generate(h = 48, times = 100) |>
  mutate(
    replicate = factor(.rep, levels = 1:100, labels = paste("Future", 1:100)),
    .rep = as.numeric(.rep)
  ) |>
  as_tibble() |>
  mutate(Month = as.Date(Month))
# Nice colors
cols <- scale_colour_manual(
  values = c("#D55E00", "#0072B2", "#009E73", "#CC79A7", "#E69F00", "#56B4E9", "#F0E442", sample(rainbow(93))),
  breaks = paste("Future", 1:100),
  name = " "
)
# Now build up some plots with alignment
p1 <- austa |>
  mutate(Visitors = if_else(Month >= yearmonth("2018 Jan"), NA_real_, Visitors)) |>
  as_tibble() |>
  mutate(Month = as.Date(Month)) |>
  ggplot(aes(x = Month, y = Visitors)) +
  geom_line() +
  labs(
    x = "Month",
    y = "Thousands of visitors",
    title = "Total short-term visitors to Australia"
  ) +
  scale_x_date(
    breaks = seq(as.Date("2000-01-01"), by = "5 years", l = 5),
    labels = paste("Jan", seq(2000, 2020, by = 5)),
    minor_breaks = seq(as.Date("2001-01-01"), by = "1 year", l = 25)
  ) +
  ylim(min(austa$Visitors, sim$.sim, na.rm = TRUE), max(austa$Visitors, sim$.sim, na.rm = TRUE))
p2 <- p1 + cols +
  geom_line(aes(y = .sim, group = replicate, colour = replicate),
    data = sim |> filter(.rep <= 1)
  )
p3 <- p1 + cols +
  geom_line(aes(y = .sim, group = replicate, colour = replicate),
    data = sim |> filter(.rep <= 2)
  )
p4 <- p1 + cols +
  geom_line(aes(y = .sim, group = replicate, colour = replicate),
    data = sim |> filter(.rep <= 3)
  )
p5 <- p1 + cols +
  geom_line(aes(y = .sim, group = replicate, colour = replicate),
    data = sim |> filter(.rep <= 10)
  )
p6 <- p1 + cols +
  geom_line(aes(y = .sim, group = replicate, colour = replicate), alpha = 0.5, data = sim) +
  guides(colour = "none")
p7 <- p1 +
  geom_line(aes(y = .sim, group = replicate, colour = replicate), alpha = 0.5, col = "gray", data = sim) +
  guides(colour = "none")
p8 <- p7 + autolayer(fc, level = c(50, 90))
p9 <- p8 + coord_cartesian(xlim = as.Date(c("2014-01-01", "2021-08-01"))) +
  scale_x_date(
    breaks = seq(as.Date("2000-01-01"), by = "2 years", l = 11),
    labels = paste("Jan", seq(2000, 2020, by = 2)),
    minor_breaks = seq(as.Date("2001-01-01"), by = "1 year", l = 25)
  )
aligned_plots <- align_patches(p1, p2, p3, p4, p5, p6, p7, p8, p9)
```

\forecast\pause

```{r austa1, dependson="austa", echo=FALSE}
aligned_plots[[1]]
```

## Random futures

\forecast

```{r austa2, dependson='austa', echo=FALSE}
aligned_plots[[2]]
```

\simfutures

## Random futures

\forecast

```{r austa3, dependson='austa', echo=FALSE}
aligned_plots[[3]]
```

\simfutures

## Random futures

\forecast

```{r austa4, dependson='austa', echo=FALSE}
aligned_plots[[4]]
```

\simfutures

## Random futures

\forecast

```{r austa5, dependson='austa', echo=FALSE}
aligned_plots[[5]]
```

\simfutures

## Random futures

\forecast

```{r austa6, dependson='austa', echo=FALSE}
aligned_plots[[6]]
```

\simfutures

## Random futures

\forecast

```{r austa7, dependson='austa', echo=FALSE}
aligned_plots[[7]]
```

\simfutures

## Random futures

\forecast

```{r austa8, dependson='austa', echo=FALSE}
aligned_plots[[8]]
```

\simfutures

## Random futures

\forecast

```{r austa9, dependson='austa', echo=FALSE}
aligned_plots[[9]]
```

\simfutures

## Random futures

\forecast

```{r austa9b, dependson='austa', echo=FALSE}
aligned_plots[[9]] +
  geom_line(
    colour = "black",
    data = austa |> filter(Month >= yearmonth("2018 Jan")) |> mutate(Month = as.Date(Month))
  )
```

\simfutures

\only<2>{\begin{textblock}{8.5}(.3,6.5)\begin{alertblock}{}
``He who sees the past as surprise-free is bound to have a future full of surprises.''\\\mbox{}\hfill{\small (Amos Tversky)}
\end{alertblock}\end{textblock}
}

## Statistical forecasting

\fontsize{14}{16}\sf

- Thing to be forecast: a random variable, $y_t$.
- Forecast distribution: If ${\cal I}$ is all observations, then $y_{t} |{\cal I}$ means "the random variable $y_{t}$ given what we know in ${\cal I}$.
- The "point forecast" is the mean (or median) of $y_{t} |{\cal I}$
- The "forecast variance" is $\text{var}[y_{t} |{\cal I}]$
- A prediction interval or "interval forecast" is a range of values of $y_t$ with high probability.
- With time series, ${y}_{t|t-1} = y_t | \{y_1,y_2,\dots,y_{t-1}\}$.
- $\hat{y}_{T+h|T} =\text{E}[y_{T+h} | y_1,\dots,y_T]$ (an $h$-step forecast taking account of all observations up to time $T$).
