---
title: "ETC3550/ETC5550 Applied&nbsp;forecasting"
author: "Ch1. Getting started"
date: "OTexts.org/fpp3/"
classoption: aspectratio=169
titlepage: title16x9.png
titlecolor: burntorange
toc: true
output:
  binb::monash:
    colortheme: monashwhite
    fig_width: 8.0
    fig_height: 3.3
    keep_tex: no
    includes:
      in_header: header.tex
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  message = FALSE,
  warning = FALSE,
  cache = TRUE,
  dev.args = list(pointsize = 11)
)
options(
  digits = 3,
  width = 80,
  ggplot2.continuous.colour="viridis",
  ggplot2.continuous.fill = "viridis",
  ggplot2.discrete.colour = c("#D55E00", "#0072B2","#009E73", "#CC79A7", "#E69F00", "#56B4E9", "#F0E442"),
  ggplot2.discrete.fill = c("#D55E00", "#0072B2","#009E73", "#CC79A7", "#E69F00", "#56B4E9", "#F0E442")
)
library(fpp3)
library(patchwork)
library(gganimate)
```

```{r datasets, include=FALSE}
austa <- readr::read_csv("http://OTexts.com/fpp3/extrafiles/austa.csv") %>%
  as_tsibble(index = Year)
melsyd <- tsibbledata::ansett %>%
  filter(Airports == "MEL-SYD")
```

# The dark history of forecasting

## Forecasting by maggots

\placefig{0.8}{1.5}{width=8cm}{sheepliverbm}

\begin{textblock}{7}(8.5,1.7)
\raggedright\large\color{black}
\textbf{Clay model of sheep's liver}\\[0.4cm]
Used by Bab\-ylon\-ian forecasters approximately 600 B.C.\\[0.7cm]
Now in \rlap{British Museum.}
\end{textblock}

## Forecasting by hallucination

\only<1->{\placefig{0}{1.35}{height=8.25cm, width=20cm}{delphitemple}}
\only<2>{\placefig{7}{1.35}{height=7.77cm, width=20cm}{1106}}

## Forecasting and the law

\placefig{10.1}{1.35}{width=6cm, height=20cm}{250px-Constance_II_Colosseo_Rome_Italy}

\begin{textblock}{8}(1,1.8)\raggedright
\begin{block}{}
\color{black}
Anyone who consults a soothsayer on account of curiosity of the future will suffer capital punishment.

\mbox{}\hfill\emph{Codex Theodosianus 9.16.4}
\end{block}
\end{textblock}

## Reputations can be made and lost

\begin{block}{}
``I think there is a world market for maybe five computers.'' \newline{\null\hfill\small (Chairman of IBM, 1943)}
\end{block}
  \pause
\begin{alertblock}{}
``There is no reason anyone would want a computer in their home.'' \newline\null\hfill{\small (President, DEC, 1977)}
\end{alertblock}
  \pause
\begin{block}{}
“There’s no chance that the iPhone is going to get any significant market share. No chance.”\hfill{\small (Steve Ballmer, CEO Microsoft, April 2007)}
\end{block}
  \pause
\begin{alertblock}{}
``We're going to be opening relatively soon \dots The virus \dots\ will go away in April.''\hfill{\small (Donald Trump, February 2020)}
\end{alertblock}

# What can we forecast?

## Forecasts that aren't forecasts

\fullheight{hopecasts2}

## What can we forecast?

\only<1>{\fullheight{AusBOM}}
\only<2>{\placefig{0}{1.35}{height=10.8cm, width=20cm}{nasdaq-stock-market}}
\only<3>{\placefig{0}{1.35}{height=10.2cm, width=20cm}{Forex2}}
\only<4>{\placefig{0}{1.35}{height=12.2cm, width=20cm, trim=0 0 0 100, clip=true}{ts22015}}
\only<5>{\placefig{0}{1.35}{height=10.2cm, width=20cm}{comet}}
\only<6>{\placefig{0}{1.35}{height=11.2cm, width=20cm}{pills}}
\only<7>{\placefig{0}{1.35}{height=9.2cm, width=20cm}{elecwires}
\begin{textblock}{12}(0.2,8.8)\tiny\textcolor{gray}{\texttt{Varistor60, CC BY-SA 4.0, https://commons.wikimedia.org/w/index.php?curid=59368531}}
\end{textblock}}
\only<8>{\placefig{0}{1.35}{height=9.2cm, width=20cm}{covid}}
\only<9>{\placefig{0}{1.35}{height=10.2cm, width=20cm}{12apostles}}

## Which is easiest to forecast?

 * daily electricity demand in 3 days time
 * timing of next Halley's comet appearance
 * time of sunrise this day next year
 * Google stock price tomorrow
 * Google stock price in 6 months time
 * maximum temperature tomorrow
 * exchange rate of \$US/AUS next week
 * total sales of drugs in Australian pharmacies next month

\vspace*{10cm}

## Which is easiest to forecast?

 1. time of sunrise this day next year
 2. timing of next Halley's comet appearance
 3. maximum temperature tomorrow
 4. daily electricity demand in 3 days time
 5. total sales of drugs in Australian pharmacies next month
 6. Google stock price tomorrow
 7. exchange rate of \$US/AUS next week
 8. Google stock price in 6 months time

\only<2>{\begin{block}{}\begin{itemize}\tightlist
 \item how do we measure ``easiest''?
 \item what makes something easy/difficult to forecast?
 \end{itemize}\end{block}}

\vspace*{10cm}

## Forecastability factors

Something is easier to forecast if:

 1. we have a good understanding of the factors that contribute to it
 2. there is lots of data available;
 3. the future is somewhat similar to the past
 4. the forecasts cannot affect the thing we are trying to forecast.

# Time series data and random futures

## Time series data

  - Four-yearly Olympic winning times
  - Annual Google profits
  - Quarterly Australian beer production
  - Monthly rainfall
  - Weekly retail sales
  - Daily IBM stock prices
  - Hourly electricity demand
  - 5-minute freeway traffic counts
  - Time-stamped stock transaction data

## Random futures

```{r austa, echo=FALSE}
# Grab ABS data
austa <- readxl::read_excel("data/340101.xlsx", sheet="Data1", skip=9) %>%
  rename(date = `Series ID`, value = A85375847A) %>%
  select(date, value) %>%
  transmute(
    Month = yearmonth(date),
    Visitors = value / 1e3
  ) %>%
  bind_rows(tibble(
    Month = yearmonth(seq(as.Date("2021-11-01"), by = "1 month", length = 2)),
    Visitors = NA_real_
  )) %>%
  as_tsibble(index = Month) %>%
  filter(Month >= yearmonth("2000 Jan"))
# Fit ETS model
fit <- austa %>%
  filter(Month < yearmonth("2018 Jan")) %>%
  model(ETS(Visitors))
# Product forecasts
fc <- forecast(fit, h = 48) %>%
  mutate(Month = as.Date(Month))
# Simulate 100 future sample paths
set.seed(1967)
sim <- fit %>%
  generate(h = 48, times = 100) %>%
  mutate(
    replicate = factor(.rep, levels = 1:100, labels = paste("Future", 1:100)),
    .rep = as.numeric(.rep)
  ) %>%
  as_tibble() %>%
  mutate(Month = as.Date(Month))
# Nice colors
cols <- scale_colour_manual(
  values = c("#D55E00", "#0072B2", "#009E73", "#CC79A7", "#E69F00", "#56B4E9", "#F0E442", sample(rainbow(93))),
  breaks = paste("Future", 1:100),
  name = " "
)
# Now build up some plots with alignment
p1 <- austa %>%
  mutate(Visitors = if_else(Month >= yearmonth("2018 Jan"), NA_real_, Visitors)) %>%
  as_tibble() %>%
  mutate(Month = as.Date(Month)) %>%
  ggplot(aes(x = Month, y = Visitors)) +
  geom_line() +
  labs(
    x = "Month",
    y = "Thousands of visitors",
    title = "Total short-term visitors to Australia"
  ) +
  scale_x_date(
    breaks = seq(as.Date("2000-01-01"), by = "5 years", l = 5),
    labels = paste("Jan", seq(2000, 2020, by = 5)),
    minor_breaks = seq(as.Date("2001-01-01"), by = "1 year", l = 25)
  ) +
  ylim(min(austa$Visitors, sim$.sim, na.rm = TRUE), max(austa$Visitors, sim$.sim, na.rm = TRUE))
p2 <- p1 + cols +
  geom_line(aes(y = .sim, group = replicate, colour = replicate),
    data = sim %>% filter(.rep <= 1)
  )
p3 <- p1 + cols +
  geom_line(aes(y = .sim, group = replicate, colour = replicate),
    data = sim %>% filter(.rep <= 2)
  )
p4 <- p1 + cols +
  geom_line(aes(y = .sim, group = replicate, colour = replicate),
    data = sim %>% filter(.rep <= 3)
  )
p5 <- p1 + cols +
  geom_line(aes(y = .sim, group = replicate, colour = replicate),
    data = sim %>% filter(.rep <= 10)
  )
p6 <- p1 + cols +
  geom_line(aes(y = .sim, group = replicate, colour = replicate), alpha = 0.5, data = sim) +
  guides(colour = "none")
p7 <- p1 +
  geom_line(aes(y = .sim, group = replicate, colour = replicate), alpha = 0.5, col = "gray", data = sim) +
  guides(colour = "none")
p8 <- p7 + autolayer(fc, level = c(50, 90))
p9 <- p8 + coord_cartesian(xlim=as.Date(c("2014-01-01","2021-08-01"))) +
  scale_x_date(
    breaks = seq(as.Date("2000-01-01"), by = "2 years", l = 11),
    labels = paste("Jan", seq(2000, 2020, by = 2)),
    minor_breaks = seq(as.Date("2001-01-01"), by = "1 year", l = 25)
  ) 
aligned_plots <- align_patches(p1, p2, p3, p4, p5, p6, p7, p8,p9)
```

\forecast\pause

```{r austa1, dependson="austa", echo=FALSE}
aligned_plots[[1]]
```

## Random futures

\forecast

```{r austa2, dependson='austa', echo=FALSE}
aligned_plots[[2]]
```

\simfutures

## Random futures

\forecast

```{r austa3, dependson='austa', echo=FALSE}
aligned_plots[[3]]
```

\simfutures

## Random futures

\forecast

```{r austa4, dependson='austa', echo=FALSE}
aligned_plots[[4]]
```

\simfutures

## Random futures

\forecast

```{r austa5, dependson='austa', echo=FALSE}
aligned_plots[[5]]
```

\simfutures

## Random futures

\forecast

```{r austa6, dependson='austa', echo=FALSE}
aligned_plots[[6]]
```

\simfutures

## Random futures

\forecast

```{r austa7, dependson='austa', echo=FALSE}
aligned_plots[[7]]
```

\simfutures

## Random futures

\forecast

```{r austa8, dependson='austa', echo=FALSE}
aligned_plots[[8]]
```

\simfutures

## Random futures

\forecast

```{r austa9, dependson='austa', echo=FALSE}
aligned_plots[[9]]
```

\simfutures

## Random futures

\forecast

```{r austa9b, dependson='austa', echo=FALSE}
aligned_plots[[9]] +
  geom_line(
    colour = "black",
    data = austa %>% filter(Month >= yearmonth("2018 Jan")) %>% mutate(Month = as.Date(Month))
  )
```

\simfutures

\only<2>{\begin{textblock}{8.5}(.3,6.5)\begin{alertblock}{}
``He who sees the past as surprise-free is bound to have a future full of surprises.''\\\mbox{}\hfill{\small (Amos Tversky)}
\end{alertblock}\end{textblock}
}

## Statistical forecasting

\fontsize{14}{16}\sf

- Thing to be forecast: a random variable, $y_t$.
- Forecast distribution: If ${\cal I}$ is all observations, then $y_{t} |{\cal I}$ means "the random variable $y_{t}$ given what we know in ${\cal I}$.
- The "point forecast" is the mean (or median) of $y_{t} |{\cal I}$
- The "forecast variance" is $\text{var}[y_{t} |{\cal I}]$
- A prediction interval or "interval forecast" is a range of values of $y_t$ with high probability.
- With time series, ${y}_{t|t-1} = y_t | \{y_1,y_2,\dots,y_{t-1}\}$.
- $\hat{y}_{T+h|T} =\text{E}[y_{T+h} | y_1,\dots,y_T]$ (an $h$-step forecast taking account of all observations up to time $T$).

# Some case studies

## CASE STUDY 1: Paperware company

\fontsize{11.5}{13}\sf

\begin{textblock}{7.6}(0.2,1.5)
\textbf{Problem:} Want forecasts of each of hundreds of
items. Series can be stationary, trended or seasonal. They currently
have a large forecasting program written in-house but it doesn't seem
to produce sensible forecasts. They want me to tell them what is
wrong and fix it.

\vspace*{0.cm}

\textbf{Additional information}\vspace*{-0.2cm}\fontsize{11.5}{13}\sf
\begin{itemize}\itemsep=0cm\parskip=0cm
\item  Program  written in COBOL making numerical calculations limited. It is not possible to do any optimisation.
\item Their programmer has little experience in numerical computing.
\item They employ no statisticians and want the program to produce forecasts \rlap{automatically.}
\end{itemize}
\end{textblock}

\placefig{8}{1.4}{width=4.8cm}{tableware2}

## CASE STUDY 1: Paperware company

### Methods currently used

A
: 12 month average

C
: 6 month average

E
: straight line regression over last 12 months

G
: straight line regression over last 6 months

H
: average slope between last year's and this year's values.
  (Equivalent to differencing at lag 12 and taking mean.)

I
: Same as H except over 6 months.

K
: I couldn't understand the explanation.

## CASE STUDY 2: PBS

\fullwidth{pills}

## CASE STUDY 2: PBS

### The Pharmaceutical Benefits Scheme (PBS) is the Australian government drugs subsidy scheme.

  * Many drugs bought from pharmacies are subsidised to allow more equitable access to modern drugs.
  * The cost to government is determined by the number and types of drugs purchased. Currently nearly 1\% of GDP.
  * The total cost is budgeted based on forecasts of drug usage.

## CASE STUDY 2: PBS

\fullheight{pbs2}

## CASE STUDY 2: PBS

  * In 2001: \$4.5 billion budget, under-forecasted by \$800 million.
  * Thousands of products. Seasonal demand.
  * Subject to covert marketing, volatile products, uncontrollable expenditure.
  * Although monthly data available for 10 years, data are aggregated to annual values, and only the first three years are used in estimating the forecasts.
  * All forecasts being done with the \texttt{FORECAST} function in MS-Excel!

## CASE STUDY 3: Car fleet company

**Client:** One of Australia's largest car fleet companies

**Problem:** how to forecast resale value of vehicles? How
should this affect leasing and sales policies?

\pause

### Additional information
 - They can provide a large amount of data on previous vehicles and their eventual resale values.
 - The resale values are currently estimated by a group of specialists. They see me as a threat and do not cooperate.

## CASE STUDY 4: Airline

\fullheight{ansettlogo}

## CASE STUDY 4: Airline

```{r, echo=FALSE}
melsyd %>%
  filter(Class == "Economy") %>%
  autoplot(Passengers) +
  labs(
    title = "Economy class passengers",
    subtitle = "Melbourne-Sydney",
    x = "Year",
    y = "Thousands"
  )
```

\only<2>{\begin{textblock}{4.2}(11,5.5)
\begin{alertblock}{}
Not the real data! Or is it?
\end{alertblock}
\end{textblock}}

## CASE STUDY 4: Airline

**Problem:** how to forecast passenger traffic on major routes?

### Additional information

  * They can provide a large amount of data on previous routes.
  * Traffic is affected by school holidays, special events such as
the Grand Prix, advertising campaigns, competition behaviour, etc.
  * They have a highly capable team of people who are able to do
most of the computing.

# Assignment 1

## Assignment 1: forecast the following series
\small

  1. Google closing stock price on 21 March 2022.
  2. Maximum temperature at Melbourne airport on 5 April 2022.
  3. The difference in points (Collingwood minus Essendon) scored in the AFL match between Collingwood and Essendon for the Anzac Day clash. 25 April 2022.
  4. The seasonally adjusted estimate of total employment for April 2022. ABS CAT 6202, to be released around mid May 2022.
  5. Google closing stock price on 23 May 2022.

\begin{block}{Due Sunday 13 March}
For each of these, give a point forecast and an 80\% prediction interval.
\end{block}\pause
\begin{alertblock}{}
Prize: \$50 Amazon gift voucher
\end{alertblock}

## Assignment 1: scoring
\small

$Y=$ actual, $F=$ point forecast, $[L,U]=$ prediction interval

### Point forecasts:

$$\text{Absolute Error} = |Y-F|
$$

 * Rank results for all students in class
 * Add ranks across all five items

### Prediction intervals:

$$
\text{Interval Score} = (U - L) + 10(L - Y)_+ + 10 (Y-U)_+
$$

 * Rank results for all students
 * Add ranks across all five items
