---
title: "10. Dynamic regression models"
author: "10.2 Dynamic regression using fable"
date: "OTexts.org/fpp3/"
classoption: aspectratio=169
titlepage: fpp3title.png
titlecolor: fpp3red
toc: false
output:
  binb::monash:
    colortheme: monashwhite
    fig_width: 7.5
    fig_height: 2.8
    keep_tex: no
    includes:
      in_header: fpp3header.tex
---

```{r setup, include=FALSE}
source("setup.R")
library(readr)

vic_elec_daily <- vic_elec |>
  filter(year(Time) == 2014) |>
  index_by(Date = date(Time)) |>
  summarise(
    Demand = sum(Demand) / 1e3,
    Temperature = max(Temperature),
    Holiday = any(Holiday)
  ) |>
  mutate(Day_Type = case_when(
    Holiday ~ "Holiday",
    wday(Date) %in% 2:6 ~ "Weekday",
    TRUE ~ "Weekend"
  ))
```


## Regression with ARIMA errors

  * In `fable`, we can specify an ARIMA($p,d,q$) for the errors, and $d$ levels of differencing will be applied to all variables ($y, x_{1,t},\dots,x_{k,t}$) during estimation.
  * Check that $\varepsilon_t$ series looks like white noise.
  * AICc can be calculated for final model.
  * Repeat procedure for all subsets of predictors to be considered, and select model with lowest AICc value.

## US personal consumption and income

```{r usconsump, fig.height=6, fig.width=9, echo=FALSE, out.height="92%"}
us_change |>
  gather(key = "variable", value = "value") |>
  ggplot(aes(y = value, x = Quarter, group = variable, colour = variable)) +
  geom_line() +
  facet_grid(variable ~ ., scales = "free_y") +
  labs(y = "", title = "Quarterly changes in US consumption and personal income") +
  guides(colour = "none")
```

## US personal consumption and income
\fontsize{11}{12}\sf

```{r usconsump2, echo=TRUE, fig.height=3}
fit <- us_change |> model(ARIMA(Consumption ~ Income))
report(fit)
```

\pause\begin{alertblock}{}
Write down the equations for the fitted model.
\end{alertblock}

## US personal consumption and income

```{r , echo=TRUE, fig.height=3}
residuals(fit, type = "regression") |>
  gg_tsdisplay(.resid, plot_type = "partial") +
  labs(title = "Regression errors")
```

## US personal consumption and income

```{r , echo=TRUE, fig.height=3}
residuals(fit, type = "innovation") |>
  gg_tsdisplay(.resid, plot_type = "partial") +
  labs(title = "ARIMA errors")
```

## US personal consumption and income
\fontsize{11}{12}\sf

```{r , echo=TRUE, fig.height=3.7}
augment(fit) |>
  features(.innov, ljung_box, dof = 3, lag = 12)
```
