---
title: "10. Dynamic regression models"
author: "10.3 Forecasting"
date: "OTexts.org/fpp3/"
classoption: aspectratio=169
titlepage: fpp3title.png
titlecolor: fpp3red
toc: false
output:
  binb::monash:
    colortheme: monashwhite
    fig_width: 7.5
    fig_height: 2.8
    keep_tex: no
    includes:
      in_header: fpp3header.tex
---

```{r setup, include=FALSE}
source("setup.R")
library(readr)

vic_elec_daily <- vic_elec |>
  filter(year(Time) == 2014) |>
  index_by(Date = date(Time)) |>
  summarise(
    Demand = sum(Demand) / 1e3,
    Temperature = max(Temperature),
    Holiday = any(Holiday)
  ) |>
  mutate(Day_Type = case_when(
    Holiday ~ "Holiday",
    wday(Date) %in% 2:6 ~ "Weekday",
    TRUE ~ "Weekend"
  ))
```


## US personal consumption and income

```{r usconsump3, echo=TRUE, fig.height=2.2}
fit <- us_change |> model(ARIMA(Consumption ~ Income))
us_change_future <- new_data(us_change, 8) |>
  mutate(Income = mean(us_change$Income))
forecast(fit, new_data = us_change_future) |>
  autoplot(us_change) +
  labs(x = "Year", y = "Percentage change",
       title = "Forecasts from regression with ARIMA(1,0,2) errors")
```

## Forecasting

  * To forecast a regression model with ARIMA errors, we need to forecast the
regression part of the model and the ARIMA part of the model and combine the
results.
  * Some predictors are known into the future (e.g., time, dummies).
  * Separate forecasting models may be needed for other predictors.
  * Forecast intervals ignore the uncertainty in forecasting the predictors.

## Daily electricity demand

```{r, echo=TRUE}
vic_elec_daily |>
  ggplot(aes(x = Temperature, y = Demand, colour = Day_Type)) +
  geom_point() +
  labs(x = "Maximum temperature", y = "Electricity demand (GW)")
```

## Daily electricity demand

```{r, echo=TRUE, fig.height=2.55}
vic_elec_daily |>
  pivot_longer(c(Demand, Temperature)) |>
  ggplot(aes(x = Date, y = value)) +
  geom_line() +
  facet_grid(name ~ ., scales = "free_y") +
  labs(y = "")
```

## Daily electricity demand
\fontsize{9}{9.5}\sf

```{r, echo=TRUE}
fit <- vic_elec_daily |>
  model(arima = ARIMA(Demand ~ Temperature + I(Temperature^2) +
    (Day_Type == "Weekday")))
report(fit)
```

## Daily electricity demand

```{r, echo=TRUE}
gg_tsresiduals(fit)
```

## Daily electricity demand
\fontsize{10}{11}\sf

```{r, echo=TRUE}
augment(fit) |>
  features(.resid, ljung_box, dof = 6, lag = 14)
```

## Daily electricity demand
\fontsize{10}{13}\sf

```{r, echo=TRUE}
# Forecast one day ahead
vic_next_day <- new_data(vic_elec_daily, 1) |>
  mutate(Temperature = 26, Day_Type = "Holiday")
forecast(fit, vic_next_day)
```

## Daily electricity demand
\fontsize{10}{11}\sf
```{r, echo=TRUE}
vic_elec_future <- new_data(vic_elec_daily, 14) |>
  mutate(
    Temperature = 26,
    Holiday = c(TRUE, rep(FALSE, 13)),
    Day_Type = case_when(
      Holiday ~ "Holiday",
      wday(Date) %in% 2:6 ~ "Weekday",
      TRUE ~ "Weekend"
    )
  )
```

## Daily electricity demand

```{r, echo = TRUE}
forecast(fit, new_data = vic_elec_future) |>
  autoplot(vic_elec_daily) + labs(y = "GW")
```
