---
title: "10. Dynamic regression models"
author: "10.1 Estimation"
date: "OTexts.org/fpp3/"
classoption: aspectratio=169
titlepage: fpp3title.png
titlecolor: fpp3red
toc: false
output:
  binb::monash:
    colortheme: monashwhite
    fig_width: 7.5
    fig_height: 2.8
    keep_tex: no
    includes:
      in_header: fpp3header.tex
---

```{r setup, include=FALSE}
source("setup.R")
library(readr)

vic_elec_daily <- vic_elec |>
  filter(year(Time) == 2014) |>
  index_by(Date = date(Time)) |>
  summarise(
    Demand = sum(Demand) / 1e3,
    Temperature = max(Temperature),
    Holiday = any(Holiday)
  ) |>
  mutate(Day_Type = case_when(
    Holiday ~ "Holiday",
    wday(Date) %in% 2:6 ~ "Weekday",
    TRUE ~ "Weekend"
  ))
```

## Stochastic \& deterministic trends

\begin{block}{Deterministic trend}
$$ y_t = \beta_0 + \beta_1 t + \eta_t  \qquad \text{where $\eta_t \sim \text{ARIMA}(p,0,q)$}
$$
\end{block}\pause

\begin{block}{Stochastic trend}
$$y_t = \beta_0 + \beta_1 t + \eta_t \qquad \text{where $\eta_t\sim \text{ARIMA}(p,1,q)$}
$$
\pause
Difference both sides:\vspace*{-0.3cm}
$$y_t = \beta_1 + \eta'_t \qquad \text{where $\eta'_t\sim \text{ARIMA}(p,0,q)$}
$$
\end{block}

## Air transport passengers Australia

```{r}
aus_airpassengers |>
  autoplot(Passengers) +
  labs(y = "Passengers (millions)", title = "Total air passengers")
```

## Air transport passengers Australia

**Deterministic trend**

\fontsize{10}{10}\sf

```{r, echo=TRUE}
fit_deterministic <- aus_airpassengers |>
  model(ARIMA(Passengers ~ 1 + trend() + pdq(d = 0)))
report(fit_deterministic)
```

```{r austaparams, echo=FALSE, dependson='deterministictrend'}
coef <- rlang::set_names(tidy(fit_deterministic)$estimate, tidy(fit_deterministic)$term)
phi1 <- coef["ar1"]
intercept <- coef["intercept"]
slope <- coef["trend()"]
sigma2 <- glance(fit_deterministic)$sigma2
```

\only<2>{\begin{textblock}{5}(9,4.5)
\begin{block}{}\vspace*{-0.2cm}\fontsize{12}{13}\sf
\begin{align*}
  y_t &= `r sprintf("%.3f", intercept)` + `r sprintf("%.3f", slope)` t + \eta_t \\
  \eta_t &= `r sprintf("%.3f", phi1)` \eta_{t-1}  + \varepsilon_t\\
  \varepsilon_t &\sim \text{NID}(0,`r sprintf("%.3f", sigma2)`).
\end{align*}
\end{block}
\end{textblock}}

\vspace*{10cm}

## Air transport passengers Australia

**Stochastic trend**

\fontsize{10}{10}\sf

```{r, echo=TRUE}
fit_stochastic <- aus_airpassengers |>
  model(ARIMA(Passengers ~ 1 + pdq(d = 1)))
report(fit_stochastic)
```

```{r austaparams2, echo=FALSE, dependson='stochastictrend'}
coef <- rlang::set_names(tidy(fit_stochastic)$estimate, tidy(fit_stochastic)$term)
drift <- coef["constant"]
sigma2 <- glance(fit_stochastic)$sigma2
```

\only<2>{\begin{textblock}{5}(9,4.5)
\begin{block}{}\vspace*{-0.2cm}\fontsize{12}{13}\sf
\begin{align*}
  y_t-y_{t-1} &= `r sprintf("%.3f", drift)` + \varepsilon_t,\\
  y_t &= y_0 + `r sprintf("%.3f", drift)` t + \eta_t \\
  \eta_t &= \eta_{t-1} + \varepsilon_{t}\\
  \varepsilon_t &\sim \text{NID}(0,`r sprintf("%.3f", sigma2)`).
\end{align*}
\end{block}
\end{textblock}}

\vspace*{10cm}

## Air transport passengers Australia

```{r, fig.height=2.2}
aus_airpassengers |>
  autoplot(Passengers) +
  autolayer(fit_stochastic |> forecast(h = 20),
    colour = "#0072B2", level = 95) +
  autolayer(fit_deterministic |> forecast(h = 20),
    colour = "#D55E00", alpha = 0.65, level = 95) +
  labs(y = "Air passengers (millions)", title = "Forecasts from trend models")
```

## Forecasting with trend

  * Point forecasts are almost identical, but prediction intervals differ.
  * Stochastic trends have much wider prediction intervals because the errors are non-stationary.
  * Be careful of forecasting with deterministic trends too far ahead.
