---
title: "10. Dynamic regression models"
author: "10.6 Lagged predictors"
date: "OTexts.org/fpp3/"
classoption: aspectratio=169
titlepage: fpp3title.png
titlecolor: fpp3red
toc: false
output:
  binb::monash:
    colortheme: monashwhite
    fig_width: 7.5
    fig_height: 2.8
    keep_tex: no
    includes:
      in_header: fpp3header.tex
---


```{r setup, include=FALSE}
source("setup.R")
library(readr)

vic_elec_daily <- vic_elec |>
  filter(year(Time) == 2014) |>
  index_by(Date = date(Time)) |>
  summarise(
    Demand = sum(Demand) / 1e3,
    Temperature = max(Temperature),
    Holiday = any(Holiday)
  ) |>
  mutate(Day_Type = case_when(
    Holiday ~ "Holiday",
    wday(Date) %in% 2:6 ~ "Weekday",
    TRUE ~ "Weekend"
  ))
```

## Lagged predictors

Sometimes a change in $x_t$ does not affect $y_t$ instantaneously\pause
\begin{block}{}
\begin{itemize}
  \item $y_t=$ sales, $x_t=$ advertising.
  \item $y_t=$ stream flow, $x_t=$ rainfall.
  \item $y_t=$ size of herd, $x_t=$ breeding stock.
\end{itemize}
\end{block}
\pause

  * These are dynamic systems with input ($x_t$) and output $(y_t)$.
  * $x_t$ is often a leading indicator.
  * There can be multiple predictors.

## Lagged predictors

The model include present and past values of predictor:
\begin{block}{}
\centerline{$
y_t = a + \gamma_0x_t + \gamma_1x_{t-1} + \dots + \gamma_kx_{t-k} + \eta_t$}
\end{block}
where $\eta_t$ is an ARIMA process.\pause

**Rewrite model as **\vspace*{-0.6cm}
\begin{align*}
y_{t} & = a+ (\gamma_{0} + \gamma_{1} B + \gamma_{2} B^{2} + \dots + \gamma_{k} B^{k}) x_{t} +\eta_t \\
      & = a+ \gamma(B) x_{t} +\eta_t.
\end{align*}\pause\vspace*{-1.2cm}

  * $\gamma(B)$ is called a \textit{transfer function} since it describes how
change in $x_t$ is transferred to $y_t$.
  * $x$ can influence $y$, but $y$ is not allowed to influence $x$.

## Example: Insurance quotes and TV adverts

```{r tvadvert, fig.height=2.4}
insurance |>
  pivot_longer(Quotes:TVadverts) |>
  ggplot(aes(x = Month, y = value)) +
  geom_line() +
  facet_grid(vars(name), scales = "free_y") +
  labs(y = NULL, title = "Insurance advertising and quotations")
```

## Example: Insurance quotes and TV adverts

```{r tvadvertpairs, dependson='tvadvertdata', echo=FALSE}
insurance |>
  mutate(
    lag1 = lag(TVadverts),
    lag2 = lag(lag1)
  ) |>
  as_tibble() |>
  select(-Month) |>
  rename(lag0 = TVadverts) |>
  pivot_longer(-Quotes, names_to = "Lag", values_to = "TV_advert") |>
  ggplot(aes(x = TV_advert, y = Quotes)) +
  geom_point() +
  facet_grid(. ~ Lag) +
  labs(title = "Insurance advertising and quotations")
```

## Example: Insurance quotes and TV adverts

```{r, echo=TRUE}
fit <- insurance |>
  # Restrict data so models use same fitting period
  mutate(Quotes = c(NA, NA, NA, Quotes[4:40])) |>
  # Estimate models
  model(
    ARIMA(Quotes ~ pdq(d = 0) + TVadverts),
    ARIMA(Quotes ~ pdq(d = 0) + TVadverts + lag(TVadverts)),
    ARIMA(Quotes ~ pdq(d = 0) + TVadverts + lag(TVadverts) +
      lag(TVadverts, 2)),
    ARIMA(Quotes ~ pdq(d = 0) + TVadverts + lag(TVadverts) +
      lag(TVadverts, 2) + lag(TVadverts, 3))
  )
```

## Example: Insurance quotes and TV adverts
\fontsize{10}{10}\sf

```{r, echo=TRUE, results = 'hide'}
glance(fit)
```
```{r, echo = FALSE}
glance(fit) |>
  transmute(`Lag order` = 0:3, sigma2, log_lik, AIC, AICc, BIC) |>
  knitr::kable()
```

## Example: Insurance quotes and TV adverts

\fontsize{10}{10}\sf

```{r tvadvertagain, echo=TRUE}
fit_best <- insurance |>
  model(ARIMA(Quotes ~ pdq(d = 0) + TVadverts + lag(TVadverts)))
report(fit_best)
```

\pause
\vspace{-1em}

```{r tvadvertparam, echo=FALSE, dependson="tvadvertagain"}
# Store coefficients
tidy_fit <- tidy(fit_best)
coef <- rlang::set_names(tidy_fit$estimate, tidy_fit$term)
phi1 <- coef["ar1"]
ma1 <- coef["ma1"]
ma2 <- coef["ma2"]
intercept <- coef["intercept"]
gamma0 <- coef["TVadverts"]
gamma1 <- coef["lag(TVadverts)"]
```
\vspace*{0.2cm}

###
\vspace*{-0.2cm}\begin{align*}
  y_t &= `r sprintf("%.3f", intercept)` +
         `r sprintf("%.3f", gamma0)` x_t +
         `r sprintf("%.3f", gamma1)` x_{t-1} + \eta_t,\\
  \eta_t &= `r sprintf("%.3f", phi1)` \eta_{t-1} +
                                     \varepsilon_t +
            `r sprintf("%.3f", ma1)` \varepsilon_{t-1} +
            `r sprintf("%.3f", ma2)` \varepsilon_{t-2},
\end{align*}

## Example: Insurance quotes and TV adverts

```{r, echo=TRUE, fig.height=3}
advert_a <- new_data(insurance, 20) |>
  mutate(TVadverts = 10)
forecast(fit_best, advert_a) |> autoplot(insurance)
```

## Example: Insurance quotes and TV adverts

```{r, echo=TRUE, fig.height=3}
advert_b <- new_data(insurance, 20) |>
  mutate(TVadverts = 8)
forecast(fit_best, advert_b) |> autoplot(insurance)
```

## Example: Insurance quotes and TV adverts

```{r, echo=TRUE, fig.height=3}
advert_c <- new_data(insurance, 20) |>
  mutate(TVadverts = 6)
forecast(fit_best, advert_c) |> autoplot(insurance)
```
