---
title: "Ch2. Time series graphics"
author: "2.1 tsibble objects"
date: "OTexts.org/fpp3/"
classoption: aspectratio=169
titlepage: fpp3title.png
titlecolor: fpp3red
toc: false
output:
  binb::monash:
    colortheme: monashwhite
    fig_width: 7.5
    fig_height: 3
    keep_tex: no
    includes:
      in_header: header.tex
---

```{r setup, include=FALSE}
source("setup.R")

global_economy <- global_economy |>
  select(Year, Country, GDP, Imports, Exports, Population)
tourism <- tourism |>
  mutate(
    State = recode(State,
      "Australian Capital Territory" = "ACT",
      "New South Wales" = "NSW",
      "Northern Territory" = "NT",
      "Queensland" = "QLD",
      "South Australia" = "SA",
      "Tasmania" = "TAS",
      "Victoria" = "VIC",
      "Western Australia" = "WA"
    )
  )
```

## `tsibble` objects

\fontsize{10}{11.2}\sf

```{r, echo = TRUE}
global_economy
```

\only<2->{\begin{textblock}{.75}(2.15,3.7)
\begin{alertblock}{}\fontsize{10}{10}\sf Index\phantom{dg}\end{alertblock}
\end{textblock}}
\only<3->{\begin{textblock}{1.6}(3.28,3.7)
\begin{alertblock}{}\fontsize{10}{10}\sf Key\phantom{dg}\end{alertblock}
\end{textblock}}
\only<4>{\begin{textblock}{6.7}(5.5,3.7)
\begin{alertblock}{}\fontsize{10}{10}\sf Measured variables\phantom{dg}\end{alertblock}
\end{textblock}}

## `tsibble` objects

\fontsize{10}{11.3}\sf

```{r, echo = TRUE}
tourism
```

\only<2->{\begin{textblock}{1.1}(2.1,3.65)
\begin{alertblock}{}\fontsize{10}{10}\sf Index\phantom{dg}\end{alertblock}
\end{textblock}}
\only<3->{\begin{textblock}{3.9}(3.65,3.65)
\begin{alertblock}{}\fontsize{10}{10}\sf Keys\phantom{dg}\end{alertblock}
\end{textblock}}
\only<4-5>{\begin{textblock}{1.5}(7.95,3.65)
\begin{alertblock}{}\fontsize{10}{10}\sf Measure\phantom{dg}\end{alertblock}
\end{textblock}}

\only<5>{\begin{textblock}{3}(9,5)
\begin{block}{}\fontsize{10}{10}\sf Domestic visitor nights in thousands by state/region and purpose.\phantom{dg}\end{block}
\end{textblock}}

## `tsibble` objects

* A `tsibble` allows storage and manipulation of multiple time series in R.

* It contains:

  + An index: time information about the observation
  + Measured variable(s): numbers of interest
  + Key variable(s): optional unique identifiers for each series

* It works with tidyverse functions.

## The `tsibble` index

### Example
\fontsize{11}{12}\sf

```{r tstable, cache=TRUE}
mydata <- tsibble(
    year = 2012:2016,
    y = c(123, 39, 78, 52, 110),
    index = year
)
mydata
```

## The `tsibble` index

### Example
\fontsize{11}{12}\sf

```{r astsibble, cache=TRUE}
mydata <- tibble(
    year = 2012:2016,
    y = c(123, 39, 78, 52, 110)
  ) |>
  as_tsibble(index = year)
mydata
```

## The `tsibble` index

\begin{block}{}
For observations more frequent than once per year, we need to use a time class function on the index.
\end{block}
\fontsize{12}{13}\sf

```{r tstablemonth, echo=FALSE}
z <- tibble(Month = paste(2019, month.abb[1:5]), Observation = c(50, 23, 34, 30, 25))
# knitr::kable(z, booktabs=TRUE)
```

```{r tstablemonth2}
z
```

## The `tsibble` index

\begin{block}{}
For observations more frequent than once per year, we need to use a time class function on the index.
\end{block}
\fontsize{12}{13}\sf

```{r month-tsibble}
z |>
  mutate(Month = yearmonth(Month)) |>
  as_tsibble(index = Month)
```

## The `tsibble` index

Common time index variables can be created with these functions:

###
```{r tstable2, echo=FALSE}
tribble(
  ~`Frequency`, ~Function,
  "Quarterly", "`yearquarter()`",
  "Monthly", "`yearmonth()`",
  "Weekly", "`yearweek()`",
  "Daily", "`as_date()`, `ymd()`",
  "Sub-daily", "`as_datetime()`"
) |>
  knitr::kable(booktabs = TRUE)
```

## Australian prison population

\fullwidth{Beechworth_prison}

## Read a csv file and convert to a tsibble {-}
\fontsize{10}{11}\sf

```{r prison}
prison <- readr::read_csv("data/prison_population.csv")
```
```{r prison2a, dependson="prison", echo=FALSE}
prison
```
## Read a csv file and convert to a tsibble {-}
\fontsize{10}{11}\sf

```{r prison3}
prison <- readr::read_csv("data/prison_population.csv") |>
  mutate(Quarter = yearquarter(date))
```

```{r prison3a, dependson="prison3", echo=FALSE}
prison
```

## Read a csv file and convert to a tsibble {-}
\fontsize{10}{11}\sf

```{r prison4}
prison <- readr::read_csv("data/prison_population.csv") |>
  mutate(Quarter = yearquarter(date)) |>
  select(-date)
```

```{r prison4a, dependson="prison4", echo=FALSE}
prison
```

## Read a csv file and convert to a tsibble {-}
\fontsize{10}{11}\sf

```{r prison5}
prison <- readr::read_csv("data/prison_population.csv") |>
  mutate(Quarter = yearquarter(date)) |>
  select(-date) |>
  as_tsibble(
    index = Quarter,
    key = c(state, gender, legal, indigenous)
  )
```

```{r prison5a, dependson="prison5", echo=FALSE}
prison
```

## Australian Pharmaceutical Benefits Scheme

\fullwidth{pills}

## Australian Pharmaceutical Benefits Scheme
\begin{block}{}
The \alert{Pharmaceutical Benefits Scheme} (PBS) is the Australian government drugs subsidy scheme.
\end{block}
\pause\fontsize{13.3}{15}\sf

  * Many drugs bought from pharmacies are subsidised to allow more equitable access to modern drugs.
  * The cost to government is determined by the number and types of drugs purchased. Currently nearly 1\% of GDP.
  * The total cost is budgeted based on forecasts of drug usage.
  * Costs are disaggregated by drug type (ATC1 x`r length(unique(PBS$ATC1))` / ATC2 `r length(unique(PBS$ATC2))`), concession category (x`r length(unique(PBS$Concession))`) and patient type (x`r length(unique(PBS$Type))`), giving $84\times2\times2=`r 84*2*2`$ time series.

## Working with `tsibble` objects {-}
\fontsize{8}{10}\sf

```{r wide, include=FALSE}
options(width = 78)
```

```{r pbs1, dependson='wide'}
PBS
```

## Working with `tsibble` objects {-}
\fontsize{12}{14}\sf

We can use the `filter()` function to select rows.

\fontsize{8}{10}\sf

```{r pbs2}
PBS |>
  filter(ATC2 == "A10")
```

## Working with `tsibble` objects {-}
\fontsize{12}{14}\sf

We can use the `select()` function to select columns.

\fontsize{8}{10}\sf

```{r pbs3}
PBS |>
  filter(ATC2 == "A10") |>
  select(Month, Concession, Type, Cost)
```

## Working with `tsibble` objects {-}
\fontsize{12}{14}\sf

We can use the `summarise()` function to summarise over keys.

\fontsize{8}{10}\sf

```{r pbs4}
PBS |>
  filter(ATC2 == "A10") |>
  select(Month, Concession, Type, Cost) |>
  summarise(TotalC = sum(Cost))
```

## Working with `tsibble` objects {-}
\fontsize{12}{14}\sf

We can use the `mutate()` function to create new variables.

\fontsize{8}{10}\sf

```{r pbs5}
PBS |>
  filter(ATC2 == "A10") |>
  select(Month, Concession, Type, Cost) |>
  summarise(TotalC = sum(Cost)) |>
  mutate(Cost = TotalC / 1e6)
```

## Working with `tsibble` objects {-}
\fontsize{12}{14}\sf

We can use the `mutate()` function to create new variables.

\fontsize{8}{10}\sf

```{r pbs6}
PBS |>
  filter(ATC2 == "A10") |>
  select(Month, Concession, Type, Cost) |>
  summarise(TotalC = sum(Cost)) |>
  mutate(Cost = TotalC / 1e6) -> a10
```

```{r a10, echo=FALSE, dependson="pbs6"}
a10
```
