---
title: "2. Time series graphics"
author: "2.3 Time series patterns"
date: "OTexts.org/fpp3/"
classoption: aspectratio=169
titlepage: fpp3title.png
titlecolor: fpp3red
toc: false
output:
  binb::monash:
    colortheme: monashwhite
    fig_width: 7.5
    fig_height: 2.8
    keep_tex: no
    includes:
      in_header: fpp3header.tex
---

```{r setup, include=FALSE}
source("setup.R")
```

## Time series patterns

Trend
  : pattern exists when there is a long-term increase or decrease in the data.

Seasonal
  : pattern exists when a series is influenced by seasonal factors (e.g., the quarter of the year, the month, or day of the week).

Cyclic
  : pattern exists when data exhibit rises and falls that are \emph{not of fixed period} (duration usually of at least 2 years).

## Time series components

### Differences between seasonal and cyclic patterns:

* seasonal pattern constant length; cyclic pattern variable length
* average length of cycle longer than length of seasonal pattern
* magnitude of cycle more variable than magnitude of seasonal pattern

## Time series patterns
\fontsize{9}{9}\sf

```{r}
aus_production |>
  filter(year(Quarter) >= 1980) |>
  autoplot(Electricity) +
  labs(y = "GWh", title = "Australian electricity production")
```

## Time series patterns
\fontsize{9}{9}\sf

```{r, warning=FALSE}
aus_production |>
  autoplot(Bricks) +
  labs(y = "million units", title = "Australian clay brick production")
```

## Time series patterns
\fontsize{9}{9}\sf

```{r}
us_employment |>
  filter(Title == "Retail Trade", year(Month) >= 1980) |>
  autoplot(Employed / 1e3) +
  labs(y = "Million people", title = "Retail employment, USA")
```

## Time series patterns
\fontsize{9}{9}\sf

```{r}
gafa_stock |>
  filter(Symbol == "AMZN", year(Date) >= 2018) |>
  autoplot(Close) +
  labs(y = "$US", title = "Amazon closing stock price")
```

## Time series patterns
\fontsize{9}{9}\sf

```{r}
pelt |>
  autoplot(Lynx) +
  labs(y = "Number trapped", title = "Annual Canadian Lynx Trappings")
```

## Seasonal or cyclic?

\alert{Differences between seasonal and cyclic patterns:}

  * seasonal pattern constant length; cyclic pattern variable length
  * average length of cycle longer than length of seasonal pattern
  * magnitude of cycle more variable than magnitude of seasonal pattern

\pause

\begin{alertblock}{}
The timing of peaks and troughs is predictable with seasonal data, but unpredictable in the long term with cyclic data.
\end{alertblock}
