---
title: "2. Time series graphics"
author: "2.4 Seasonal plots"
date: "OTexts.org/fpp3/"
classoption: aspectratio=169
titlepage: fpp3title.png
titlecolor: fpp3red
toc: false
output:
  binb::monash:
    colortheme: monashwhite
    fig_width: 7.5
    fig_height: 3
    keep_tex: no
    includes:
      in_header: header.tex
---

```{r setup, include=FALSE}
source("setup.R")
PBS |>
  filter(ATC2 == "A10") |>
  select(Month, Concession, Type, Cost) |>
  summarise(TotalC = sum(Cost)) |>
  mutate(Cost = TotalC / 1e6) -> a10
```

## Seasonal plots

```{r, echo=TRUE, dependson="a10", fig.height=3.3}
a10 |> gg_season(Cost, labels = "both") +
  labs(y = "$ million", title = "Seasonal plot: antidiabetic drug sales")
```

## Seasonal plots

  * Data plotted against the individual "seasons" in which the data were observed.  (In this case a "season" is a month.)
  * Something like a time plot except that the data from each season are overlapped.
  * Enables the underlying seasonal pattern to be seen more clearly, and also allows any substantial departures from the seasonal pattern to be easily identified.
  * In R: `gg_season()`

## Quarterly Australian Beer Production

```{r fig.height=2.8}
beer <- aus_production |>
  select(Quarter, Beer) |>
  filter(year(Quarter) >= 1992)
beer |> autoplot(Beer)
```

## Quarterly Australian Beer Production

```{r}
beer |> gg_season(Beer, labels = "right")
```

## Multiple seasonal periods
\fontsize{9}{9}\sf

```{r}
vic_elec
```

## Multiple seasonal periods

```{r, dev = "png", dpi = 180}
vic_elec |> gg_season(Demand)
```

## Multiple seasonal periods

```{r, dev = "png", dpi = 180}
vic_elec |> gg_season(Demand, period = "week")
```

## Multiple seasonal periods

```{r, dev = "png", dpi = 180}
vic_elec |> gg_season(Demand, period = "day")
```
