---
title: "Ch2. Time series graphics"
author: "2.5 Seasonal subseries plots"
date: "OTexts.org/fpp3/"
classoption: aspectratio=169
titlepage: fpp3title.png
titlecolor: fpp3red
toc: false
output:
  binb::monash:
    colortheme: monashwhite
    fig_width: 7.5
    fig_height: 3
    keep_tex: no
    includes:
      in_header: header.tex
---

```{r setup, include=FALSE}
source("setup.R")
PBS |>
  filter(ATC2 == "A10") |>
  select(Month, Concession, Type, Cost) |>
  summarise(TotalC = sum(Cost)) |>
  mutate(Cost = TotalC / 1e6) -> a10

tourism <- tourism |>
  mutate(
    State = recode(State,
      "Australian Capital Territory" = "ACT",
      "New South Wales" = "NSW",
      "Northern Territory" = "NT",
      "Queensland" = "QLD",
      "South Australia" = "SA",
      "Tasmania" = "TAS",
      "Victoria" = "VIC",
      "Western Australia" = "WA"
    )
  )
```

## Seasonal subseries plots
\fontsize{10}{10}\sf

```{r, echo=TRUE, dependson="a10"}
a10 |>
  gg_subseries(Cost) +
  labs(y = "$ million", title = "Subseries plot: antidiabetic drug sales")
```

## Seasonal subseries plots

  * Data for each season collected together in time plot as separate time series.
  * Enables the underlying seasonal pattern to be seen clearly, and changes in seasonality over time to be visualized.
  * In R: `gg_subseries()`

## Quarterly Australian Beer Production

```{r fig.height = 2.8}
beer <- aus_production |>
  select(Quarter, Beer) |>
  filter(year(Quarter) >= 1992)
beer |> autoplot(Beer)
```

## Quarterly Australian Beer Production

```{r}
beer |> gg_subseries(Beer)
```

## Australian holidays

```{r holidays}
holidays <- tourism |>
  filter(Purpose == "Holiday") |>
  group_by(State) |>
  summarise(Trips = sum(Trips))
```

\fontsize{10}{10}\sf

```{r, echo=FALSE}
holidays
```

## Australian holidays

```{r holidays-plot, echo=TRUE, dependson="holidays"}
holidays |> autoplot(Trips) +
  labs(y = "thousands of trips", title = "Australian domestic holiday nights")
```

## Seasonal plots

```{r graphics1, fig.width=5, fig.height=4.7, out.width="45%"}
holidays |> gg_season(Trips) +
  labs(y = "thousands of trips", title = "Australian domestic holiday nights")
```

## Seasonal subseries plots

```{r graphics2, fig.height=3.1}
holidays |>
  gg_subseries(Trips) +
  labs(y = "thousands of trips", title = "Australian domestic holiday nights")
```
