---
title: "2. Time series graphics"
author: "2.8 Autocorrelation"
date: "OTexts.org/fpp3/"
classoption: aspectratio=169
titlepage: fpp3title.png
titlecolor: fpp3red
toc: false
output:
  binb::monash:
    colortheme: monashwhite
    fig_width: 7.5
    fig_height: 3
    keep_tex: no
    includes:
      in_header: header.tex
---

```{r setup, include=FALSE}
source("setup.R")
library(patchwork)
new_production <- aus_production |>
  filter(year(Quarter) >= 1992)
```

## Example: Beer production

```{r, fig.height=6, fig.width=6, out.width="6.5cm"}
new_production |> gg_lag(Beer, geom = "point")
```

\only<2>{\begin{textblock}{5.8}(9,2.5)\fontsize{12}{14}\sf
\begin{block}{}
\begin{itemize}\tightlist
\item Each graph shows $y_t$ plotted against $y_{t-k}$ for
different values of $k$.
\item The autocorrelations are the correlations associated
with these scatterplots.
\item $r_1=\text{Correlation}(y_{t}, y_{t-1})$
\item $r_2=\text{Correlation}(y_{t}, y_{t-2})$
\item $r_3=\text{Correlation}(y_{t}, y_{t-3})$ \\ \hspace*{0.5cm} $\vdots$
\end{itemize}
\end{block}
\end{textblock}}

## Autocorrelation

We denote the sample autocovariance at lag $k$ by $c_k$ and the sample autocorrelation at lag $k$ by $r_k$.  Then define

\begin{block}{}
\begin{align*}
c_k &= \frac{1}{T}\sum_{t=k+1}^T (y_t-\bar{y})(y_{t-k}-\bar{y}) \\[0.cm]
\text{and}\qquad
r_{k} &= c_k/c_0
\end{align*}
\end{block}\pause

  * $r_1$ indicates how successive values of $y$ relate to each other
  * $r_2$ indicates how $y$ values two periods apart relate to each other
  * $r_k$ is \textit{almost} the same as the sample correlation between $y_t$ and $y_{t-k}$.

## Autocorrelation

Results for first 9 lags for beer data:

\fontsize{11}{13}\sf

```{r, echo=TRUE}
new_production |> ACF(Beer, lag_max = 9)
```

## Autocorrelation

Results for first 9 lags for beer data:

```{r beeracf, fig.height=1.8}
new_production |> ACF(Beer, lag_max = 9) |> autoplot()
```

\vspace*{-0.2cm}

  * Together, the autocorrelations at lags 1, 2, \dots, make up the \emph{autocorrelation} or ACF.
  * The plot is known as a **correlogram**

\vspace*{10cm}

## Autocorrelation

\fontsize{14}{14}\sf

```{r beeracf2, fig.height=1.8}
new_production |> ACF(Beer) |> autoplot()
```

  * $r_{4}$  higher than for the other lags due to **the seasonal pattern in the data**: peaks tend to be **4 quarters** apart and troughs tend to be **4 quarters** apart.
  * $r_2$ is more negative than for the other lags because troughs tend to be 2 quarters behind peaks.

## Autocorrelation functions

\only<1>{\centering\includegraphics[width=19cm, height=7.4cm]{acf_1}}
\only<2>{\centering\includegraphics[width=19cm, height=7.4cm]{acf_2}}
\only<3>{\centering\includegraphics[width=19cm, height=7.4cm]{acf_3}}
\only<4>{\centering\includegraphics[width=19cm, height=7.4cm]{acf_4}}
\only<5>{\centering\includegraphics[width=19cm, height=7.4cm]{acf_5}}
\only<6>{\centering\includegraphics[width=19cm, height=7.4cm]{acf_6}}
\only<7>{\centering\includegraphics[width=19cm, height=7.4cm]{acf_7}}
\only<8>{\centering\includegraphics[width=19cm, height=7.4cm]{acf_8}}
\only<9>{\centering\includegraphics[width=19cm, height=7.4cm]{acf_9}}

\vspace*{10cm}

\begin{textblock}{3}(0.3,8.7)\fontsize{7}{9}\sf
Artwork by @allison\_horst
\end{textblock}

## Trend and seasonality in ACF plots

- When data have a trend, the autocorrelations for small lags tend to be large and positive.
- When data are seasonal, the autocorrelations will be larger at the seasonal lags (i.e., at multiples of the seasonal frequency)
- When data are trended and seasonal, you see a combination of these effects.

## US retail trade employment
\fontsize{10}{10}\sf

```{r}
retail <- us_employment |>
  filter(Title == "Retail Trade", year(Month) >= 1980)
retail |> autoplot(Employed)
```

## US retail trade employment
\fontsize{10}{10}\sf

```{r}
retail |>
  ACF(Employed, lag_max = 48) |>
  autoplot()
```

## Google stock price
\fontsize{10}{10}\sf

```{r}
google_2015 <- gafa_stock |>
  filter(Symbol == "GOOG", year(Date) == 2015) |>
  select(Date, Close)
google_2015
```

## Google stock price
\fontsize{10}{10}\sf

```{r}
google_2015 |> autoplot(Close)
```

## Google stock price
\fontsize{10}{10}\sf

```{r}
google_2015 |>
  ACF(Close, lag_max = 100)
```

## Google stock price
\fontsize{10}{10}\sf

```{r}
google_2015 |>
  ACF(Close, lag_max = 100) |>
  autoplot()
```
