---
title: "2. Time series graphics"
author: "2.9 White noise"
date: "OTexts.org/fpp3/"
classoption: aspectratio=169
titlepage: fpp3title.png
titlecolor: fpp3red
toc: false
output:
  binb::monash:
    colortheme: monashwhite
    fig_width: 7.5
    fig_height: 3
    keep_tex: no
    includes:
      in_header: header.tex
---

```{r setup, include=FALSE}
source("setup.R")
```

## Example: White noise
\fontsize{10}{10}\sf

```{r wn}
set.seed(30)
wn <- tsibble(t = 1:50, y = rnorm(50), index = t)
wn |> autoplot(y)
```

\only<2>{
\begin{textblock}{10}(.4,6.6)\fontsize{12}{13}\sf
\begin{alertblock}{}
White noise data is uncorrelated across time with zero mean and constant variance.

(Technically, we require independence as well.)
\end{alertblock}
\end{textblock}}

\vspace*{10cm}

## Example: White noise
\fontsize{10}{10}\sf

```r
wn |> ACF(y)
```

\fontsize{10}{10}\sf\tabcolsep=0.1cm

```{r wnacf, echo=FALSE, dependson="wn"}
wn |>
  ACF(y, lag_max = 10) |>
  as_tibble() |>
  mutate(lag = as.numeric(lag)) |>
  pivot_wider(names_from = lag, values_from = acf) |>
  rename_all(function(x) {
    paste("$r_{", x, "}$", sep = "")
  }) |>
  knitr::kable(
    booktabs = TRUE,
    escape = FALSE, align = "c", digits = 3,
    format.args = list(nsmall = 3)
  )
```

```{r, echo=FALSE, fig.height=1.5}
wn |>
  ACF(y) |>
  autoplot()
```

\pause

 * Sample autocorrelations for white noise series.
 * Expect each autocorrelation to be close to zero.
 * Blue lines show 95% critical values.

\vspace*{10cm}

## \large Sampling distribution of autocorrelations

Sampling distribution of $r_k$ for white noise data is asymptotically N(0,$1/T$).\pause

  *  95% of all $r_k$ for white noise must lie within $\pm 1.96/\sqrt{T}$.
  * If this is not the case, the series is probably not WN.
  * Common to plot lines at $\pm 1.96/\sqrt{T}$ when plotting ACF.
These are the \alert{critical values}.

## Example: Pigs slaughtered
\fontsize{10}{10}\sf

```{r, fig.height=2.7}
pigs <- aus_livestock |>
  filter(State == "Victoria", Animal == "Pigs", year(Month) >= 2014)
pigs |> autoplot(Count / 1e3) +
  labs(y = "Thousands", title = "Number of pigs slaughtered in Victoria")
```

## Example: Pigs slaughtered
\fontsize{10}{10}\sf

```{r}
pigs |>
  ACF(Count) |>
  autoplot()
```

## Example: Pigs slaughtered

Monthly total number of pigs slaughtered
in the state of Victoria, Australia, from January 2014 through December 2018
(Source: Australian Bureau of Statistics.)\pause

  * Difficult to detect pattern in time plot.
  * ACF shows significant autocorrelation for lag 2 and 12.
  * Indicate some slight seasonality.

\pause

These show the series is **not a white noise series**.
