---
title: "ETC3550/ETC5550 Applied&nbsp;forecasting"
author: "Ch2. Time series graphics"
date: "OTexts.org/fpp3/"
classoption: aspectratio=169
titlepage: title16x9.png
titlecolor: burntorange
toc: true
output:
  binb::monash:
    colortheme: monashwhite
    fig_width: 7.5
    fig_height: 3.5
    keep_tex: no
    includes:
      in_header: header.tex
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  message = FALSE,
  warning = FALSE,
  cache = TRUE,
  dev.args = list(pointsize = 11)
)
options(
  digits = 3,
  width = 80,
  ggplot2.continuous.colour="viridis",
  ggplot2.continuous.fill = "viridis",
  ggplot2.discrete.colour = c("#D55E00", "#0072B2","#009E73", "#CC79A7", "#E69F00", "#56B4E9", "#F0E442"),
  ggplot2.discrete.fill = c("#D55E00", "#0072B2","#009E73", "#CC79A7", "#E69F00", "#56B4E9", "#F0E442")
)
library(fpp3)
library(patchwork)

global_economy <- global_economy %>%
  select(Year, Country, GDP, Imports, Exports, Population)
tourism <- tourism %>%
  mutate(
    State = recode(State,
      "Australian Capital Territory" = "ACT",
      "New South Wales" = "NSW",
      "Northern Territory" = "NT",
      "Queensland" = "QLD",
      "South Australia" = "SA",
      "Tasmania" = "TAS",
      "Victoria" = "VIC",
      "Western Australia" = "WA"
    )
  )
```

# Time series in R

## `tsibble` objects

\fontsize{10}{11.2}\sf

```{r, echo = TRUE}
global_economy
```

\only<2->{\begin{textblock}{.75}(2.15,3.75)
\begin{alertblock}{}\fontsize{10}{10}\sf Index\phantom{dg}\end{alertblock}
\end{textblock}}
\only<3->{\begin{textblock}{1.6}(3.28,3.75)
\begin{alertblock}{}\fontsize{10}{10}\sf Key\phantom{dg}\end{alertblock}
\end{textblock}}
\only<4>{\begin{textblock}{6.7}(5.5,3.75)
\begin{alertblock}{}\fontsize{10}{10}\sf Measured variables\phantom{dg}\end{alertblock}
\end{textblock}}

## `tsibble` objects

\fontsize{10}{11.3}\sf

```{r, echo = TRUE}
tourism
```

\only<2->{\begin{textblock}{1.1}(2.1,3.7)
\begin{alertblock}{}\fontsize{10}{10}\sf Index\phantom{dg}\end{alertblock}
\end{textblock}}
\only<3->{\begin{textblock}{3.9}(3.65,3.7)
\begin{alertblock}{}\fontsize{10}{10}\sf Keys\phantom{dg}\end{alertblock}
\end{textblock}}
\only<4-5>{\begin{textblock}{1.5}(7.95,3.7)
\begin{alertblock}{}\fontsize{10}{10}\sf Measure\phantom{dg}\end{alertblock}
\end{textblock}}

\only<5>{\begin{textblock}{3}(9,5)
\begin{block}{}\fontsize{10}{10}\sf Domestic visitor nights in thousands by state/region and purpose.\phantom{dg}\end{block}
\end{textblock}}

## `tsibble` objects

* A `tsibble` allows storage and manipulation of multiple time series in R.

* It contains:

  + An index: time information about the observation
  + Measured variable(s): numbers of interest
  + Key variable(s): optional unique identifiers for each series

* It works with tidyverse functions.

## The `tsibble` index

### Example
\fontsize{11}{12}\sf

```{r tstable, cache=TRUE}
mydata <- tsibble(
    year = 2012:2016,
    y = c(123, 39, 78, 52, 110),
    index = year
)
mydata
```

## The `tsibble` index

### Example
\fontsize{11}{12}\sf

```{r astsibble, cache=TRUE}
mydata <- tibble(
    year = 2012:2016,
    y = c(123, 39, 78, 52, 110)
  ) %>%
  as_tsibble(index = year)
mydata
```

## The `tsibble` index

\begin{block}{}
For observations more frequent than once per year, we need to use a time class function on the index.
\end{block}
\fontsize{12}{13}\sf

```{r tstablemonth, echo=FALSE}
z <- tibble(Month = paste(2019, month.abb[1:5]), Observation = c(50, 23, 34, 30, 25))
# knitr::kable(z, booktabs=TRUE)
```

```{r tstablemonth2}
z
```

## The `tsibble` index

\begin{block}{}
For observations more frequent than once per year, we need to use a time class function on the index.
\end{block}
\fontsize{12}{13}\sf

```{r month-tsibble}
z %>%
  mutate(Month = yearmonth(Month)) %>%
  as_tsibble(index = Month)
```

## The `tsibble` index

Common time index variables can be created with these functions:

###
```{r tstable2, echo=FALSE}
tribble(
  ~`Frequency`, ~Function,
  "Annual", "`start:end`",
  "Quarterly", "`yearquarter()`",
  "Monthly", "`yearmonth()`",
  "Weekly", "`yearweek()`",
  "Daily", "`as_date()`, `ymd()`",
  "Sub-daily", "`as_datetime()`"
) %>%
  knitr::kable(booktabs = TRUE)
```

# Example: Australian prison population

## Australian prison population

\fullwidth{Beechworth_prison}

## Read a csv file and convert to a tsibble {-}
\fontsize{10}{11}\sf

```{r prison}
prison <- readr::read_csv("data/prison_population.csv")
```
```{r prison2a, dependson="prison", echo=FALSE}
prison
```
## Read a csv file and convert to a tsibble {-}
\fontsize{10}{11}\sf

```{r prison3}
prison <- readr::read_csv("data/prison_population.csv") %>%
  mutate(Quarter = yearquarter(date))
```

```{r prison3a, dependson="prison3", echo=FALSE}
prison
```

## Read a csv file and convert to a tsibble {-}
\fontsize{10}{11}\sf

```{r prison4}
prison <- readr::read_csv("data/prison_population.csv") %>%
  mutate(Quarter = yearquarter(date)) %>%
  select(-date)
```

```{r prison4a, dependson="prison4", echo=FALSE}
prison
```

## Read a csv file and convert to a tsibble {-}
\fontsize{10}{11}\sf

```{r prison5}
prison <- readr::read_csv("data/prison_population.csv") %>%
  mutate(Quarter = yearquarter(date)) %>%
  select(-date) %>%
  as_tsibble(
    index = Quarter,
    key = c(state, gender, legal, indigenous)
  )
```

```{r prison5a, dependson="prison5", echo=FALSE}
prison
```

# Example: Australian pharmaceutical sales

## Australian Pharmaceutical Benefits Scheme

\fullwidth{pills}

## Australian Pharmaceutical Benefits Scheme
\begin{block}{}
The \alert{Pharmaceutical Benefits Scheme} (PBS) is the Australian government drugs subsidy scheme.
\end{block}
\pause\fontsize{13.3}{15}\sf

  * Many drugs bought from pharmacies are subsidised to allow more equitable access to modern drugs.
  * The cost to government is determined by the number and types of drugs purchased. Currently nearly 1\% of GDP.
  * The total cost is budgeted based on forecasts of drug usage.
  * Costs are disaggregated by drug type (ATC1 x`r length(unique(PBS$ATC1))` / ATC2 `r length(unique(PBS$ATC2))`), concession category (x`r length(unique(PBS$Concession))`) and patient type (x`r length(unique(PBS$Type))`), giving $84\times2\times2=`r 84*2*2`$ time series.

## Working with `tsibble` objects {-}
\fontsize{8}{10}\sf

```{r wide, include=FALSE}
options(width = 78)
```

```{r pbs1, dependson='wide'}
PBS
```

## Working with `tsibble` objects {-}
\fontsize{12}{14}\sf

We can use the `filter()` function to select rows.

\fontsize{8}{10}\sf

```{r pbs2}
PBS %>%
  filter(ATC2 == "A10")
```

## Working with `tsibble` objects {-}
\fontsize{12}{14}\sf

We can use the `select()` function to select columns.

\fontsize{8}{10}\sf

```{r pbs3}
PBS %>%
  filter(ATC2 == "A10") %>%
  select(Month, Concession, Type, Cost)
```

## Working with `tsibble` objects {-}
\fontsize{12}{14}\sf

We can use the `summarise()` function to summarise over keys.

\fontsize{8}{10}\sf

```{r pbs4}
PBS %>%
  filter(ATC2 == "A10") %>%
  select(Month, Concession, Type, Cost) %>%
  summarise(total_cost = sum(Cost))
```

## Working with `tsibble` objects {-}
\fontsize{12}{14}\sf

We can use the `mutate()` function to create new variables.

\fontsize{8}{10}\sf

```{r pbs5}
PBS %>%
  filter(ATC2 == "A10") %>%
  select(Month, Concession, Type, Cost) %>%
  summarise(total_cost = sum(Cost)) %>%
  mutate(total_cost = total_cost / 1e6)
```

## Working with `tsibble` objects {-}
\fontsize{12}{14}\sf

We can use the `mutate()` function to create new variables.

\fontsize{8}{10}\sf

```{r pbs6}
PBS %>%
  filter(ATC2 == "A10") %>%
  select(Month, Concession, Type, Cost) %>%
  summarise(total_cost = sum(Cost)) %>%
  mutate(total_cost = total_cost / 1e6) -> a10
```

```{r a10, echo=FALSE, dependson="pbs6"}
a10
```

```{r narrow, include=FALSE}
options(width = 60)
```

# Time plots

## Time plots
\fontsize{11}{12}\sf

```{r, echo=TRUE}
a10 %>%
  autoplot(total_cost)
```

## Ansett airlines
\fontsize{11}{12}\sf

```{r, echo=TRUE}
ansett %>%
  autoplot(Passengers)
```

## Ansett airlines
\fontsize{11}{12}\sf

```{r, echo=TRUE}
ansett %>%
  filter(Class == "Economy") %>%
  autoplot(Passengers)
```

## Ansett airlines
\fontsize{11}{12}\sf

```{r, echo=TRUE}
ansett %>%
  filter(Airports == "MEL-SYD") %>%
  autoplot(Passengers)
```

## Time series patterns

Trend
  : pattern exists when there is a long-term increase or decrease in the data.

Seasonal
  : pattern exists when a series is influenced by seasonal factors (e.g., the quarter of the year, the month, or day of the week).

Cyclic
  : pattern exists when data exhibit rises and falls that are \emph{not of fixed period} (duration usually of at least 2 years).

## Time series components

### Differences between seasonal and cyclic patterns:

* seasonal pattern constant length; cyclic pattern variable length
* average length of cycle longer than length of seasonal pattern
* magnitude of cycle more variable than magnitude of seasonal pattern


## Time series patterns
\fontsize{10}{10}\sf

```{r, fig.height=3}
aus_production %>%
  filter(year(Quarter) >= 1980) %>%
  autoplot(Electricity) +
  labs(y = "GWh",
       title = "Australian electricity production")
```

## Time series patterns
\fontsize{10}{10}\sf

```{r, warning=FALSE, fig.height=3}
aus_production %>%
  autoplot(Bricks) +
  labs(y = "million units",
       title = "Australian clay brick production")
```

## Time series patterns
\fontsize{10}{10}\sf

```{r, fig.height=3}
us_employment %>%
  filter(Title == "Retail Trade", year(Month) >= 1980) %>%
  autoplot(Employed / 1e3) +
  labs(y = "Million people",
       title = "Retail employment, USA")
```

## Time series patterns
\fontsize{10}{10}\sf

```{r, fig.height=3}
gafa_stock %>%
  filter(Symbol == "AMZN", year(Date) >= 2018) %>%
  autoplot(Close) +
  labs(y = "$US",
       title = "Amazon closing stock price")
```

## Time series patterns
\fontsize{10}{10}\sf

```{r, fig.height=3}
pelt %>%
  autoplot(Lynx) +
  labs(y="Number trapped",
       title = "Annual Canadian Lynx Trappings")
```

## Seasonal or cyclic?

\alert{Differences between seasonal and cyclic patterns:}

  * seasonal pattern constant length; cyclic pattern variable length
  * average length of cycle longer than length of seasonal pattern
  * magnitude of cycle more variable than magnitude of seasonal pattern

\pause

\begin{alertblock}{}
The timing of peaks and troughs is predictable with seasonal data, but unpredictable in the long term with cyclic data.
\end{alertblock}

# Seasonal and subseries plots

## Seasonal plots

\footnotesize

```{r, echo=TRUE, dependson="a10"}
a10 %>% gg_season(total_cost, labels = "both") +
  labs(y = "$ million",
       title = "Seasonal plot: antidiabetic drug sales")
```

## Seasonal plots

  * Data plotted against the individual "seasons" in which the data were observed.  (In this case a "season" is a month.)
  * Something like a time plot except that the data from each season are overlapped.
  * Enables the underlying seasonal pattern to be seen more clearly, and also allows any substantial departures from the seasonal pattern to be easily identified.
  * In R: `gg_season()`

## Seasonal subseries plots

\fontsize{10}{11}\sf

```{r, echo=TRUE, dependson="a10"}
a10 %>%
  gg_subseries(total_cost) +
  labs(y = "$ million",
       title = "Subseries plot: antidiabetic drug sales")
```

## Seasonal subseries plots

  * Data for each season collected together in time plot as separate time series.
  * Enables the underlying seasonal pattern to be seen clearly, and changes in seasonality over time to be visualized.
  * In R: `gg_subseries()`

## Quarterly Australian Beer Production
\fontsize{13}{15}\sf

```{r, fig.height=3}
beer <- aus_production %>%
  select(Quarter, Beer) %>%
  filter(year(Quarter) >= 1992)
beer %>% autoplot(Beer)
```

## Quarterly Australian Beer Production

```{r}
beer %>% gg_season(Beer, labels="right")
```

## Quarterly Australian Beer Production

```{r}
beer %>% gg_subseries(Beer)
```

## Your turn

Look at the quarterly tourism data for the Snowy Mountains

```r
snowy <- tourism %>%
  filter(Region == "Snowy Mountains")
```

  - Use `autoplot()`, `gg_season()` and `gg_subseries()` to explore the data.
  - What do you learn?


## Multiple seasonal periods
\fontsize{10}{11}\sf

```{r}
vic_elec
```

## Multiple seasonal periods
\fontsize{12}{13}\sf

```{r, dev = "png", dpi = 180}
vic_elec %>% gg_season(Demand)
```

## Multiple seasonal periods
\fontsize{12}{13}\sf

```{r, dev = "png", dpi = 180}
vic_elec %>% gg_season(Demand, period = "week")
```

## Multiple seasonal periods
\fontsize{12}{13}\sf

```{r, dev = "png", dpi = 180}
vic_elec %>% gg_season(Demand, period = "day")
```


## Australian holidays
\fontsize{9}{10}\sf

```{r holidays}
holidays <- tourism %>%
  filter(Purpose == "Holiday") %>%
  group_by(State) %>%
  summarise(Trips = sum(Trips))
```

```{r, echo=FALSE}
holidays
```

## Australian holidays
\fontsize{9}{10}\sf

```{r holidays-plot, echo=TRUE, dependson="holidays", fig.height=3.9}
holidays %>% autoplot(Trips) +
  labs(y = "thousands of trips",
       title = "Australian domestic holiday nights")
```

## Seasonal plots
\fontsize{9}{10}\sf

```{r graphics1, fig.width=4, fig.height=5, out.width="45%"}
holidays %>% gg_season(Trips) +
  labs(y = "thousands of trips",
       title = "Australian domestic holiday nights")
```

## Seasonal subseries plots
\fontsize{9}{10}\sf

```{r graphics2, fig.height=4.1}
holidays %>%
  gg_subseries(Trips) +
    labs(y = "thousands of trips",
       title = "Australian domestic holiday nights")
```

# Lag plots and autocorrelation

## Example: Beer production
\fontsize{11}{13}\sf

```{r}
new_production <- aus_production %>%
  filter(year(Quarter) >= 1992)
new_production
```

## Example: Beer production
\fontsize{13}{15}\sf

```{r, fig.height=6, fig.width=6, out.width="6.5cm"}
new_production %>% gg_lag(Beer)
```

## Example: Beer production
\fontsize{13}{15}\sf

```{r, fig.height=6, fig.width=6, out.width="6.5cm"}
new_production %>% gg_lag(Beer, geom='point')
```


## Lagged scatterplots

  * Each graph shows $y_t$ plotted against $y_{t-k}$ for
different values of $k$.
  * The autocorrelations are the correlations associated
with these scatterplots.
  * ACF (autocorrelation function):
     * $r_1=\text{Correlation}(y_{t}, y_{t-1})$
     * $r_2=\text{Correlation}(y_{t}, y_{t-2})$
     * $r_3=\text{Correlation}(y_{t}, y_{t-3})$
     * etc.

## Autocorrelation

**Covariance** and **correlation**: measure extent of **linear relationship** between two variables ($y$ and $X$).\pause

**Autocovariance** and **autocorrelation**: measure linear relationship between **lagged values** of a time series $y$.\pause

We measure the relationship between:

  * $y_{t}$ and $y_{t-1}$
  * $y_{t}$ and $y_{t-2}$
  * $y_{t}$ and $y_{t-3}$
  * etc.

## Autocorrelation

We denote the sample autocovariance at lag $k$ by $c_k$ and the sample autocorrelation at lag $k$ by $r_k$.  Then define

\begin{block}{}
\begin{align*}
c_k &= \frac{1}{T}\sum_{t=k+1}^T (y_t-\bar{y})(y_{t-k}-\bar{y}) \\[0.cm]
\text{and}\qquad
r_{k} &= c_k/c_0
\end{align*}
\end{block}\pause\small

  * $r_1$ indicates how successive values of  $y$  relate to each other
  * $r_2$ indicates how  $y$ values two periods apart relate to each other
  * $r_k$ is \textit{almost} the same as the sample correlation between $y_t$ and $y_{t-k}$.

## Autocorrelation

Results for first 9 lags for beer data:

\fontsize{11}{13}\sf

```{r, echo=TRUE}
new_production %>% ACF(Beer, lag_max = 9)
```

## Autocorrelation

Results for first 9 lags for beer data:

\fontsize{11}{13}\sf

```{r beeracf, fig.height=2.5}
new_production %>% ACF(Beer, lag_max = 9) %>% autoplot()
```

  * Together, the autocorrelations at lags 1, 2, \dots, make up the \emph{autocorrelation} or ACF.
  * The plot is known as a **correlogram**


\vspace*{10cm}

## Autocorrelation

\fontsize{11}{13}\sf

```{r beeracf2, fig.height=2.5}
new_production %>% ACF(Beer) %>% autoplot()
```

  * $r_{4}$  higher than for the other lags. This is due to **the seasonal pattern in the data**: the peaks tend to be **4 quarters** apart and the troughs tend to be **2 quarters** apart.
  * $r_2$ is more negative than for the other lags because troughs tend to be 2 quarters behind peaks.


## Trend and seasonality in ACF plots

- When data have a trend, the autocorrelations for small lags tend to be large and positive.
- When data are seasonal, the autocorrelations will be larger at the seasonal lags (i.e., at multiples of the seasonal frequency)
- When data are trended and seasonal, you see a combination of these effects.

## Autocorrelation functions

\only<1>{\centering\includegraphics[width=17cm]{acf_1}}
\only<2>{\centering\includegraphics[width=17cm]{acf_2}}
\only<3>{\centering\includegraphics[width=17cm]{acf_3}}
\only<4>{\centering\includegraphics[width=17cm]{acf_4}}
\only<5>{\centering\includegraphics[width=17cm]{acf_5}}
\only<6>{\centering\includegraphics[width=17cm]{acf_6}}
\only<7>{\centering\includegraphics[width=17cm]{acf_7}}
\only<8>{\centering\includegraphics[width=17cm]{acf_8}}
\only<9>{\centering\includegraphics[width=17cm]{acf_9}}

\vspace*{10cm}

\begin{textblock}{9}(6,9.2)\fontsize{7}{9}\sf
Artwork by @allison\_horst
\end{textblock}

## US retail trade employment
\fontsize{11}{12}\sf

```{r}
retail <- us_employment %>%
  filter(Title == "Retail Trade", year(Month) >= 1980)
retail %>% autoplot(Employed)
```

## US retail trade employment
\fontsize{11}{12}\sf

```{r}
retail %>%
  ACF(Employed, lag_max = 48) %>%
  autoplot()
```

## Google stock price
\fontsize{12}{14}\sf

```{r google-2015}
google_2015 <- gafa_stock %>%
  filter(Symbol == "GOOG", year(Date) == 2015) %>%
  select(Date, Close)
google_2015
```

## Google stock price

```{r}
google_2015 %>% autoplot(Close)
```

## Google stock price
\fontsize{12}{16}\sf

```r
google_2015 %>%
  ACF(Close, lag_max=100)
# Error: Can't handle tsibble of irregular interval.
```

\pause

```{r}
google_2015
```

## Google stock price
\fontsize{11}{13}\sf

```{r}
google_2015 <- google_2015 %>%
  mutate(trading_day = row_number()) %>%
  update_tsibble(index = trading_day, regular = TRUE)
google_2015
```

## Google stock price

```{r}
google_2015 %>%
  ACF(Close, lag_max = 100) %>%
  autoplot()
```

## Your turn

We have introduced the following functions:

  - `gg_lag`
  - `ACF`

Use these functions to explore the following time series:

  - Bricks from `aus_production`
  - Lynx from `pelt`
  - Victorian Electricity Demand from `aus_elec`

Can you spot any seasonality, cyclicity and trend? What do you learn about the series?

## Which is which?

```{r, fig.height=6, fig.width=12, echo=FALSE, warning=FALSE, out.width="11.5cm"}
cowtemp <- as_tsibble(fma::cowtemp)
USAccDeaths <- as_tsibble(USAccDeaths)
AirPassengers <- as_tsibble(AirPassengers)
mink <- as_tsibble(fma::mink)
tp1 <- autoplot(cowtemp, value) +
  labs(x = "", y = "chirps per minute", title = "1. Daily temperature of cow")
tp2 <- autoplot(USAccDeaths, value) +
  labs(x = "", y = "thousands", title = "2. Monthly accidental deaths")
tp3 <- autoplot(AirPassengers, value) +
  labs(x = "", y = "thousands", title = "3. Monthly air passengers")
tp4 <- autoplot(mink, value) +
  labs(x = "", y = "thousands", title = "4. Annual mink trappings")
acfb <- ACF(cowtemp, value) %>% autoplot() +
  labs(x="", title="B") + ylim(-0.4,1)
acfa <- ACF(USAccDeaths, value) %>% autoplot() +
  labs(x = "", title = "A") + ylim(-0.4,1)
acfd <- ACF(AirPassengers, value) %>% autoplot() +
  labs(x = "", title = "D") + ylim(-0.4,1)
acfc <- ACF(mink, value) %>% autoplot() +
  labs(x = "", title ="C") + ylim(-0.4,1)
(tp1 | tp2 | tp3 | tp4) / (acfa | acfb | acfc | acfd)
```

# White noise

## Example: White noise
\fontsize{11}{13}\sf

```{r wn, fig.height=3.3}
set.seed(30)
wn <- tsibble(t = 1:50, y = rnorm(50), index = t)
wn %>% autoplot(y)
```

\only<2>{
\begin{textblock}{10}(1.4,7.2)\fontsize{13}{15}\sf
\begin{alertblock}{}
White noise data is uncorrelated across time with zero mean and constant variance.

(Technically, we require independence as well.)
\end{alertblock}
\end{textblock}}

\vspace*{10cm}


## Example: White noise
\fontsize{11}{11}\sf

```r
wn %>% ACF(y)
```

\fontsize{10}{11}\sf\tabcolsep=0.1cm

```{r wnacf, echo=FALSE, dependson="wn"}
wn %>%
  ACF(y, lag_max = 10) %>%
  as_tibble() %>%
  mutate(lag = as.numeric(lag)) %>%
  pivot_wider(names_from = lag, values_from = acf) %>%
  rename_all(function(x) {
    paste("$r_{", x, "}$", sep = "")
  }) %>%
  knitr::kable(
    booktabs = TRUE,
    escape = FALSE, align = "c", digits = 3,
    format.args = list(nsmall = 3)
  )
```

```{r, echo=FALSE, fig.height=2}
wn %>%
  ACF(y) %>%
  autoplot()
```

\pause\fontsize{11}{12}\sf

 * Sample autocorrelations for white noise series.
 * Expect each autocorrelation to be close to zero.
 * Blue lines show 95% critical values.

\vspace*{10cm}

## \large Sampling distribution of autocorrelations

Sampling distribution of $r_k$ for white noise data is asymptotically N(0,$1/T$).\pause

  *  95% of all $r_k$ for white noise must lie within $\pm 1.96/\sqrt{T}$.
  * If this is not the case, the series is probably not WN.
  * Common to plot lines at $\pm 1.96/\sqrt{T}$ when plotting ACF.
These are the \alert{critical values}.

## Example: Pigs slaughtered

\fontsize{10}{12}\sf

```{r, fig.height=3}
pigs <- aus_livestock %>%
  filter(State == "Victoria", Animal == "Pigs",
         year(Month) >= 2014)
pigs %>% autoplot(Count/1e3) +
  labs(y = "Thousands",
       title = "Number of pigs slaughtered in Victoria")
```

## Example: Pigs slaughtered

```{r}
pigs %>% ACF(Count) %>% autoplot()
```

## Example: Pigs slaughtered

Monthly total number of pigs slaughtered
in the state of Victoria, Australia, from January 2014 through December 2018
(Source: Australian Bureau of Statistics.)\pause

  * Difficult to detect pattern in time plot.
  * ACF shows significant autocorrelation for lag 2 and 12.
  * Indicate some slight seasonality.

\pause

These show the series is **not a white noise series**.

## Your turn

You can compute the daily changes in the Google stock price in 2018 using

\fontsize{11.5}{15}\sf


```{r, eval = FALSE}
dgoog <- gafa_stock %>%
  filter(Symbol == "GOOG", year(Date) >= 2018) %>%
  mutate(trading_day = row_number()) %>%
  update_tsibble(index=trading_day, regular=TRUE) %>%
  mutate(diff = difference(Close))
```

\fontsize{14}{16}\sf

Does `diff` look like white noise?
