---
title: "3. Time series decomposition"
author: "3.2 Time series components"
date: "OTexts.org/fpp3/"
classoption: aspectratio=169
titlepage: fpp3title.png
titlecolor: fpp3red
toc: false
output:
  binb::monash:
    colortheme: monashwhite
    fig_width: 7.5
    fig_height: 3
    keep_tex: no
    includes:
      in_header: header.tex
---

```{r setup, include=FALSE}
source("setup.R")
```

## Time series patterns

**Recall**

Trend
:  pattern exists when there is a long-term increase or decrease in the data.

Cyclic
: pattern exists when data exhibit rises and falls that are *not of fixed period* (duration usually of at least 2 years).

Seasonal
: pattern exists when a series is influenced by seasonal factors (e.g., the quarter of the year, the month, or day of the week).

## Time series decomposition

\begin{block}{}\vspace*{-0.3cm}
\[ y_t = f(S_t, T_t, R_t) \]
\end{block}
\begin{tabular}{@{}llp{8cm}@{}}
where & $y_t=$ & data at period $t$ \\
      & $T_t=$ & trend-cycle component at period $t$\\
      & $S_t=$ & seasonal component at period $t$ \\
      & $R_t=$ & remainder component at period $t$
\end{tabular}
\pause

**Additive decomposition:** $y_t = S_t + T_t + R_t.$

**Multiplicative decomposition:** $y_t = S_t \times T_t \times R_t.$

## Time series decomposition
\fontsize{13}{15}\sf

  *  Additive model  appropriate if  magnitude of  seasonal fluctuations does not vary with level.
  *  If seasonal are proportional to level of series, then multiplicative model appropriate.
  *  Multiplicative decomposition more prevalent with economic series
  *  Alternative: use a Box-Cox transformation, and then use additive decomposition.
  *  Logs turn multiplicative relationship into an additive relationship:

$$y_t = S_t \times T_t \times R_t \quad\Rightarrow\quad
\log y_t = \log S_t + \log T_t + \log R_t.
$$

## US Retail Employment
\fontsize{11}{11}\sf

```{r usretail}
us_retail_employment <- us_employment |>
  filter(year(Month) >= 1990, Title == "Retail Trade") |>
  select(-Series_ID)
us_retail_employment
```

\vspace*{10cm}

## US Retail Employment

```{r dable1}
us_retail_employment |>
  autoplot(Employed) +
  labs(y="Persons (thousands)", title="Total employment in US retail")
```

\vspace*{10cm}

## US Retail Employment
\fontsize{11}{11}\sf

```{r dable2}
us_retail_employment |>
  model(stl = STL(Employed))
```

\vspace*{10cm}

## US Retail Employment
\fontsize{11}{11}\sf

```{r dable3}
dcmp <- us_retail_employment |>
  model(stl = STL(Employed))
components(dcmp)
```

\vspace*{10cm}

## US Retail Employment

```{r usretail-stl, fig.width=8, fig.height=5}
components(dcmp) |> autoplot()
```

## US Retail Employment

```{r dable4}
us_retail_employment |>
  autoplot(Employed, color='gray') +
  autolayer(components(dcmp), trend, color='#D55E00') +
  labs(y="Persons (thousands)", title="Total employment in US retail")
```

\vspace*{10cm}

## US Retail Employment

```{r usretail3}
components(dcmp) |> gg_subseries(season_year)
```

## Seasonal adjustment

  *  Useful by-product of decomposition:  an easy way to calculate seasonally adjusted data.
  *  Additive decomposition: seasonally adjusted data given by
$$y_t - S_t = T_t + R_t$$
  *  Multiplicative decomposition: seasonally adjusted data given by
$$y_t / S_t = T_t \times R_t$$

## US Retail Employment

```{r usretail-sa}
us_retail_employment |>
  autoplot(Employed, color='gray') +
  autolayer(components(dcmp), season_adjust, color='#0072B2') +
  labs(y="Persons (thousands)", title="Total employment in US retail")
```

## Seasonal adjustment

  * We use estimates of $S$ based on past values to seasonally adjust a current value.
  *  Seasonally adjusted series reflect **remainders** as well as **trend**. Therefore they are not "smooth" and "downturns" or "upturns" can be misleading.
  *  It is better to use the trend-cycle component to look for turning points.
