---
title: "ETC3550/ETC5550 Applied&nbsp;forecasting"
author: "Ch3. Time series decomposition"
date: "OTexts.org/fpp3/"
toc: true
colortheme: monashwhite
output:
  binb::monash:
    fig_width: 7
    fig_height: 3.5
    includes:
      in_header: header.tex
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, cache=TRUE)

library(tidyverse)
library(fpp3)
library(purrr)
library(gganimate)
library(latex2exp)

options(width=65)
```

# Transformations and adjustments

## Per capita adjustments
\fontsize{13}{14}\sf

```{r gdp-per-capita}
global_economy %>%
  filter(Country == "Australia") %>%
  autoplot(GDP)
```

## Per capita adjustments
\fontsize{13}{14}\sf

```{r gdp-per-capita2}
global_economy %>%
  filter(Country == "Australia") %>%
  autoplot(GDP / Population)
```

## Your turn

Consider the GDP information in `global_economy`. Plot the GDP per capita for each country over time. Which country has the highest GDP per capita? How has this changed over time?

## Inflation adjustments
\fontsize{10}{10}\sf

```{r retail_cpi, message=FALSE, warning=FALSE, fig.show='hide'}
print_retail <- aus_retail %>%
  filter(Industry == "Newspaper and book retailing") %>%
  group_by(Industry) %>%
  index_by(Year = year(Month)) %>%
  summarise(Turnover = sum(Turnover))
aus_economy <- filter(global_economy, Code == "AUS")
print_retail %>%
  left_join(aus_economy, by = "Year") %>%
  mutate(Adj_turnover = Turnover / CPI) %>%
  pivot_longer(c(Turnover, Adj_turnover),
    names_to = "Type", values_to = "Turnover"
  ) %>%
  ggplot(aes(x = Year, y = Turnover)) +
  geom_line() +
  facet_grid(vars(Type), scales = "free_y") +
  xlab("Years") + ylab(NULL) +
  ggtitle("Turnover: Australian print media industry")
```

## Inflation adjustments
\fontsize{10}{10}\sf

```{r ref.label = 'retail_cpi', message=FALSE, warning=FALSE, echo=FALSE, fig.height=5}
```

## Mathematical transformations
\fontsize{13}{15}\sf

If the data show different variation at different levels of the series, then a transformation can be useful.
\pause

Denote original observations as $y_1,\dots,y_n$ and transformed
observations as $w_1, \dots, w_n$.
\pause

\begin{block}{\footnotesize Mathematical transformations for stabilizing
variation}
\begin{tabular}{llc}
Square root & $w_t = \sqrt{y_t}$ & $\downarrow$ \\[0.2cm]
Cube root & $w_t = \sqrt[3]{y_t}$ & Increasing \\[0.2cm]
Logarithm & $w_t = \log(y_t)$  & strength
\end{tabular}
\end{block}
\pause

Logarithms, in particular, are useful because they are more interpretable: changes in a log value are **relative (percent) changes on the original scale**.

## Mathematical transformations
\fontsize{12}{14}\sf

```{r food, echo=TRUE}
food <- aus_retail %>%
  filter(Industry == "Food retailing") %>%
  summarise(Turnover = sum(Turnover))
```

```{r food-plot, echo = FALSE, fig.height=3.4}
food %>% autoplot(Turnover) +
  labs(y = "Turnover ($AUD)")
```

## Mathematical transformations
\fontsize{12}{14}\sf

```{r food-sqrt1, echo=TRUE, fig.height=4}
food %>% autoplot(sqrt(Turnover)) +
  labs(y = "Square root turnover")
```

## Mathematical transformations
\fontsize{12}{14}\sf

```{r food-cbrt, echo=TRUE, fig.height=4}
food %>% autoplot(Turnover^(1/3)) +
  labs(y = "Cube root turnover")
```

## Mathematical transformations
\fontsize{12}{14}\sf

```{r food-log, echo=TRUE, fig.height=4}
food %>% autoplot(log(Turnover)) +
  labs(y = "Log turnover")
```

## Mathematical transformations
\fontsize{12}{14}\sf

```{r food-inverse, echo=TRUE, fig.height=4}
food %>% autoplot(-1/Turnover) +
  labs(y = "Inverse turnover")
```

## Box-Cox transformations

Each of these transformations is close to a member of the
family of \textbf{Box-Cox transformations}:
$$w_t = \left\{\begin{array}{ll}
        \log(y_t),      & \quad \lambda = 0; \\
        (y_t^\lambda-1)/\lambda ,         & \quad \lambda \ne 0.
\end{array}\right.
$$\pause

* $\lambda=1$: (No substantive transformation)
* $\lambda=\frac12$: (Square root plus linear transformation)
* $\lambda=0$: (Natural logarithm)
* $\lambda=-1$: (Inverse plus 1)

## Box-Cox transformations

```{r food-anim, cache=TRUE, echo=FALSE, fig.show='animate', interval=1/10, message=FALSE, fig.height=5, fig.width=8, aniopts='controls,buttonsize=0.3cm,width=11.5cm'}
food %>%
  mutate(!!!set_names(map(seq(0, 1, 0.01), ~ expr(fabletools::box_cox(Turnover, !!.x))), seq(0, 1, 0.01))) %>%
  select(-Turnover) %>%
  pivot_longer(-Month, names_to = "lambda", values_to = "Turnover") %>%
  mutate(lambda = as.numeric(lambda)) %>%
  ggplot(aes(x = Month, y = Turnover)) +
  geom_line() +
  transition_states(1 - lambda, state_length = 0) +
  view_follow() +
  ggtitle("Box-Cox transformed food retailing turnover (lambda = {format(1 - as.numeric(closest_state), digits = 2)})")
```

## Box-Cox transformations
\fontsize{12}{14}\sf


```{r food-lambda, echo=TRUE}
food %>%
  features(Turnover, features = guerrero)
```

\pause\fontsize{13}{15}\sf

* This attempts to balance the seasonal fluctuations and random variation across the series.
* Always check the results.
* A low value of $\lambda$ can give extremely large prediction intervals.

## Box-Cox transformations
\fontsize{13}{14}\sf

```{r food-bc, echo=TRUE,fig.height=4}
food %>% autoplot(box_cox(Turnover, 0.0524)) +
  labs(y = "Box-Cox transformed turnover")
```

## Transformations
\fontsize{13}{15}\sf

* Often no transformation needed.
* Simple transformations are easier to explain and work well enough.
* Transformations can have very large effect on PI.
* If the data contains zeros, then don't take logs.
* `logp1()` can be useful for data with zeros.
* If some data are negative, no power transformation is possible unless a constant is added to all values.
* Choosing logs is a simple way to force forecasts to be positive
* Transformations must be reversed to obtain forecasts on the original scale. (Handled automatically by `fable`.)

## Your turn

\fontsize{13}{14}\sf

1. For the following series, find an appropriate transformation in order to stabilise the variance.

    * United States GDP from `global_economy`
    * Slaughter of Victorian “Bulls, bullocks and steers” in `aus_livestock`
    * Victorian Electricity Demand from `vic_elec`.
    * Gas production from `aus_production`

2. Why is a Box-Cox transformation unhelpful for the `canadian_gas` data?

# Time series components

## Time series patterns

**Recall**

Trend
:  pattern exists when there is a long-term increase or decrease in the data.

Cyclic
: pattern exists when data exhibit rises and falls that are *not of fixed period* (duration usually of at least 2 years).

Seasonal
: pattern exists when a series is influenced by seasonal factors (e.g., the quarter of the year, the month, or day of the week).

## Time series decomposition

\begin{block}{}\vspace*{-0.3cm}
\[ y_t = f(S_t, T_t, R_t) \]
\end{block}
\begin{tabular}{@{}llp{8cm}@{}}
where & $y_t=$ & data at period $t$ \\
      & $T_t=$ & trend-cycle component at period $t$\\
      & $S_t=$ & seasonal component at period $t$ \\
      & $R_t=$ & remainder component at period $t$
\end{tabular}
\pause

**Additive decomposition:** $y_t = S_t + T_t + R_t.$

**Multiplicative decomposition:** $y_t = S_t \times T_t \times R_t.$

## Time series decomposition
\fontsize{13}{15}\sf

  *  Additive model  appropriate if  magnitude of  seasonal fluctuations does not vary with level.
  *  If seasonal are proportional to level of series, then multiplicative model appropriate.
  *  Multiplicative decomposition more prevalent with economic series
  *  Alternative: use a Box-Cox transformation, and then use additive decomposition.
  *  Logs turn multiplicative relationship into an additive relationship:

$$y_t = S_t \times T_t \times E_t \quad\Rightarrow\quad
\log y_t = \log S_t + \log T_t + \log R_t.
$$

## US Retail Employment
\fontsize{10}{11}\sf

```{r usretail}
us_retail_employment <- us_employment %>%
  filter(year(Month) >= 1990, Title == "Retail Trade") %>%
  select(-Series_ID)
us_retail_employment
```

\vspace*{10cm}

## US Retail Employment
\fontsize{10}{11}\sf

```{r dable1}
us_retail_employment %>%
  autoplot(Employed) +
  xlab("Year") + ylab("Persons (thousands)") +
  ggtitle("Total employment in US retail")
```

\vspace*{10cm}

## US Retail Employment
\fontsize{10}{11}\sf

```{r dable2}
us_retail_employment %>%
  model(stl = STL(Employed))
```

\vspace*{10cm}

## US Retail Employment
\fontsize{10}{11}\sf

```{r dable3}
dcmp <- us_retail_employment %>%
  model(stl = STL(Employed))
components(dcmp)
```

\vspace*{10cm}

## US Retail Employment
\fontsize{10}{11}\sf

```{r dable4}
us_retail_employment %>%
  autoplot(Employed, color='gray') +
  autolayer(components(dcmp), trend, color='red') +
  xlab("Year") + ylab("Persons (thousands)") +
  ggtitle("Total employment in US retail")
```

\vspace*{10cm}

## US Retail Employment
\fontsize{10}{11}\sf

```{r usretail-stl, fig.width=8, fig.height=5}
components(dcmp) %>% autoplot() + xlab("Year")
```

## US Retail Employment
\fontsize{10}{11}\sf

```{r usretail3}
components(dcmp) %>% gg_subseries(season_year)
```

## Seasonal adjustment

  *  Useful by-product of decomposition:  an easy way to calculate seasonally adjusted data.
  *  Additive decomposition: seasonally adjusted data given by
$$y_t - S_t = T_t + R_t$$
  *  Multiplicative decomposition: seasonally adjusted data given by
$$y_t / S_t = T_t \times R_t$$

## US Retail Employment
\fontsize{10}{11}\sf

```{r usretail-sa}
us_retail_employment %>%
  autoplot(Employed, color='gray') +
  autolayer(components(dcmp), season_adjust, color='blue') +
  xlab("Year") + ylab("Persons (thousands)") +
  ggtitle("Total employment in US retail")
```

## Seasonal adjustment

  * We use estimates of $S$ based on past values to seasonally adjust a current value.
  *  Seasonally adjusted series reflect **remainders** as well as **trend**. Therefore they are not "smooth" and "downturns" or "upturns" can be misleading.
  *  It is better to use the trend-cycle component to look for turning points.

# History of time series decomposition

## History of time series decomposition

\fontsize{13}{15}\sf

  *  Classical method originated in 1920s.
  *  Census II method introduced in 1957. Basis for X-11 method and variants (including X-12-ARIMA, X-13-ARIMA)
  *  STL method introduced in 1983
  *  TRAMO/SEATS introduced in 1990s.
\pause

### National Statistics Offices
 * ABS uses X-12-ARIMA
 * US Census Bureau uses X-13ARIMA-SEATS
 * Statistics Canada uses X-12-ARIMA
 * ONS (UK) uses X-12-ARIMA
 * EuroStat use X-13ARIMA-SEATS

## X-11 decomposition

**Advantages**

  *  Relatively robust to outliers
  *  Completely automated choices for trend and seasonal changes
  *  Very widely tested on economic data over a long period of time.

\pause

**Disadvantages**

  *  No prediction/confidence intervals
  *  Ad hoc method with no underlying model
  *  Only developed for quarterly and monthly data

## Extensions: X-12-ARIMA and X-13-ARIMA

  *  The X-11, X-12-ARIMA and X-13-ARIMA methods are based on Census II decomposition.
  *  These allow adjustments for trading days and other explanatory variables.
  *  Known outliers can be omitted.
  *  Level shifts and ramp effects can be modelled.
  *  Missing values estimated and replaced.
  *  Holiday factors (e.g., Easter, Labour Day) can be estimated.

## X-13ARIMA-SEATS

**Advantages**

  * Model-based
  * Smooth trend estimate
  * Allows estimates at end points
  * Allows changing seasonality
  * Developed for economic data

\pause

**Disadvantages**

  *  Only developed for quarterly and monthly data

# STL decomposition

## STL decomposition

\fontsize{13}{14}\sf

  *  STL: "Seasonal and Trend decomposition using Loess"
  *  Very versatile and robust.
  *  Unlike X-12-ARIMA, STL will handle any type of seasonality.
  *  Seasonal component allowed to change over time, and rate of change controlled by user.
  *  Smoothness of trend-cycle also controlled by user.
  *  Robust to outliers
  *  Not trading day or calendar adjustments.
  *  Only additive.
  *  Take logs to get multiplicative decomposition.
  *  Use Box-Cox transformations to get other decompositions.

## STL decomposition
\fontsize{10}{11}\sf

```{r stlwindow9, echo=TRUE, warning=FALSE, fig.width=8, fig.height=4}
us_retail_employment %>%
  model(STL(Employed ~ season(window=9), robust=TRUE)) %>%
  components() %>% autoplot() +
    ggtitle("STL decomposition: US retail employment")
```

## STL decomposition
\fontsize{10}{11}\sf

```{r stlwindowanim, echo=FALSE, warning=FALSE, message=FALSE, fig.show='animate', interval=1/10,  fig.height=5.35, fig.width=8, aniopts='controls,buttonsize=0.3cm,width=11.5cm', eval=TRUE}
s_windows <- seq(5,55,by=2)
stl_defs <- purrr::map(s_windows, function(s_window){
  STL(Employed ~ season(window=s_window), robust=TRUE)
})
names(stl_defs) <- sprintf("season(window=%02d)", s_windows)

us_retail_employment %>%
  model(!!!stl_defs) %>%
  components() %>%
  as_tibble() %>%
  pivot_longer(Employed:remainder,
               names_to = "component", names_ptypes = list(component = factor(levels = c("Employed", "trend", "season_year", "remainder"))),
               values_to = "Employed") %>%
  ggplot(aes(x = Month, y = Employed)) +
  geom_line() +
  facet_grid(rows = vars(component), scales = "free_y") +
  labs(title = "STL decomposition of US retail employment",
       subtitle = "{closest_state}") +
  transition_states(.model)
```

\vspace*{10cm}

## STL decomposition
\fontsize{10}{11}\sf

```{r echo = TRUE, results = 'hide'}
us_retail_employment %>%
  model(STL(Employed ~ season(window=5))) %>%
  components()

us_retail_employment %>%
  model(STL(Employed ~ trend(window=15) +
                       season(window="periodic"),
            robust = TRUE)
  ) %>% components()
```

\fontsize{12}{13}\sf

  *  `trend(window = ?)` controls wiggliness of trend component.
  *  `season(window = ?)` controls variation on seasonal component.
  *  `season(window = 'periodic')` is equivalent to an infinite window.

## STL decomposition
\fontsize{10}{11}\sf

```{r mstl, fig.width=8, fig.height=4}
us_retail_employment %>%
  model(STL(Employed)) %>%
  components() %>%
  autoplot()
```

* `STL()` chooses `season(window=13)` by default
* Can include transformations.

## STL decomposition
\fontsize{13}{14.5}\sf

* Algorithm that updates trend and seasonal components iteratively.
* Starts with $\hat{T}_t=0$
* Uses a mixture of loess and moving averages to successively refine the trend and seasonal estimates.
* The trend window controls loess bandwidth applied to deasonalised values.
* The season window controls loess bandwidth applied to detrended subseries.
* Robustness weights based on remainder.
* Default season `window = 13`
* Default trend `window = nextodd(` \newline\mbox{}\hfill `ceiling((1.5*period)/(1-(1.5/s.window)))`

# When things go wrong

## The ABS stuff-up

\fullwidth{abs1}

## The ABS stuff-up

\fullwidth{abs2}

## The ABS stuff-up

\fullwidth{abs3}

## The ABS stuff-up

\fontsize{10}{12}\sf

```{r abs1, echo=FALSE}
employed <- tsibble(
  Time = yearmonth("1978 Feb") + 0:439,
  Employed = c(
    5985.7, 6040.6, 6054.2, 6038.3, 6031.3, 6036.1, 6005.4, 6024.3, 6045.9, 6033.8, 6125.4, 5971.3,
    6050.7, 6096.2, 6087.7, 6075.6, 6095.7, 6103.9, 6078.5, 6157.8, 6164.0, 6188.8, 6257.2, 6112.9,
    6207.2, 6278.7, 6224.9, 6273.4, 6269.9, 6314.1, 6281.4, 6360.0, 6320.2, 6342.0, 6426.6, 6253.0,
    6356.5, 6428.1, 6426.3, 6412.4, 6413.9, 6425.3, 6393.7, 6502.7, 6445.3, 6433.3, 6506.9, 6355.5,
    6432.4, 6497.4, 6431.6, 6440.9, 6414.3, 6425.9, 6379.3, 6443.5, 6421.1, 6366.8, 6370.1, 6172.0,
    6263.9, 6310.3, 6254.5, 6272.8, 6266.5, 6295.0, 6241.2, 6358.2, 6336.1, 6377.5, 6456.5, 6251.4,
    6365.4, 6503.2, 6477.6, 6489.7, 6499.0, 6528.7, 6466.1, 6579.8, 6553.2, 6576.1, 6636.0, 6452.4,
    6595.7, 6657.4, 6588.8, 6657.9, 6659.4, 6703.4, 6675.5, 6814.7, 6771.1, 6881.9, 6910.8, 6753.6,
    6861.9, 6961.9, 6997.9, 6979.0, 7007.7, 6991.5, 6918.5, 7040.6, 7030.4, 7034.2, 7116.8, 6902.5,
    7022.3, 7133.4, 7109.6, 7103.5, 7128.9, 7175.6, 7092.3, 7186.5, 7177.4, 7182.2, 7330.7, 7169.4,
    7247.3, 7397.4, 7383.4, 7354.8, 7378.3, 7383.1, 7353.3, 7503.2, 7477.3, 7508.6, 7622.9, 7423.8,
    7566.5, 7634.6, 7678.4, 7720.8, 7711.0, 7740.8, 7715.3, 7841.6, 7806.5, 7862.4, 7935.5, 7707.7,
    7803.0, 7874.1, 7887.9, 7908.5, 7900.3, 7919.4, 7808.0, 7905.5, 7848.9, 7826.9, 7915.5, 7641.3,
    7708.7, 7715.4, 7717.2, 7703.7, 7678.1, 7583.0, 7620.7, 7713.2, 7638.0, 7614.9, 7712.2, 7518.9,
    7597.2, 7646.2, 7644.1, 7631.4, 7637.3, 7668.3, 7613.4, 7709.7, 7665.7, 7587.4, 7693.4, 7533.7,
    7531.0, 7645.7, 7572.6, 7620.5, 7627.9, 7646.5, 7589.4, 7747.6, 7738.8, 7744.9, 7842.1, 7646.8,
    7738.6, 7824.2, 7827.4, 7857.9, 7878.4, 7966.0, 7861.7, 8054.4, 7997.2, 8003.3, 8135.5, 7928.4,
    8049.9, 8118.1, 8174.6, 8165.2, 8205.6, 8229.0, 8165.9, 8300.4, 8232.6, 8300.3, 8395.7, 8166.7,
    8246.6, 8280.4, 8248.0, 8297.1, 8311.7, 8332.1, 8265.9, 8373.0, 8319.4, 8314.4, 8431.4, 8235.2,
    8291.4, 8347.5, 8343.1, 8330.2, 8345.6, 8374.9, 8250.3, 8474.0, 8405.2, 8462.1, 8540.5, 8334.7,
    8413.0, 8460.0, 8499.9, 8482.5, 8516.8, 8541.9, 8455.2, 8653.2, 8601.0, 8554.3, 8696.5, 8477.4,
    8556.5, 8618.9, 8631.9, 8606.5, 8673.2, 8706.7, 8603.6, 8777.1, 8755.3, 8763.7, 8900.7, 8628.2,
    8754.4, 8830.7, 8882.2, 8865.0, 8922.0, 9020.0, 8911.6, 9061.3, 8973.1, 8912.7, 9059.6, 8834.9,
    8920.9, 8956.0, 9023.6, 9004.6, 9021.9, 9048.9, 8971.9, 9105.9, 9058.7, 9055.6, 9177.1, 8993.4,
    9092.3, 9128.5, 9129.5, 9134.7, 9180.8, 9194.5, 9150.3, 9303.5, 9249.1, 9286.7, 9439.7, 9281.7,
    9372.6, 9362.1, 9365.6, 9380.1, 9370.4, 9363.9, 9327.0, 9486.1, 9447.8, 9427.7, 9573.6, 9363.8,
    9434.5, 9506.4, 9512.0, 9533.5, 9543.3, 9553.1, 9462.1, 9668.6, 9662.2, 9652.9, 9807.8, 9634.4,
    9744.6, 9828.3, 9856.3, 9850.8, 9896.6, 9912.3, 9870.3, 10004.6, 9949.7, 9945.0, 10074.7, 9842.7,
    9961.1, 10048.7, 10041.0, 10082.1, 10120.8, 10170.8, 10105.8, 10299.5, 10212.4, 10201.6, 10404.3,
    10156.1, 10277.0, 10349.2, 10362.9, 10412.0, 10436.3, 10456.8, 10406.4, 10588.8, 10520.5, 10535.0,
    10710.1, 10524.9, 10622.9, 10677.4, 10706.2, 10690.3, 10745.0, 10761.9, 10710.4, 10854.5, 10807.4,
    10757.3, 10915.6, 10681.0, 10776.7, 10775.2, 10792.7, 10786.8, 10770.9, 10808.8, 10707.3, 10882.1,
    10845.2, 10829.2, 11010.9, 10809.9, 10889.2, 10928.9, 10940.1, 10957.4, 11009.3, 11030.5, 10973.8,
    11159.4, 11129.0, 11144.5, 11295.0, 11063.7, 11146.2, 11217.0, 11186.5, 11196.2, 11221.3, 11227.5,
    11130.7, 11321.2, 11274.0, 11240.6, 11354.8, 11159.0, 11236.2, 11332.4, 11328.3, 11389.0, 11350.6,
    11363.7, 11259.8, 11452.6, 11401.9, 11375.0, 11518.4, 11304.0, 11424.3, 11436.3, 11482.2, 11495.6,
    11497.8, 11486, 11369, 11547, 11499, 11472, 11571, 11354, 11493, 11562, 11589, 11595, 11602, 11590,
    11622, 11593
  ),
  index = Time
) %>%
  mutate(
    Month = month(Time, label = TRUE),
    Year = year(Time)
  ) %>%
  select(Time, Month, Year, Employed)
```

```{r abs2}
employed
```

## The ABS stuff-up

\fontsize{9}{10}\sf

```{r abs3, fig.height=4}
employed %>%
  autoplot(Employed) +
  ggtitle("Total employed") + ylab("Thousands") + xlab("Year")
```

## The ABS stuff-up

\fontsize{9}{10}\sf

```{r abs4, fig.height=4}
employed %>%
  filter(Year >= 2005) %>%
  autoplot(Employed) +
  ggtitle("Total employed") + ylab("Thousands") + xlab("Year")
```

## The ABS stuff-up

\fontsize{9}{10}\sf

```{r abs5, fig.height=4}
employed %>%
  filter(Year >= 2005) %>%
  gg_season(Employed, label = "right") +
  ggtitle("Total employed") + ylab("Thousands")
```

## The ABS stuff-up

\fontsize{9}{10}\sf

```{r abs6, fig.height=2}
employed %>%
  mutate(diff = difference(Employed)) %>%
  filter(Month == "Sep") %>%
  ggplot(aes(y = diff, x = 1)) +
  geom_boxplot() + coord_flip() +
  ggtitle("Sep - Aug: total employed") +
  xlab("") + ylab("Thousands") +
  scale_x_continuous(breaks = NULL, labels = NULL)
```

## The ABS stuff-up

\fontsize{9}{10}\sf

```{r abs7, fig.height=3.85}
dcmp <- employed %>%
  filter(Year >= 2005) %>%
  model(stl = STL(Employed ~ season(window = 11), robust = TRUE))
components(dcmp) %>% autoplot()
```

## The ABS stuff-up

\fontsize{9}{10}\sf

```{r abs8, fig.height=3.5}
components(dcmp) %>%
  filter(year(Time) == 2013) %>%
  gg_season(season_year) +
  ggtitle("Seasonal component") +
  guides(colour = "none")
```

## The ABS stuff-up

\fontsize{9}{10}\sf

```{r abs9}
components(dcmp) %>%
  as_tsibble() %>%
  autoplot(season_adjust)
```

## The ABS stuff-up
\fontsize{13}{15}\sf

  *  August 2014 employment numbers higher than expected.
  *  Supplementary survey usually conducted in August for employed people.
  *  Most likely, some employed people were claiming to be unemployed in August to avoid supplementary questions.
  *  Supplementary survey not run in 2014, so no motivation to lie about employment.
  *  In previous years, seasonal adjustment fixed the problem.
  *  The ABS has now adopted a new method to avoid the bias.

