---
title: "5. The forecaster's toolbox"
author: "5.1 A tidy forecasting workflow"
date: "OTexts.org/fpp3/"
classoption: aspectratio=169
titlepage: fpp3title.png
titlecolor: fpp3red
toc: false
output:
  binb::monash:
    colortheme: monashwhite
    fig_width: 7.5
    fig_height: 3
    keep_tex: no
    includes:
      in_header: fpp3header.tex
---

```{r setup, include=FALSE}
source("setup.R")
```

## A tidy forecasting workflow

The process of producing forecasts can be split up into a few fundamental steps.

1. Preparing data
2. Data visualisation
3. Specifying a model
4. Model estimation
5. Accuracy \& performance evaluation
6. Producing forecasts

## A tidy forecasting workflow

```{r workflow, echo = FALSE}
line_curve <- function(x, y, xend, yend, ...) {
  geom_curve(
    aes(x = x, y = y, xend = xend, yend = yend),
    arrow = arrow(type = "closed", length = unit(0.03, "npc")),
    ...
  )
}

ggplot() +
  geom_text(
    aes(x = x, y = y, label = label),
    data = tribble(
      ~x, ~y, ~label,
      1, 0, "Tidy",
      7 / 3, 0, "Visualise",
      3, 0.5, "Specify",
      11 / 3, 0, "Estimate",
      3, -0.5, "Evaluate",
      5, 0, "Forecast"
    ),
    size = 5
  ) +
  geom_segment(
    aes(x = x, y = y, xend = xend, yend = yend),
    data = tribble(
      ~x, ~y, ~xend, ~yend,
      1.3, 0, 1.9, 0,
      4.1, 0, 4.6, 0
    ),
    arrow = arrow(type = "closed", length = unit(0.03, "npc"))
  ) +
  line_curve(7 / 3, 0.1, 8 / 3, 0.5, angle = 250, curvature = -0.3) +
  line_curve(10 / 3, 0.5, 11 / 3, 0.1, angle = 250, curvature = -0.3) +
  line_curve(8 / 3, -0.5, 7 / 3, -0.1, angle = 250, curvature = -0.3) +
  line_curve(11 / 3, -0.1, 10 / 3, -0.5, angle = 250, curvature = -0.3) +
  theme_void() +
  xlim(0.8, 5.2) +
  ylim(-0.6, 0.6) +
  coord_equal(ratio = 1)
```

## Data preparation (tidy)
\fontsize{10}{13}\sf

```{r GDPpc, fig.height = 3.2}
gdppc <- global_economy |>
  mutate(GDP_per_capita = GDP / Population) |>
  select(Year, Country, GDP, Population, GDP_per_capita)
gdppc
```

## Data visualisation

```{r GDP-plot, fig.height = 2.8}
gdppc |>
  filter(Country == "Sweden") |>
  autoplot(GDP_per_capita) +
  labs(title = "GDP per capita for Sweden", y = "$US")
```

## Model estimation

The `model()` function trains models to data.

\fontsize{10}{13}\sf

```{r GDP-model, warning=FALSE}
fit <- gdppc |>
  model(trend_model = TSLM(GDP_per_capita ~ trend()))
fit
```

\only<2>{\begin{textblock}{5}(9,4)
\begin{alertblock}{}
A \texttt{mable} is a model table, each cell corresponds to a fitted model.
\end{alertblock}
\end{textblock}}

## Producing forecasts

\fontsize{10}{13}\sf

```{r GDP-fc, echo = TRUE, dependson='GDP-model', warning=FALSE}
fit |> forecast(h = "3 years")
```

\only<2>{\begin{textblock}{5}(9,2)
\begin{alertblock}{}
A \texttt{fable} is a forecast table with point forecasts and distributions.
\end{alertblock}
\end{textblock}}

## Visualising forecasts

\footnotesize

```{r GDP-fc-plot, warning=FALSE, message=FALSE, fig.height=2.8}
fit |>
  forecast(h = "3 years") |>
  filter(Country == "Sweden") |>
  autoplot(gdppc) + labs(title = "GDP per capita for Sweden", y = "$US")
```
