---
title: "5. The forecaster's toolbox"
author: "5.10 Time series cross-validation"
date: "OTexts.org/fpp3/"
classoption: aspectratio=169
titlepage: fpp3title.png
titlecolor: fpp3red
toc: false
output:
  binb::monash:
    colortheme: monashwhite
    fig_width: 7.5
    fig_height: 3
    keep_tex: no
    includes:
      in_header: header.tex
---

```{r setup, include=FALSE}
source("setup.R")
```

```{r tscvplots, echo=FALSE}
tscv_plot <- function(.init, .step, h = 1) {
  expand.grid(
    time = seq(26),
    .id = seq(trunc(20 / .step))
  ) |>
    group_by(.id) |>
    mutate(
      observation = case_when(
        time <= ((.id - 1) * .step + .init) ~ "train",
        time %in% c((.id - 1) * .step + .init + h) ~ "test",
        TRUE ~ "unused"
      )
    ) |>
    ungroup() |>
    filter(.id <= 26 - .init) |>
    ggplot(aes(x = time, y = .id)) +
    geom_segment(
      aes(x = 0, xend = 27, y = .id, yend = .id),
      arrow = arrow(length = unit(0.015, "npc")),
      col = "black", size = .25
    ) +
    geom_point(aes(col = observation), size = 2) +
    scale_y_reverse() +
    scale_color_manual(values = c(train = "#0072B2", test = "#D55E00", unused = "gray")) +
    # theme_void() +
    # geom_label(aes(x = 28.5, y = 1, label = "time")) +
    guides(col = FALSE) +
    labs(x = "time", y = "") +
    theme_void() +
    theme(axis.title = element_text())
}
```

## Time series cross-validation {-}

**Traditional evaluation**

```{r traintest1, fig.height=1, echo=FALSE, dependson="tscvplots"}
tscv_plot(.init = 18, .step = 20, h = 1:8) +
  geom_text(aes(x = 10, y = 0.8, label = "Training data"), color = "#0072B2") +
  geom_text(aes(x = 21, y = 0.8, label = "Test data"), color = "#D55E00") +
  ylim(1, 0)
```

\pause

**Time series cross-validation**

```{r tscvggplot1, echo=FALSE}
tscv_plot(.init = 3, .step = 1, h = 1) +
  geom_text(aes(x = 21, y = 0, label = "h = 1"), color = "#D55E00")
```

## Time series cross-validation {-}

**Traditional evaluation**

```{r traintest2, ref.label="traintest1", fig.height=1, echo=FALSE}
```

**Time series cross-validation**

```{r tscvggplot2, echo=FALSE,  dependson="tscvplots"}
tscv_plot(.init = 3, .step = 1, h = 2) +
  geom_text(aes(x = 21, y = 0, label = "h = 2"), color = "#D55E00")
```

## Time series cross-validation {-}

**Traditional evaluation**

```{r traintest3, ref.label="traintest1", fig.height=1, echo=FALSE}
```

**Time series cross-validation**

```{r tscvggplot3, echo=FALSE,  dependson="tscvplots"}
tscv_plot(.init = 3, .step = 1, h = 3) +
  geom_text(aes(x = 21, y = 0, label = "h = 3"), color = "#D55E00")
```

## Time series cross-validation {-}

**Traditional evaluation**

```{r traintest4, ref.label="traintest1", fig.height=1, echo=FALSE}
```

**Time series cross-validation**

```{r tscvggplot4, echo=FALSE,  dependson="tscvplots"}
tscv_plot(.init = 3, .step = 1, h = 4) +
  geom_text(aes(x = 21, y = 0, label = "h = 4"), color = "#D55E00")
```

\only<2>{\begin{textblock}{8}(.5,6.5)\begin{block}{}\fontsize{12}{13}\sf
\begin{itemize}\tightlist
\item Forecast accuracy averaged over test sets.
\item Also known as "evaluation on a rolling forecasting origin"
\end{itemize}\end{block}\end{textblock}}

\vspace*{10cm}

<!--
## Creating the rolling training sets {-}

\fontsize{13}{14}\sf

There are three main rolling types which can be used.

* Stretch: extends a growing length window with new data.
* Slide: shifts a fixed length window through the data.
* Tile: moves a fixed length window without overlap.

Three functions to roll a tsibble: `stretch_tsibble()`, `slide_tsibble()`,
and `tile_tsibble()`.

For time series cross-validation, stretching windows are most commonly used. -->

<!-- ## Creating the rolling training sets {-}

```{r animate, echo = FALSE, warning = FALSE, message = FALSE, fig.show='animate', interval=1/10, fig.height=4, fig.width=8, aniopts='controls,buttonsize=0.3cm,width=11.5cm'}
library(gganimate)
tourism_melb <- tourism |>
  filter(Region == "Melbourne", Purpose == "Holiday") |>
  select(Quarter, Trips)
slide_window <- slide_tsibble(tourism_melb, .size = 4) |>
  mutate(type = "Slide") |>
  as_tibble()
tile_window <- tile_tsibble(tourism_melb, .size = 4) |>
  mutate(type = "Tile") |>
  as_tibble()
tile_window <- bind_rows(
  tile_window |> mutate(.id = 4 * (.id - 1) + 1),
  tile_window |> mutate(.id = 4 * (.id - 1) + 2),
  tile_window |> mutate(.id = 4 * (.id - 1) + 3),
  tile_window |> mutate(.id = 4 * (.id - 1) + 4),
)
stretch_window <- stretch_tsibble(tourism_melb, .init = 4) |>
  mutate(type = "Stretch") |>
  as_tibble()
window <- bind_rows(slide_window, tile_window, stretch_window) |>
  group_by(.id, type) |>
  mutate(xmin = min(Quarter), xmax = max(Quarter), ymin = -Inf, ymax = Inf) |>
  ungroup() |>
  select(-Trips) |>
  mutate(type = factor(type, levels = c("Stretch", "Slide", "Tile")))

ggplot() +
  geom_line(aes(x = Quarter, y = Trips), data = tourism_melb, colour = "grey", size = 1.2) +
  geom_rect(
    aes(
      xmin = xmin, xmax = xmax,
      ymin = ymin, ymax = ymax,
      group = .id
    ),
    data = window,
    fill = "#9ecae1", colour = "#9ecae1", size = 1.5, alpha = 0.6
  ) +
  labs(x = "Quarter", y = "Trips") +
  facet_wrap(~type, ncol = 1) +
  theme_bw() +
  transition_manual(.id)
```
 -->

## Time series cross-validation {-}

\fontsize{12}{13}\sf

Stretch with a minimum length of 3, growing by 1 each step.

```{r fbf2, include=FALSE}
# Extract training data
fb_stock <- gafa_stock |>
  filter(Symbol == "FB") |>
  mutate(trading_day = row_number()) |>
  update_tsibble(index = trading_day, regular = TRUE)
```

```{r google-stretch, cache=TRUE}
fb_stretch <- fb_stock |>
  stretch_tsibble(.init = 3, .step = 1) |>
  filter(.id != max(.id))
```
\fontsize{10}{11}\sf
```{r google-stretch-print, echo = FALSE}
options(width = 60)
fb_stretch |>
  select(Date, Close, trading_day, .id) |>
  print(n = 7)
```

## Time series cross-validation {-}

\small

Estimate RW w/ drift models for each window.

```{r google-fit, cache = TRUE}
fit_cv <- fb_stretch |>
  model(RW(Close ~ drift()))
```

\fontsize{10}{11}\sf
```{r google-fit-print, echo = FALSE}
print(fit_cv, n = 4)
```

## Time series cross-validation {-}

\fontsize{9.5}{11}\sf

Produce one step ahead forecasts from all models.

```{r google-fc, cache = TRUE}
fc_cv <- fit_cv |>
  forecast(h = 1)
```

```{r google-fc-print, echo = FALSE}
fc_cv |>
  select(-.model) |>
  print(n = 4)
```

## Time series cross-validation {-}

\fontsize{11}{11}\sf

```{r google-accuracy, cache = TRUE, results = 'hide', eval = FALSE}
# Cross-validated
fc_cv |> accuracy(fb_stock)
# Training set
fb_stock |>
  model(RW(Close ~ drift())) |>
  accuracy()
```

\fontsize{13}{15}\sf

```{r, echo = FALSE, warning = FALSE}
fc_cv |>
  accuracy(fb_stock) |>
  mutate(.type = "Cross-validation") |>
  bind_rows(
    fb_stock |>
      model(RW(Close ~ drift())) |>
      accuracy()
  ) |>
  transmute(Type = .type, RMSE, MAE, MAPE) |>
  gt::gt("Type") |>
  gt::fmt_number(columns = 2:4, decimals = 3) |>
  gt::as_latex()
```

A good way to choose the best forecasting model is to find the model with the smallest RMSE computed using time series cross-validation.
