---
title: "5. The forecaster's toolbox"
author: "5.2 Some simple forecasting methods"
date: "OTexts.org/fpp3/"
classoption: aspectratio=169
titlepage: fpp3title.png
titlecolor: fpp3red
toc: false
output:
  binb::monash:
    colortheme: monashwhite
    fig_width: 7.5
    fig_height: 3
    keep_tex: no
    includes:
      in_header: header.tex
---

```{r setup, include=FALSE}
source("setup.R")
```

## Some simple forecasting methods
\fontsize{13}{14}\sf

### `MEAN(y)`: Average method

  * Forecast of all future values is equal to mean of historical data $\{y_1,\dots,y_T\}$.
  * Forecasts: $\hat{y}_{T+h|T} = \bar{y} = (y_1+\dots+y_T)/T$

```{r mean-method-explained, echo=FALSE, message=FALSE, warning=FALSE, fig.height = 2.7}
bricks <- aus_production |>
  filter(!is.na(Bricks)) |>
  mutate(average = mean(Bricks))

fc <- bricks |>
  filter(row_number() == n()) |>
  as_tibble() |>
  unnest(Quarter = list(as.Date(Quarter) + months(c(0, 12 * 5))))

bricks |>
  ggplot(aes(x = Quarter, y = Bricks)) +
  geom_line() +
  geom_line(aes(y = average), colour = "#0072B2", linetype = "dashed") +
  geom_line(aes(y = average), data = fc, colour = "#0072B2") +
  labs(title = "Clay brick production in Australia")
```

\vspace*{10cm}

## Some simple forecasting methods
\fontsize{13}{14}\sf

### `NAIVE(y)`: Naïve method

  * Forecasts equal to last observed value.
  * Forecasts: $\hat{y}_{T+h|T} =y_T$.
  * Consequence of efficient market hypothesis.

```{r naive-method-explained, echo = FALSE, warning = FALSE, fig.height = 2.75}
bricks |>
  filter(!is.na(Bricks)) |>
  model(NAIVE(Bricks)) |>
  forecast(h = "5 years") |>
  autoplot(filter(bricks, year(Quarter) > 1990), level = NULL) +
  geom_point(data = slice(bricks, n()), aes(y = Bricks), colour = "#0072B2") +
  labs(title = "Clay brick production in Australia")
```

\vspace*{10cm}

## Some simple forecasting methods
\fontsize{13}{14}\sf

### `SNAIVE(y ~ lag(m))`: Seasonal naïve method

  * Forecasts equal to last value from same season.
  * Forecasts: $\hat{y}_{T+h|T} =y_{T+h-m(k+1)}$, where $m=$ seasonal period and $k$ is the integer part of $(h-1)/m$.

```{r snaive-method-explained, echo = FALSE, warning = FALSE, fig.height = 2.75}
bricks |>
  model(SNAIVE(Bricks ~ lag("year"))) |>
  forecast(h = "5 years") |>
  autoplot(filter(bricks, year(Quarter) > 1990), level = NULL) +
  geom_point(data = slice(bricks, (n() - 3):n()), aes(y = Bricks), colour = "#0072B2") +
  labs(title = "Clay brick production in Australia")
```

\vspace*{10cm}

## Some simple forecasting methods
\fontsize{13}{14}\sf

### `RW(y ~ drift())`: Drift method

 * Forecasts equal to last value plus average change.
 * Forecasts:\vspace*{-.7cm}

 \begin{align*}
 \hat{y}_{T+h|T} & =  y_{T} + \frac{h}{T-1}\sum_{t=2}^T (y_t-y_{t-1})\\
                 & = y_T + \frac{h}{T-1}(y_T -y_1).
 \end{align*}\vspace*{-0.2cm}

   * Equivalent to extrapolating a line drawn between first and last observations.

## Some simple forecasting methods

### Drift method

```{r drift-method-explained, echo = FALSE, warning = FALSE}
aus_production |>
  filter(!is.na(Bricks)) |>
  model(RW(Bricks ~ drift())) |>
  forecast(h = "5 years") |>
  autoplot(aus_production, level = NULL) +
  geom_line(
    data = slice(aus_production, range(cumsum(!is.na(Bricks)))),
    aes(y = Bricks), linetype = "dashed", colour = "#0072B2"
  ) +
  labs(title = "Clay brick production in Australia")
```

\vspace*{10cm}

## Model fitting

The `model()` function trains models to data.

\fontsize{10}{11}\sf

```{r brick-model}
brick_fit <- aus_production |>
  filter(!is.na(Bricks)) |>
  model(
    Seasonal_naive = SNAIVE(Bricks),
    Naive = NAIVE(Bricks),
    Drift = RW(Bricks ~ drift()),
    Mean = MEAN(Bricks)
  )
```

```{r brick-model2, echo=FALSE, dependson='brick-model'}
brick_fit
```

## Producing forecasts

\fontsize{10}{13}\sf

```{r brick-fc, echo = TRUE, dependson='brick-model'}
brick_fc <- brick_fit |>
  forecast(h = "5 years")
```

```{r brick-fbl, echo = FALSE, dependson='brick-fc'}
print(brick_fc, n = 4)
```

## Visualising forecasts

\footnotesize

```{r brick-fc-plot, warning=FALSE, message=FALSE, fig.height=2.6, dependson='brick-fc'}
brick_fc |>
  autoplot(aus_production, level = NULL) +
  labs(title = "Clay brick production in Australia",
       y = "Millions of bricks") +
  guides(colour = guide_legend(title = "Forecast"))
```

## Facebook closing stock price

\fontsize{9}{10}\sf

```{r fbf2, fig.show='hide'}
# Extract training data
fb_stock <- gafa_stock |>
  filter(Symbol == "FB") |>
  mutate(trading_day = row_number()) |>
  update_tsibble(index = trading_day, regular = TRUE)

# Specify, estimate and forecast
fb_stock |>
  model(
    Mean = MEAN(Close),
    Naive = NAIVE(Close),
    Drift = RW(Close ~ drift())
  ) |>
  forecast(h = 42) |>
  autoplot(fb_stock, level = NULL) +
  labs(title = "Facebook closing stock price", y = "$US") +
  guides(colour = guide_legend(title = "Forecast"))
```

## Facebook closing stock price

```{r ref.label='fbf2', echo=FALSE, fig.height=4}
```
