---
title: "5. The forecaster's toolbox"
author: "5.4 Residual diagnostics"
date: "OTexts.org/fpp3/"
classoption: aspectratio=169
titlepage: fpp3title.png
titlecolor: fpp3red
toc: false
output:
  binb::monash:
    colortheme: monashwhite
    fig_width: 7.5
    fig_height: 3
    keep_tex: no
    includes:
      in_header: header.tex
---

```{r setup, include=FALSE}
source("setup.R")
```

## Facebook closing stock price


```{r fbf2, include=FALSE}
# Extract training data
fb_stock <- gafa_stock |>
  filter(Symbol == "FB") |>
  mutate(trading_day = row_number()) |>
  update_tsibble(index=trading_day, regular=TRUE)
```

```{r fbf}
fb_stock |> autoplot(Close)
```

## Facebook closing stock price
\fontsize{10}{10}\sf

```{r augment}
fit <- fb_stock |> model(NAIVE(Close))
augment(fit)
```

\only<2>{\begin{textblock}{6}(.5,6.2)\fontsize{14}{16}\sf
\begin{alertblock}{Na\"{\i}ve forecasts:}\vspace*{-0.4cm}
\begin{align*}
\hat{y}_{t|t-1} & = y_{t-1}\\
e_t & = y_t - \hat{y}_{t|t-1} = y_t-y_{t-1}
\end{align*}
\end{alertblock}\end{textblock}}

\only<2>{\begin{textblock}{1}(9,2.7)\fontsize{14}{16}\sf
\begin{alertblock}{}\centerline{$\hat{y}_{t|t-1}$}\end{alertblock}\end{textblock}}
\only<2>{\begin{textblock}{.8}(10.5,2.7)\fontsize{14}{16}\sf
\begin{alertblock}{}\centerline{$\phantom{\hat{y}_{|}}{e}_{t}\phantom{\hat{y}_{|}}$}\end{alertblock}\end{textblock}}

## Facebook closing stock price

```{r dj4, echo=TRUE, warning=FALSE, fig.height=2.8}
augment(fit) |>
  ggplot(aes(x = trading_day)) +
  geom_line(aes(y = Close, colour = "Data")) +
  geom_line(aes(y = .fitted, colour = "Fitted"))
```

## Facebook closing stock price

```{r dj4a, echo=TRUE, warning=FALSE, fig.height=2.6}
augment(fit) |>
  filter(trading_day > 1100) |>
  ggplot(aes(x = trading_day)) +
  geom_line(aes(y = Close, colour = "Data")) +
  geom_line(aes(y = .fitted, colour = "Fitted"))
```

## Facebook closing stock price

```{r dj5, echo=TRUE, warning = FALSE, fig.height=2.8}
augment(fit) |>
  autoplot(.resid) +
  labs(y = "$US",
       title = "Residuals from naïve method")
```

## Facebook closing stock price

```{r dj6, warning=FALSE, fig.height=2.8}
augment(fit) |>
  ggplot(aes(x = .resid)) +
  geom_histogram(bins = 150) +
  labs(title = "Histogram of residuals")
```

## Facebook closing stock price

```{r dj7, fig.height=2.8}
augment(fit) |>
  ACF(.resid) |>
  autoplot() + labs(title = "ACF of residuals")
```

## `gg_tsresiduals()` function

```{r dj10, echo=TRUE, warning = FALSE}
gg_tsresiduals(fit)
```

## ACF of residuals

  * We assume that the residuals are white noise (uncorrelated, mean zero, constant variance). If they aren't, then there is information left in  the residuals that should be used in computing forecasts.

  * So a standard residual diagnostic is to check the ACF of the residuals of a forecasting method.

  * We *expect* these to look like white noise.

## Portmanteau tests
\begin{block}{}
$r_k = $ autocorrelation of residual at lag $k$
\end{block}\vspace*{-0.3cm}

Consider a *whole set* of $r_{k}$ values, and develop a test to see whether the set is significantly different from a zero set.\pause

\begin{block}{Box-Pierce test\phantom{g}}
\centerline{$\displaystyle
Q = T \sum_{k=1}^\ell r_k^2$}
where $\ell$  is max lag being considered and $T$ is number of observations.
\end{block}

  * If each $r_k$ close to zero, $Q$ will be **small**.
  * If some $r_k$ values large (positive or negative), $Q$ will be **large**.

\vspace*{10cm}

## Portmanteau tests
\begin{block}{}
$r_k = $ autocorrelation of residual at lag $k$
\end{block}\vspace*{-0.3cm}

Consider a *whole set* of $r_{k}$  values, and develop a test to see whether the set is significantly different from a zero set.

\begin{block}{Ljung-Box test}
\centerline{$\displaystyle
 Q^* = T(T+2) \sum_{k=1}^\ell (T-k)^{-1}r_k^2$}
where $\ell$  is max lag being considered and $T$ is number of observations.
\end{block}

  * My preferences: $\ell=10$ for non-seasonal data, $h=2m$ for seasonal data (where $m$ is seasonal period).
  * Better performance, especially in small samples.

\vspace*{10cm}

## Portmanteau tests

  * If data are WN, $Q^*$ has $\chi^2$ distribution with  $(\ell - K)$ degrees of freedom where $K=$ no.\ parameters in model.
  * When applied to raw data, set $K=0$.
  * `lag` $= \ell$, `dof` $= K$

\fontsize{11}{12}\sf

```{r dj9, echo=TRUE}
augment(fit) |>
  features(.resid, ljung_box, lag=10, dof=0)
```

<!-- ## Your turn

Compute seasonal naïve forecasts for quarterly Australian beer production from 1992.

\fontsize{10}{12}\sf

```{r, results = 'hide', fig.show='hide'}
recent <- aus_production |> filter(year(Quarter) >= 1992)
fit <- recent |> model(SNAIVE(Beer))
fit |> forecast() |> autoplot(recent)
```

\fontsize{14}{15}\sf

Test if the residuals are white noise.

\fontsize{10}{12}\sf

```{r, results = 'hide', fig.show='hide', warning = FALSE}
augment(fit) |> features(.resid, ljung_box, lag=10, dof=0)
gg_tsresiduals(fit)
```

\fontsize{14}{15}\sf

What do you conclude? -->
