---
title: "5. The forecaster's toolbox"
author: "5.5 Distributional forecasts"
date: "OTexts.org/fpp3/"
classoption: aspectratio=169
titlepage: fpp3title.png
titlecolor: fpp3red
toc: false
output:
  binb::monash:
    colortheme: monashwhite
    fig_width: 7.5
    fig_height: 3
    keep_tex: no
    includes:
      in_header: header.tex
---

```{r setup, include=FALSE}
source("setup.R")
```


```{r brick-model}
brick_fit <-  aus_production |>
  filter(!is.na(Bricks)) |>
  model(
    Seasonal_naive = SNAIVE(Bricks),
    Naive = NAIVE(Bricks),
    Drift = RW(Bricks ~ drift()),
    Mean = MEAN(Bricks)
  )
```

## Forecast distributions

 * A forecast $\hat{y}_{T+h|T}$ is (usually) the mean of the conditional distribution $y_{T+h} \mid y_1, \dots, y_{T}$.
 * Most time series models produce normally distributed forecasts.
 * The forecast distribution describes the probability of observing any future value.

## Forecast distributions

Assuming residuals are normal, uncorrelated, sd = $\hat\sigma$:

\begin{block}{}
\begin{tabular}{ll}
\bf Mean: & $y_{T+h|T} \sim N(\bar{y}, (1 + 1/T)\hat{\sigma}^2)$\\[0.2cm]
\bf Naïve: & $y_{T+h|T} \sim N(y_T, h\hat{\sigma}^2)$\\[0.2cm]
\bf Seasonal naïve: & $y_{T+h|T} \sim N(y_{T+h-m(k+1)}, (k+1)\hat{\sigma}^2)$\\[0.2cm]
\bf Drift: & $y_{T+h|T} \sim N(y_T + \frac{h}{T-1}(y_T - y_1),h\frac{T+h}{T}\hat{\sigma}^2)$
\end{tabular}
\end{block}

where $k$ is the integer part of $(h-1)/m$.

Note that when $h=1$ and $T$ is large, these all give the same approximate forecast variance: $\hat{\sigma}^2$.

## Prediction intervals

 * A prediction interval gives a region within which we expect $y_{T+h}$ to lie with a specified probability.
 * Assuming forecast errors are normally distributed, then a 95% PI is
 \begin{alertblock}{}
\centerline{$
  \hat{y}_{T+h|T} \pm 1.96 \hat\sigma_h
$}
\end{alertblock}
where $\hat\sigma_h$ is the st dev of the $h$-step distribution.

 * When $h=1$, $\hat\sigma_h$ can be estimated from the residuals.

## Prediction intervals
\fontsize{10}{12}\sf

```{r brick-fc-interval}
brick_fit <-  aus_production |>
  filter(!is.na(Bricks)) |>
  model(Seasonal_naive = SNAIVE(Bricks)) |>
  forecast(h = "5 years") |>
  hilo(level = 95)
```

## Prediction intervals

 * Point forecasts often useless without a measure of uncertainty (such as prediction intervals).
 * Prediction intervals require a stochastic model (with random errors, etc).
 * For most models, prediction intervals get wider as the forecast horizon increases.
 * Use `level` argument to control coverage.
 * Check residual assumptions before believing them.
 * Usually too narrow due to unaccounted uncertainty.
