---
title: "5. The forecaster's toolbox"
author: "5.6 Forecasting using transformations"
date: "OTexts.org/fpp3/"
classoption: aspectratio=169
titlepage: fpp3title.png
titlecolor: fpp3red
toc: false
output:
  binb::monash:
    colortheme: monashwhite
    fig_width: 7.5
    fig_height: 3
    keep_tex: no
    includes:
      in_header: fpp3header.tex
---

```{r setup, include=FALSE}
source("setup.R")
```

## Modelling with transformations

```{r food, echo=TRUE, fig.height=2.65}
eggs <- prices |>
  filter(!is.na(eggs)) |>
  select(eggs)
eggs |> autoplot() +
  labs(title = "Annual egg prices", y = "$US (adjusted for inflation)")
```

## Modelling with transformations

Transformations used in the left of the formula will be automatically back-transformed. To model log-transformed egg prices, you could use:

\fontsize{13}{15}\sf

```{r food-bt-fit, dependson='food'}
fit <- eggs |>
  model(RW(log(eggs) ~ drift()))
fit
```

## Forecasting with transformations

\fontsize{11}{12}\sf

```{r food-bt-fc, dependson='food-bt-fit'}
fc <- fit |>
  forecast(h = 50)
fc
```

## Forecasting with transformations

```{r elec9,echo=TRUE,fig.height=3, dependson='food-bt-fc'}
fc |> autoplot(eggs) +
  labs(title = "Annual egg prices",
       y = "US$ (adjusted for inflation)")
```

## Bias adjustment

  * Back-transformed point forecasts are medians.
  * Back-transformed PI have the correct coverage.

\pause

**Back-transformed means**

Let $X$ be have mean $\mu$ and variance $\sigma^2$.

Let $f(x)$ be back-transformation function, and $Y=f(X)$.

Taylor series expansion about $\mu$:
$$f(X) = f(\mu) + (X-\mu)f'(\mu) + \frac{1}{2}(X-\mu)^2f''(\mu).$$\pause\vspace*{-0.4cm}

\begin{alertblock}{}
\centerline{$\E[Y] = \E[f(X)] = f(\mu) + \frac12 \sigma^2 f''(\mu)$}
\end{alertblock}

## Bias adjustment

\fontsize{13}{15}\sf

**Box-Cox back-transformation:**
\begin{align*}
y_t &= \left\{\begin{array}{ll}
        \exp(w_t)      & \quad \lambda = 0; \\
        (\lambda W_t+1)^{1/\lambda}  & \quad \lambda \ne 0.
\end{array}\right. \\
f(x) &= \begin{cases}
                        e^x & \quad\lambda=0;\\
 (\lambda x + 1)^{1/\lambda} & \quad\lambda\ne0.
 \end{cases}\\
f''(x) &= \begin{cases}
                        e^x & \quad\lambda=0;\\
 (1-\lambda)(\lambda x + 1)^{1/\lambda-2} & \quad\lambda\ne0.
 \end{cases}
\end{align*}\pause
\begin{alertblock}{}
\centerline{$\E[Y] = \begin{cases}
                        e^\mu\left[1+\frac{\sigma^2}{2}\right] & \quad\lambda=0;\\
 (\lambda \mu + 1)^{1/\lambda}\left[1+\frac{\sigma^2(1-\lambda)}{2(\lambda\mu+1)^2}\right] & \quad\lambda\ne0.
 \end{cases}$}
\end{alertblock}

## Bias adjustment

```{r biasadj, fig.height=2.8, message=FALSE}
fc |>
  autoplot(eggs, level = 80, point_forecast = lst(mean, median)) +
  labs(title = "Annual egg prices",
       y = "US$ (adjusted for inflation)")
```
