---
title: "5. The forecaster's toolbox"
author: "5.7 Forecasting with decomposition"
date: "OTexts.org/fpp3/"
classoption: aspectratio=169
titlepage: fpp3title.png
titlecolor: fpp3red
toc: false
output:
  binb::monash:
    colortheme: monashwhite
    fig_width: 7.5
    fig_height: 3
    keep_tex: no
    includes:
      in_header: header.tex
---

```{r setup, include=FALSE}
source("setup.R")
```

## Forecasting and decomposition

\begin{block}{}
$$y_t = \hat{S}_t + \hat{A}_t$$
\vspace*{-0.5cm}\begin{itemize}\tightlist
  \item $\hat{A}_t$ is seasonally adjusted component
  \item $\hat{S}_t$ is seasonal component.
\end{itemize}
\end{block}

  *  Forecast $\hat{S}_t$ using SNAIVE.
  *  Forecast $\hat{A}_t$ using non-seasonal time series method.
  *  Combine forecasts of $\hat{S}_t$ and $\hat{A}_t$ to get forecasts of original data.

## US Retail Employment
\fontsize{10}{11}\sf

```{r usretail}
us_retail_employment <- us_employment |>
  filter(year(Month) >= 1990, Title == "Retail Trade") |>
  select(-Series_ID)
us_retail_employment
```

\vspace*{10cm}

## US Retail Employment
\fontsize{10}{11}\sf

```{r usretail1, echo=TRUE, fig.height=3.2}
dcmp <- us_retail_employment |>
  model(STL(Employed)) |>
  components() |> select(-.model)
dcmp
```

\vspace*{10cm}

## US Retail Employment

```{r usretail2, echo=TRUE, fig.height=2.6}
dcmp |>
  model(NAIVE(season_adjust)) |>
  forecast() |>
  autoplot(dcmp) +
  labs(title = "Naive forecasts of seasonally adjusted data")
```

\vspace*{10cm}

## US Retail Employment
\fontsize{10}{11}\sf

```{r usretail3, echo=TRUE, fig.height=2.4}
us_retail_employment |>
  model(stlf = decomposition_model(
    STL(Employed ~ trend(window = 7), robust = TRUE),
    NAIVE(season_adjust)
  )) |>
  forecast() |>
  autoplot(us_retail_employment)
```

\vspace*{10cm}

## Decomposition models

`decomposition_model()` creates a decomposition model

 * You must provide a method for forecasting the `season_adjust` series.
 * A seasonal naive method is used by default for the `seasonal` components.
 * The variances from both the seasonally adjusted and seasonal forecasts are combined.
