---
title: "ETC3550 Applied&nbsp;forecasting&nbsp;for business&nbsp;and&nbsp;economics"
author: "Ch7. Time series decomposition"
date: "OTexts.org/fpp3/"
toc: true
colortheme: monashwhite
output:
  binb::monash:
    fig_width: 7
    fig_height: 3.5
    includes:
      in_header: header.tex
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, cache=TRUE)

library(tidyverse)
library(fable)
library(tsibble)
library(feasts)
library(lubridate)

elecequip <- as_tsibble(fpp2::elecequip)

options(width=65)
```

# Time series components

## Time series patterns

**Recall**

Trend
:  pattern exists when there is a long-term increase or decrease in the data.

Cyclic
: pattern exists when data exhibit rises and falls that are *not of fixed period* (duration usually of at least 2 years).

Seasonal
: pattern exists when a series is influenced by seasonal factors (e.g., the quarter of the year, the month, or day of the week).

## Time series decomposition

\begin{block}{}\vspace*{-0.3cm}
\[ y_t = f(S_t, T_t, R_t) \]
\end{block}
\begin{tabular}{@{}llp{8cm}@{}}
where & $y_t=$ & data at period $t$ \\
      & $T_t=$ & trend-cycle component at period $t$\\
      & $S_t=$ & seasonal component at period $t$ \\
      & $R_t=$ & remainder component at period $t$
\end{tabular}
\pause

**Additive decomposition:** $y_t = S_t + T_t + R_t.$

**Multiplicative decomposition:** $y_t = S_t \times T_t \times R_t.$

## Time series decomposition
\fontsize{13}{15}\sf

  *  Additive model  appropriate if  magnitude of  seasonal fluctuations does not vary with level.
  *  If seasonal are proportional to level of series, then multiplicative model appropriate.
  *  Multiplicative decomposition more prevalent with economic series
  *  Alternative: use a Box-Cox transformation, and then use additive decomposition.
  *  Logs turn multiplicative relationship into an additive relationship:

$$y_t = S_t \times T_t \times E_t \quad\Rightarrow\quad
\log y_t = \log S_t + \log T_t + \log R_t.
$$

## Decomposition dable

\fontsize{10}{11}\sf

```{r dable}
dcmp <- elecequip %>% STL(value ~ season(window = 7))
dcmp
```

## Euro electrical equipment

\fontsize{11}{14}\sf

```{r elecequip-stl, fig.width=8, fig.height=5}
autoplot(dcmp) + xlab("Year")
```

## Euro electrical equipment

```{r elecequip3}
dcmp %>% gg_subseries(season_year)
```

## Euro electrical equipment

\fontsize{10}{13}\sf

```{r elecequip-trend, message=FALSE, warning=FALSE}
autoplot(elecequip, series="Data") +
  autolayer(dcmp, trend, series="Trend-cycle")
```


## Your turn

Repeat the decomposition using

\fontsize{11}{14}\sf

```r
elecequip %>%
  STL(value ~ season(window=7) + trend(window=11)) %>%
  autoplot()
```

\fontsize{14}{15}\sf

What happens as you change `s.window` and `t.window`?

# Seasonal adjustment

## Seasonal adjustment

  *  Useful by-product of decomposition:  an easy way to calculate seasonally adjusted data.
  *  Additive decomposition: seasonally adjusted data given by
$$y_t - S_t = T_t + R_t$$
  *  Multiplicative decomposition: seasonally adjusted data given by
$$y_t / S_t = T_t \times R_t$$

## Euro electrical equipment

\fontsize{11}{12}\sf

```r
dcmp <- elecequip %>% STL(value ~ season(window=7))
elecequip %>% 
  autoplot(value, series="Data") +
    autolayer(dcmp, trend + remainder, 
      series="Seasonally Adjusted")
```

```{r elecequip-sa, echo=FALSE}
elecequip %>% autoplot(value, series="Data") +
  autolayer(dcmp, seas_adjust, series="Seasonally Adjusted") +
  xlab("Year") + ylab("New orders index") +
  ggtitle("Electrical equipment manufacturing (Euro area)") +
  scale_colour_manual(breaks = c("Seasonally adjusted", "Data"), values=c("blue","grey"))
```

## Seasonal adjustment

  * We use estimates of $S$ based on past values to seasonally adjust a current value.
  *  Seasonally adjusted  series reflect **remainders** as well as **trend**. Therefore they are not "smooth"" and "downturns"" or "upturns" can be misleading.
  *  It is better to use the trend-cycle component to look for turning points.

## The ABS stuff-up

\fullheight{abs1}

## The ABS stuff-up

\fullheight{abs2}

## The ABS stuff-up

\fullheight{abs3}

## The ABS stuff-up

\fontsize{11}{14}\sf

```{r abs, echo=FALSE, fig.width=6, fig.height=4}
x <- tsibble(
  Time = yearmonth("1978 Feb") + 0:439,
  Employed = c(5985.7,6040.6,6054.2,6038.3,6031.3,6036.1,6005.4,6024.3,6045.9,6033.8,6125.4,5971.3,
  6050.7,6096.2,6087.7,6075.6,6095.7,6103.9,6078.5,6157.8,6164.0,6188.8,6257.2,6112.9,
  6207.2,6278.7,6224.9,6273.4,6269.9,6314.1,6281.4,6360.0,6320.2,6342.0,6426.6,6253.0,
  6356.5,6428.1,6426.3,6412.4,6413.9,6425.3,6393.7,6502.7,6445.3,6433.3,6506.9,6355.5,
  6432.4,6497.4,6431.6,6440.9,6414.3,6425.9,6379.3,6443.5,6421.1,6366.8,6370.1,6172.0,
  6263.9,6310.3,6254.5,6272.8,6266.5,6295.0,6241.2,6358.2,6336.1,6377.5,6456.5,6251.4,
  6365.4,6503.2,6477.6,6489.7,6499.0,6528.7,6466.1,6579.8,6553.2,6576.1,6636.0,6452.4,
  6595.7,6657.4,6588.8,6657.9,6659.4,6703.4,6675.5,6814.7,6771.1,6881.9,6910.8,6753.6,
  6861.9,6961.9,6997.9,6979.0,7007.7,6991.5,6918.5,7040.6,7030.4,7034.2,7116.8,6902.5,
  7022.3,7133.4,7109.6,7103.5,7128.9,7175.6,7092.3,7186.5,7177.4,7182.2,7330.7,7169.4,
  7247.3,7397.4,7383.4,7354.8,7378.3,7383.1,7353.3,7503.2,7477.3,7508.6,7622.9,7423.8,
  7566.5,7634.6,7678.4,7720.8,7711.0,7740.8,7715.3,7841.6,7806.5,7862.4,7935.5,7707.7,
  7803.0,7874.1,7887.9,7908.5,7900.3,7919.4,7808.0,7905.5,7848.9,7826.9,7915.5,7641.3,
  7708.7,7715.4,7717.2,7703.7,7678.1,7583.0,7620.7,7713.2,7638.0,7614.9,7712.2,7518.9,
  7597.2,7646.2,7644.1,7631.4,7637.3,7668.3,7613.4,7709.7,7665.7,7587.4,7693.4,7533.7,
  7531.0,7645.7,7572.6,7620.5,7627.9,7646.5,7589.4,7747.6,7738.8,7744.9,7842.1,7646.8,
  7738.6,7824.2,7827.4,7857.9,7878.4,7966.0,7861.7,8054.4,7997.2,8003.3,8135.5,7928.4,
  8049.9,8118.1,8174.6,8165.2,8205.6,8229.0,8165.9,8300.4,8232.6,8300.3,8395.7,8166.7,
  8246.6,8280.4,8248.0,8297.1,8311.7,8332.1,8265.9,8373.0,8319.4,8314.4,8431.4,8235.2,
  8291.4,8347.5,8343.1,8330.2,8345.6,8374.9,8250.3,8474.0,8405.2,8462.1,8540.5,8334.7,
  8413.0,8460.0,8499.9,8482.5,8516.8,8541.9,8455.2,8653.2,8601.0,8554.3,8696.5,8477.4,
  8556.5,8618.9,8631.9,8606.5,8673.2,8706.7,8603.6,8777.1,8755.3,8763.7,8900.7,8628.2,
  8754.4,8830.7,8882.2,8865.0,8922.0,9020.0,8911.6,9061.3,8973.1,8912.7,9059.6,8834.9,
  8920.9,8956.0,9023.6,9004.6,9021.9,9048.9,8971.9,9105.9,9058.7,9055.6,9177.1,8993.4,
  9092.3,9128.5,9129.5,9134.7,9180.8,9194.5,9150.3,9303.5,9249.1,9286.7,9439.7,9281.7,
  9372.6,9362.1,9365.6,9380.1,9370.4,9363.9,9327.0,9486.1,9447.8,9427.7,9573.6,9363.8,
  9434.5,9506.4,9512.0,9533.5,9543.3,9553.1,9462.1,9668.6,9662.2,9652.9,9807.8,9634.4,
  9744.6,9828.3,9856.3,9850.8,9896.6,9912.3,9870.3,10004.6,9949.7,9945.0,10074.7,9842.7,
  9961.1,10048.7,10041.0,10082.1,10120.8,10170.8,10105.8,10299.5,10212.4,10201.6,10404.3,
  10156.1,10277.0,10349.2,10362.9,10412.0,10436.3,10456.8,10406.4,10588.8,10520.5,10535.0,
  10710.1,10524.9,10622.9,10677.4,10706.2,10690.3,10745.0,10761.9,10710.4,10854.5,10807.4,
  10757.3,10915.6,10681.0,10776.7,10775.2,10792.7,10786.8,10770.9,10808.8,10707.3,10882.1,
  10845.2,10829.2,11010.9,10809.9,10889.2,10928.9,10940.1,10957.4,11009.3,11030.5,10973.8,
  11159.4,11129.0,11144.5,11295.0,11063.7,11146.2,11217.0,11186.5,11196.2,11221.3,11227.5,
  11130.7,11321.2,11274.0,11240.6,11354.8,11159.0,11236.2,11332.4,11328.3,11389.0,11350.6,
  11363.7,11259.8,11452.6,11401.9,11375.0,11518.4,11304.0,11424.3,11436.3,11482.2,11495.6,
  11497.8,11486,11369,11547,11499,11472,11571,11354,11493,11562,11589,11595,11602,11590,
  11622,11593),
  index = Time)

x %>% autoplot(Employed) +
  ggtitle("Total employed") +
  ylab("Thousands") + xlab("Year")
```

## The ABS stuff-up

```{r, echo=FALSE, fig.width=6, fig.height=4}
x %>% 
  filter(year(Time) >= 2005) %>% 
  autoplot(Employed) +
  ggtitle("Total employed") +
  ylab("Thousands") + xlab("Year")
```

## The ABS stuff-up

\fontsize{10}{14}\sf

```{r}
x %>% 
  filter(year(Time) >= 2005) %>% 
  gg_season(Employed, label='right') +
  ggtitle("Total employed") + ylab("Thousands")
```

## The ABS stuff-up

```{r, fig.height=2, echo=FALSE}
x %>% 
  mutate(Month = month(Time, label = TRUE), Year = year(Time)) %>% 
  filter(Month %in% c("Aug", "Sep")) %>% 
  update_tsibble(key = id(Month), index = Year) %>% 
  select(-Time) %>% 
  spread(Month, Employed) %>% 
  ggplot() +
  geom_boxplot(aes(y= Sep - Aug, x = 1)) + coord_flip() +
  ggtitle("Sep - Aug: total employed") + xlab("") + ylab("Thousands") +
  scale_x_continuous(breaks=NULL, labels=NULL)
```

## The ABS stuff-up

\fontsize{11}{11}\sf

```{r, fig.height=3.5}
dcmp <- x %>% 
  filter(year(Time) >= 2009) %>%
  STL(Employed ~ season(window = 11), robust = TRUE)
autoplot(dcmp)
```

## The ABS stuff-up

```{r, echo=FALSE}
dcmp %>% 
  filter(year(Time) == 2013) %>% 
  gg_season(season_year) +
  ggtitle("Seasonal component") +
  guides(colour="none")
```

## The ABS stuff-up

```{r}
dcmp %>% as_tsibble %>% 
  autoplot(trend + remainder)
```

## The ABS stuff-up
\fontsize{13}{15}\sf

  *  August 2014 employment numbers higher than expected.
  *  Supplementary survey usually conducted in August for employed people.
  *  Most likely, some employed people were claiming to be unemployed in August to avoid supplementary questions.
  *  Supplementary survey not run in 2014, so no motivation to lie about employment.
  *  In previous years, seasonal adjustment fixed the problem.
  *  The ABS has now adopted a new method to avoid the bias.

## History of time series decomposition

\fontsize{13}{15}\sf

  *  Classical method originated in 1920s.
  *  Census II method introduced in 1957. Basis for X-11 method and variants (including X-12-ARIMA, X-13-ARIMA)
  *  STL method introduced in 1983
  *  TRAMO/SEATS introduced in 1990s.
\pause

### National Statistics Offices
 * ABS uses X-12-ARIMA
 * US Census Bureau uses X-13-ARIMA-SEATS
 * Statistics Canada uses X-12-ARIMA
 * ONS (UK) uses X-12-ARIMA
 * EuroStat use X-13-ARIMA-SEATS

# X-11 decomposition

## X-11 decomposition

\fontsize{10}{12}\sf

```{r x11, warning=FALSE, fig.width=8, fig.height=4.5}
elecequip %>% X11(value, type="multiplicative") %>%
  autoplot() +
  ggtitle("X11 decomposition of electrical equipment index")
```

## (Dis)advantages of X-11

**Advantages**

  *  Relatively robust to outliers
  *  Completely automated choices for trend and seasonal changes
  *  Very widely tested on economic data over a long period of time.

\pause

**Disadvantages**

  *  No prediction/confidence intervals
  *  Ad hoc method with no underlying model
  *  Only developed for quarterly and monthly data

## Extensions: X-12-ARIMA and X-13-ARIMA

  *  The X-11, X-12-ARIMA and X-13-ARIMA methods are based on Census II decomposition.
  *  These allow adjustments for trading days and other explanatory variables.
  *  Known outliers can be omitted.
  *  Level shifts and ramp effects can be modelled.
  *  Missing values estimated and replaced.
  *  Holiday factors (e.g., Easter, Labour Day) can be estimated.

# SEATS decomposition

## SEATS decomposition

\fontsize{10}{12}\sf

```{r seats, warning=FALSE, fig.width=8, fig.height=4.5}
elecequip %>% SEATS(value) %>%
  autoplot() +
  ggtitle("SEATS decomposition of electrical equipment index")
```

## SEATS decomposition

```{r seats-seasadj, echo=FALSE, warning=FALSE, fig.width=9, fig.height=5.5}
fit <- elecequip %>% SEATS(value)
autoplot(elecequip, value, series="Data") +
  autolayer(fit, trend, series="Trend") +
  autolayer(fit, seas_adjust, series="Seasonally Adjusted") +
  xlab("Year") + ylab("New orders index") +
  ggtitle("Electrical equipment manufacturing (Euro area)") +
  scale_colour_manual(values=c("gray","blue","red"),
                     breaks=c("Data","Seasonally Adjusted","Trend"))
```

## SEATS decomposition

\fontsize{11}{14}\sf

```{r elecequip-seats, echo=TRUE}
fit %>% 
  gg_subseries(seasonal) + ylab("Seasonal")
```

## (Dis)advantages of SEATS

**Advantages**

  * Model-based
  * Smooth trend estimate
  * Allows estimates at end points
  * Allows changing seasonality
  * Developed for economic data

\pause

**Disadvantages**

  *  Only developed for quarterly and monthly data

# STL decomposition

## STL decomposition

\fontsize{13}{14}\sf

  *  STL: "Seasonal and Trend decomposition using Loess"
  *  Very versatile and robust.
  *  Unlike X-12-ARIMA, STL will handle any type of seasonality.
  *  Seasonal component allowed to change over time, and rate of change controlled by user.
  *  Smoothness of trend-cycle also controlled by user.
  *  Robust to outliers
  *  Not trading day or calendar adjustments.
  *  Only additive.
  *  Take logs to get multiplicative decomposition.
  *  Use Box-Cox transformations to get other decompositions.

## STL decomposition

\fontsize{10}{14}\sf

```{r stlagain, echo=TRUE, warning=FALSE, fig.width=8, fig.height=4}
dcmp <- elecequip %>% 
  STL(value ~ season(window = 5), robust = TRUE)
autoplot(dcmp) +
  ggtitle("STL decomposition of electrical equipment index")
```

## STL decomposition

\fontsize{10}{14}\sf

```{r stlagain2, echo=TRUE, warning=FALSE, fig.width=8, fig.height=4}
fit <- stl(elecequip, s.window="periodic", robust=TRUE)
autoplot(fit) +
  ggtitle("STL decomposition of electrical equipment index")
```

## STL decomposition

\fontsize{10}{14}\sf

```r
elecequip %>% 
  STL(value ~ season(window = 5))

elecequip %>% 
  STL(value ~ trend(window=15) + season(window="periodic"),
      robust = TRUE)
```

\fontsize{14}{16}\sf

  *  `trend(window = ?)` controls wiggliness of trend component.
  *  `season(window = ?)` controls variation on seasonal component.

## STL decomposition
\fontsize{12}{14}\sf

```{r mstl, fig.width=8, fig.height=4}
elecequip %>% STL(value) %>% autoplot()
```

* `STL()` chooses `s.window=13` by default
* Can include transformations.

# Forecasting and decomposition

## Forecasting and decomposition

  *  Forecast seasonal component by repeating the last year
  *  Forecast seasonally adjusted data using non-seasonal time series method.
  *  Combine forecasts of seasonal component with forecasts of seasonally adjusted data to get forecasts of original data.
  *  Sometimes a decomposition is useful just for understanding the data before building a separate forecasting model.

## Electrical equipment

\fontsize{10}{11}\sf

```{r elecequip4, echo=TRUE, fig.height=3.2}
dcmp <- elecequip %>% 
  STL(value ~ trend(window=13) + season(window="periodic"))
dcmp %>% 
  model(NAIVE(seas_adjust)) %>% forecast() %>% 
  autoplot(dcmp) + ylab("New orders index") +
  ggtitle("ETS forecasts of seasonally adjusted data")
```

\vspace*{10cm}

## Electrical equipment

\fontsize{10}{11}\sf

```{r elecequip5, echo=TRUE, fig.height=3.2}
dcmp_def <- dcmp_model(STL, 
  value ~ trend(window = 13) + season(window = "periodic"),
  NAIVE(seas_adjust))
elecequip %>%
  model(STLM = dcmp_def) %>% forecast() %>%
  autoplot(elecequip) + ylab("New orders index") + xlab("Year")
```

\vspace*{10cm}

## Decomposition and prediction intervals

  *  It is common to take the prediction intervals from the seasonally adjusted forecasts and modify them with the seasonal component.
  *  This ignores the uncertainty in the seasonal component estimate.
  *  It also ignores the uncertainty in the future seasonal pattern.
