---
title: "7. Time series regression models"
author: "7.1 The linear model"
date: "OTexts.org/fpp3/"
classoption: aspectratio=169
titlepage: fpp3title.png
titlecolor: fpp3red
toc: false
output:
  binb::monash:
    colortheme: monashwhite
    fig_width: 7.5
    fig_height: 3
    keep_tex: no
    includes:
      in_header: fpp3header.tex
---

```{r setup, include=FALSE}
source("setup.R")
```

## Multiple regression and forecasting

\begin{block}{}\vspace*{-0.3cm}
\[
  y_t = \beta_0 + \beta_1 x_{1,t} + \beta_2 x_{2,t} + \cdots + \beta_kx_{k,t} + \varepsilon_t.
\]
\end{block}

* $y_t$ is the variable we want to predict: the "response" variable
* Each $x_{j,t}$ is numerical and is called a "predictor".
 They are usually assumed to be known for all past and future times.
* The coefficients $\beta_1,\dots,\beta_k$ measure the effect of each
predictor after taking account of the effect of all other predictors
in the model.

That is, the coefficients measure the **marginal effects**.

* $\varepsilon_t$ is a white noise error term

## Example: US consumption expenditure

```{r MultiPredictors, echo=FALSE, fig.height=5, fig.width=10, out.height="120%"}
us_change |>
  gather("Measure", "Change", Consumption, Income, Production, Savings, Unemployment) |>
  ggplot(aes(x = Quarter, y = Change, colour = Measure)) +
  geom_line() +
  facet_grid(vars(Measure), scales = "free_y") +
  labs(y = "") +
  guides(colour = "none")
```

## Example: US consumption expenditure

```{r ScatterMatrix, echo=FALSE, fig.height=5, fig.width=10, out.height="100%"}
us_change |>
  as_tibble() |>
  select(-Quarter) |>
  GGally::ggpairs()
```

## Example: US consumption expenditure

\fontsize{7}{6.8}\sf

```{r usestim, echo=TRUE}
fit_consMR <- us_change |>
  model(lm = TSLM(Consumption ~ Income + Production + Unemployment + Savings))
report(fit_consMR)
```

## Example: US consumption expenditure

```{r usfitted1, echo=FALSE}
augment(fit_consMR) |>
  ggplot(aes(x = Quarter)) +
  geom_line(aes(y = Consumption, colour = "Data")) +
  geom_line(aes(y = .fitted, colour = "Fitted")) +
  labs(y = NULL, title = "Percent change in US consumption expenditure") +
  scale_colour_manual(values = c(Data = "black", Fitted = "#D55E00")) +
  guides(colour = guide_legend(title = NULL))
```

## Example: US consumption expenditure

```{r usfitted2, echo=FALSE, message=FALSE, warning=FALSE}
augment(fit_consMR) |>
  ggplot(aes(y = .fitted, x = Consumption)) +
  geom_point() +
  labs(
    y = "Fitted (predicted values)",
    x = "Data (actual values)",
    title = "Percentage change in US consumption expenditure"
  ) +
  geom_abline(intercept = 0, slope = 1)
```

## Example: US consumption expenditure

```{r, echo=TRUE}
fit_consMR |> gg_tsresiduals()
```
