---
title: "7. Time series regression models"
author: "7.4 Some useful predictors"
date: "OTexts.org/fpp3/"
classoption: aspectratio=169
titlepage: fpp3title.png
titlecolor: fpp3red
toc: false
output:
  binb::monash:
    colortheme: monashwhite
    fig_width: 7.5
    fig_height: 3
    keep_tex: no
    includes:
      in_header: fpp3header.tex
---

```{r setup, include=FALSE}
source("setup.R")
```

## Trend

**Linear trend**
\[
  x_t = t
\]

* $t=1,2,\dots,T$
* Strong assumption that trend will continue.

## Nonlinear trend

**Piecewise linear trend with bend at $\tau$**
\begin{align*}
x_{1,t} &= t \\
x_{2,t} &= \left\{ \begin{array}{ll}
  0 & t <\tau\\
  (t-\tau) & t \ge \tau
\end{array}\right.
\end{align*}
\pause\vspace*{-.6cm}

**Quadratic or higher order trend**
\[
  x_{1,t} =t,\quad x_{2,t}=t^2,\quad \dots
\]
\pause\vspace*{-0.5cm}

\centerline{\textcolor{orange}{\textbf{NOT RECOMMENDED!}}}

## Dummy variables

\begin{textblock}{6}(0.4,1.5)
If a categorical variable takes only two values (e.g., `Yes'
or `No'), then an equivalent numerical variable can be constructed
taking value 1 if yes and 0 if no. This is called a \textbf{dummy variable}.
\end{textblock}

\placefig{7.7}{1.5}{width=5cm}{dummy2}

## Dummy variables

\begin{textblock}{6}(0.4,1.5)
If there are more than two categories, then the variable can
be coded using several dummy variables (one fewer than the total
number of categories).

\end{textblock}

\placefig{7.7}{1.5}{width=7.3cm}{dummy3}

## Beware of the dummy variable trap!
* Using one dummy for each category gives too many dummy variables!

* The regression will then be singular and inestimable.

* Either omit the constant, or omit the dummy for one category.

* The coefficients of the dummies are relative to the omitted category.

## Uses of dummy variables
\fontsize{13}{15}\sf

**Seasonal dummies**

* For quarterly data: use 3 dummies
* For monthly data: use 11 dummies
* For daily data: use 6 dummies
* What to do with weekly data?

\pause

**Outliers**

* If there is an outlier, you can use a dummy variable to remove its effect.

\pause

**Public holidays**

* For daily data: if it is a public holiday, dummy=1, otherwise dummy=0.

## Beer production revisited

```{r, echo=FALSE, fig.height=2.7}
recent_production <- aus_production |> filter(year(Quarter) >= 1992)
recent_production |> autoplot(Beer) +
  labs(y = "Megalitres", title = "Australian quarterly beer production")
```

\pause

\begin{block}{Regression model}
\centering
$y_t = \beta_0 + \beta_1 t + \beta_2d_{2,t} + \beta_3 d_{3,t} + \beta_4 d_{4,t} + \varepsilon_t$
\end{block}\vspace*{-0.3cm}

* $d_{i,t} = 1$ if $t$ is quarter $i$ and 0 otherwise.

## Beer production revisited
\fontsize{7}{7}\sf

```{r, echo=TRUE}
fit_beer <- recent_production |> model(TSLM(Beer ~ trend() + season()))
report(fit_beer)
```

## Beer production revisited

```{r, fig.height=2.4}
augment(fit_beer) |>
  ggplot(aes(x = Quarter)) +
  geom_line(aes(y = Beer, colour = "Data")) +
  geom_line(aes(y = .fitted, colour = "Fitted")) +
  labs(y = "Megalitres", title = "Australian quarterly beer production") +
  scale_colour_manual(values = c(Data = "black", Fitted = "#D55E00"))
```

## Beer production revisited

```{r, fig.height=2.4}
augment(fit_beer) |>
  ggplot(aes(x = Beer, y = .fitted, colour = factor(quarter(Quarter)))) +
  geom_point() +
  labs(y = "Fitted", x = "Actual values", title = "Quarterly beer production") +
  scale_colour_brewer(palette = "Dark2", name = "Quarter") +
  geom_abline(intercept = 0, slope = 1)
```

## Beer production revisited

```{r, echo=TRUE}
fit_beer |> gg_tsresiduals()
```

## Beer production revisited

```{r, echo=TRUE}
fit_beer |>
  forecast() |>
  autoplot(recent_production)
```

## Fourier series

Periodic seasonality can be handled using pairs of Fourier terms:
$$
s_{k}(t) = \sin\left(\frac{2\pi k t}{m}\right)\qquad c_{k}(t) = \cos\left(\frac{2\pi k t}{m}\right)
$$
$$
y_t = a + bt + \sum_{k=1}^K \left[\alpha_k s_k(t) + \beta_k c_k(t)\right] + \varepsilon_t$$

* Every periodic function can be approximated by sums of sin and cos terms for large enough $K$.
* Choose $K$ by minimizing AICc.
* Called "harmonic regression"

```r
TSLM(y ~ trend() + fourier(K))
```

## Harmonic regression: beer production

\fontsize{7}{7}\sf

```{r fourierbeer, echo=TRUE}
fourier_beer <- recent_production |> model(TSLM(Beer ~ trend() + fourier(K = 2)))
report(fourier_beer)
```

## Harmonic regression: eating-out expenditure

```{r cafe, echo=TRUE, fig.height=2.6}
aus_cafe <- aus_retail |>
  filter(Industry == "Cafes, restaurants and takeaway food services",
         year(Month) %in% 2004:2018) |>
  summarise(Turnover = sum(Turnover))
aus_cafe |> autoplot(Turnover)
```

## Harmonic regression: eating-out expenditure

\fontsize{7}{7}\sf

```{r cafefit, dependson='cafe', fig.height=5, echo=TRUE}
fit <- aus_cafe |>
  model(
    K1 = TSLM(log(Turnover) ~ trend() + fourier(K = 1)),
    K2 = TSLM(log(Turnover) ~ trend() + fourier(K = 2)),
    K3 = TSLM(log(Turnover) ~ trend() + fourier(K = 3)),
    K4 = TSLM(log(Turnover) ~ trend() + fourier(K = 4)),
    K5 = TSLM(log(Turnover) ~ trend() + fourier(K = 5)),
    K6 = TSLM(log(Turnover) ~ trend() + fourier(K = 6))
  )
glance(fit) |> select(.model, r_squared, adj_r_squared, AICc)
```

## Harmonic regression: eating-out expenditure

```{r, echo=FALSE}
cafe_plot <- function(...) {
  fit |>
    select(...) |>
    forecast() |>
    autoplot(aus_cafe) +
    labs(title = sprintf("Log transformed %s, trend() + fourier(K = %s)", model_sum(select(fit, ...)[[1]][[1]]), deparse(..1))) +
    geom_label(
      aes(x = yearmonth("2007 Jan"), y = 4250, label = paste0("AICc = ", format(AICc))),
      data = glance(select(fit, ...))
    ) +
    geom_line(aes(y = .fitted), colour = "red", augment(select(fit, ...))) +
    ylim(c(1500, 5100))
}
```

```{r cafe1, dependson='cafe', fig.height=5, echo=FALSE}
cafe_plot(K = 1)
```

## Harmonic regression: eating-out expenditure

```{r cafe2, dependson='cafe', fig.height=5, echo=FALSE}
cafe_plot(K = 2)
```

## Harmonic regression: eating-out expenditure

```{r cafe3, dependson='cafe', fig.height=5, echo=FALSE}
cafe_plot(K = 3)
```

## Harmonic regression: eating-out expenditure

```{r cafe4, dependson='cafe', fig.height=5, echo=FALSE}
cafe_plot(K = 4)
```

## Harmonic regression: eating-out expenditure

```{r cafe5, dependson='cafe', fig.height=5, echo=FALSE}
cafe_plot(K = 5)
```

## Harmonic regression: eating-out expenditure

```{r cafe6, dependson='cafe', fig.height=5, echo=FALSE}
cafe_plot(K = 6)
```

## Intervention variables

**Spikes**

* Equivalent to a dummy variable for handling an outlier.
\pause

**Steps**

* Variable takes value 0 before the intervention and 1 afterwards.
\pause

**Change of slope**

* Variables take values 0 before the intervention and values $\{1,2,3,\dots\}$ afterwards.

## Holidays

**For monthly data**

* Christmas: always in December so part of monthly seasonal effect
* Easter: use a dummy variable $v_t=1$ if any part of Easter is in that month, $v_t=0$ otherwise.
* Ramadan and Chinese new year similar.

## Distributed lags

Lagged values of a predictor.

Example: $x$ is advertising which has a delayed effect

\begin{align*}
  x_{1} &= \text{advertising for previous month;} \\
  x_{2} &= \text{advertising for two months previously;} \\
        & \vdots \\
  x_{m} &= \text{advertising for $m$ months previously.}
\end{align*}

## Example: Boston marathon winning times

```{r, fig.height=2.6, echo=TRUE}
marathon <- boston_marathon |>
  filter(Event == "Men's open division") |>
  select(-Event) |>
  mutate(Minutes = as.numeric(Time) / 60)
marathon |> autoplot(Minutes) + labs(y = "Winning times in minutes")
```

## Example: Boston marathon winning times

\fontsize{9}{9}\sf

```{r, echo=TRUE}
fit_trends <- marathon |>
  model(
    # Linear trend
    linear = TSLM(Minutes ~ trend()),
    # Exponential trend
    exponential = TSLM(log(Minutes) ~ trend()),
    # Piecewise linear trend
    piecewise = TSLM(Minutes ~ trend(knots = c(1940, 1980)))
  )
```

```{r}
fit_trends
```

## Example: Boston marathon winning times

```{r, echo=TRUE, eval=FALSE, fig.height=2.7}
fit_trends |>
  forecast(h = 10) |>
  autoplot(marathon)
```

```{r, echo=FALSE, message=TRUE, warning=FALSE, fig.height=3}
fc_trends <- fit_trends |> forecast(h = 10)
marathon |>
  autoplot(Minutes) +
  geom_line(
    data = fitted(fit_trends),
    aes(y = .fitted, colour = .model)
  ) +
  autolayer(fc_trends, alpha = 0.5, level = 95) +
  labs(
    y = "Minutes",
    title = "Boston marathon winning times"
  )
```

## Example: Boston marathon winning times

```{r residPiecewise, message=FALSE, warning=FALSE}
fit_trends |>
  select(piecewise) |>
  gg_tsresiduals()
```
