---
title: "ETC3550/ETC5550 Applied&nbsp;forecasting"
author: "Ch7. Regression models"
institute: "OTexts.org/fpp3/"
pdf-engine: pdflatex
fig-width: 7.5
fig-height: 3.5
format:
  beamer:
    theme: monash
    aspectratio: 169
    fontsize: 14pt
    section-titles: false
    knitr:
      opts_chunk:
        dev: "cairo_pdf"
include-in-header: header.tex
execute:
  echo: false
  message: false
  warning: false
---

```{r setup, include=FALSE}
source("setup.R")
```

## Multiple regression and forecasting

\vspace*{0.2cm}\begin{block}{}\vspace*{-0.3cm}
$$
  y_t = \beta_0 + \beta_1 x_{1,t} + \beta_2 x_{2,t} + \cdots + \beta_kx_{k,t} + \varepsilon_t.
$$
\end{block}

* $y_t$ is the variable we want to predict: the "response" variable
* Each $x_{j,t}$ is numerical and is called a "predictor".
 They are usually assumed to be known for all past and future times.
* The coefficients $\beta_1,\dots,\beta_k$ measure the effect of each
predictor *after taking account of the effect of all other predictors
in the model*.
* $\varepsilon_t$ is a white noise error term

## Trend

**Linear trend**

\centerline{$x_t = t,\qquad t = 1,2,\dots,$}\pause

**Piecewise linear trend with bend at $\tau$**
\vspace*{-0.6cm}
\begin{align*}
x_{1,t} &= t \\
x_{2,t} &= \left\{ \begin{array}{ll}
  0 & t <\tau\\
  (t-\tau) & t \ge \tau
\end{array}\right.
\end{align*}
\pause\vspace*{-0.8cm}

**Quadratic or higher order trend**

\centerline{$x_{1,t} =t,\quad x_{2,t}=t^2,\quad \dots$}

\pause\vspace*{-0.1cm}
\centerline{\textcolor{orange}{\textbf{NOT RECOMMENDED!}}}

## Uses of dummy variables
\fontsize{13}{14}\sf

**Seasonal dummies**

* For quarterly data: use 3 dummies
* For monthly data: use 11 dummies
* For daily data: use 6 dummies
* What to do with weekly data?

\pause

**Outliers**

* A dummy variable can remove its effect.

\pause

**Public holidays**

* For daily data: if it is a public holiday, dummy=1, otherwise dummy=0.

## Holidays

**For monthly data**

* Christmas: always in December so part of monthly seasonal effect
* Easter: use a dummy variable $v_t=1$ if any part of Easter is in that month, $v_t=0$ otherwise.
* Ramadan and Chinese New Year similar.

## Fourier series

Periodic seasonality can be handled using pairs of Fourier \rlap{terms:}\vspace*{-0.3cm}
$$
s_{k}(t) = \sin\left(\frac{2\pi k t}{m}\right)\qquad c_{k}(t) = \cos\left(\frac{2\pi k t}{m}\right)
$$
$$
y_t = a + bt + \sum_{k=1}^K \left[\alpha_k s_k(t) + \beta_k c_k(t)\right] + \varepsilon_t$$\vspace*{-0.8cm}

* Every periodic function can be approximated by sums of sin and cos terms for large enough $K$.
* Choose $K$ by minimizing AICc or CV.
* Called "harmonic regression"

## Distributed lags

Lagged values of a predictor.

Example: $x$ is advertising which has a delayed effect

\vspace*{-0.8cm}\begin{align*}
  x_{1} &= \text{advertising for previous month;} \\
  x_{2} &= \text{advertising for two months previously;} \\
        & \vdots \\
  x_{m} &= \text{advertising for $m$ months previously.}
\end{align*}

## Comparing regression models
\fontsize{13}{14}\sf

* $R^2$  does not allow for ``degrees of freedom''.
* Adding *any* variable tends to increase the value of $R^2$, even if that variable is irrelevant.
\pause

To overcome this problem, we can use *adjusted $R^2$*:
\begin{block}{}
$$
\bar{R}^2 = 1-(1-R^2)\frac{T-1}{T-k-1}
$$
where $k=$ no.\ predictors and $T=$ no.\ observations.
\end{block}

\pause

\begin{alertblock}{Maximizing $\bar{R}^2$ is equivalent to minimizing $\hat\sigma^2$.}
\centerline{$\displaystyle
\hat{\sigma}^2 = \frac{1}{T-k-1}\sum_{t=1}^T \varepsilon_t^2$
}
\end{alertblock}

## Akaike's Information Criterion

\vspace*{0.2cm}\begin{block}{}
\centerline{$\text{AIC} = -2\log(L) + 2(k+2)$}
\end{block}\vspace*{-0.5cm}

* $L=$ likelihood
* $k=$ \# predictors in model.
* AIC penalizes terms more heavily than $\bar{R}^2$.

\pause\begin{block}{}
\centerline{$\text{AIC}_{\text{C}} = \text{AIC} + \frac{2(k+2)(k+3)}{T-k-3}$}
\end{block}

* Minimizing the AIC or AICc is asymptotically equivalent to minimizing MSE via **leave-one-out cross-validation** (for any linear regression).

## Leave-one-out cross-validation

For regression, leave-one-out cross-validation is faster and more efficient than time-series cross-validation.

* Select one observation for test set, and use *remaining* observations in training set. Compute error on test observation.
* Repeat using each possible observation as the test set.
* Compute accuracy measure over all errors.

```{r tscvplots, echo=FALSE}
tscv_plot <- function(.init, .step, h = 1) {
  expand.grid(
    time = seq(26),
    .id = seq(trunc(20 / .step))
  ) |>
    group_by(.id) |>
    mutate(
      observation = case_when(
        time <= ((.id - 1) * .step + .init) ~ "train",
        time %in% c((.id - 1) * .step + .init + h) ~ "test",
        TRUE ~ "unused"
      )
    ) |>
    ungroup() |>
    filter(.id <= 26 - .init) |>
    ggplot(aes(x = time, y = .id)) +
    geom_segment(
      aes(x = 0, xend = 27, y = .id, yend = .id),
      arrow = arrow(length = unit(0.015, "npc")),
      col = "black", size = .25
    ) +
    geom_point(aes(col = observation), size = 2) +
    scale_y_reverse() +
    scale_color_manual(values = c(train = "#0072B2", test = "#D55E00", unused = "gray")) +
    # theme_void() +
    # geom_label(aes(x = 28.5, y = 1, label = "time")) +
    guides(col = FALSE) +
    labs(x = "time", y = "") +
    theme_void() +
    theme(axis.title = element_text())
}
loocv_plot <- function() {
  expand.grid(time = seq(26), .id = seq(26)) |>
    group_by(.id) |>
    mutate(observation = if_else(time == .id, "test", "train")) |>
    ungroup() |>
    filter(.id <= 20) |>
    ggplot(aes(x = time, y = .id)) +
    geom_segment(
      aes(x = 0, xend = 27, y = .id, yend = .id),
      arrow = arrow(length = unit(0.015, "npc")),
      col = "black", size = .25
    ) +
    geom_point(aes(col = observation), size = 2) +
    scale_y_reverse() +
    scale_color_manual(values = c(train = "#0072B2", test = "#D55E00", unused = "gray")) +
    guides(col = FALSE) +
    labs(x = "time", y = "") +
    theme_void() +
    theme(axis.title = element_text())
}
```

## Cross-validation {-}

**Traditional evaluation**

```{r traintest1, fig.height=1, echo=FALSE, dependson="tscvplots"}
tscv_plot(.init = 18, .step = 20, h = 1:8) +
  geom_text(aes(x = 10, y = 0.8, label = "Training data"), color = "#0072B2") +
  geom_text(aes(x = 21, y = 0.8, label = "Test data"), color = "#D55E00") +
  ylim(1, 0)
```

\pause

**Time series cross-validation**

```{r tscvggplot1, echo=FALSE}
tscv_plot(.init = 3, .step = 1, h = 1) +
  geom_text(aes(x = 21, y = 0, label = "h = 1"), color = "#D55E00")
```

## Cross-validation {-}

**Traditional evaluation**

```{r traintest1a, fig.height=1, echo=FALSE, dependson="tscvplots"}
tscv_plot(.init = 18, .step = 20, h = 1:8) +
  geom_text(aes(x = 10, y = 0.8, label = "Training data"), color = "#0072B2") +
  geom_text(aes(x = 21, y = 0.8, label = "Test data"), color = "#D55E00") +
  ylim(1, 0)
```

**Leave-one-out cross-validation**

```{r, echo=FALSE}
loocv_plot() +
  geom_text(aes(x = 21, y = 0, label = "h = 1"), color = "#ffffff")
```

\only<2>{\begin{textblock}{4}(6,6)\begin{block}{}\fontsize{13}{15}\sf
CV = MSE on \textcolor[HTML]{D55E00}{test sets}\end{block}\end{textblock}}

## Bayesian Information Criterion

\begin{block}{}
$$
\text{BIC} = -2\log(L) + (k+2)\log(T)
$$
\end{block}
where $L$ is the likelihood and $k$ is the number of predictors in the model.\pause

* BIC penalizes terms more heavily than AIC
* Also called SBIC and SC.
* Minimizing BIC is asymptotically equivalent to leave-$v$-out cross-validation when $v = T[1-1/(log(T)-1)]$.

## Choosing regression variables
\fontsize{14}{15}\sf

**Best subsets regression**

* Fit all possible regression models using one or more of the predictors.
* Choose the best model based on one of the measures of predictive ability (CV, AIC, AICc).
\pause

**Backwards stepwise regression**

* Start with a model containing all variables.
* Subtract one variable at a time. Keep model if lower CV.
* Iterate until no further improvement.
* Not guaranteed to lead to best model.

## Ex-ante versus ex-post forecasts

 * *Ex ante forecasts* are made using only information available in advance.
    - require forecasts of predictors
 * *Ex post forecasts* are made using later information on the predictors.
    - useful for studying behaviour of forecasting models.

 * trend, seasonal and calendar variables are all known in advance, so these don't need to be forecast.
