---
title: "ETC3550/ETC5550 Applied&nbsp;forecasting"
author: "Ch7. Regression models"
date: "OTexts.org/fpp3/"
classoption: aspectratio=169
titlepage: title16x9.png
titlecolor: burntorange
toc: true
output:
  binb::monash:
    colortheme: monashwhite
    fig_width: 7.5
    fig_height: 3.5
    keep_tex: no
    includes:
      in_header: header.tex
---

```{r setup, include=FALSE}
source("setup.R")
```

# The linear model with time series

## Multiple regression and forecasting

\fontsize{13}{15}\sf

\begin{block}{}\vspace*{-0.3cm}
\[
  y_t = \beta_0 + \beta_1 x_{1,t} + \beta_2 x_{2,t} + \cdots + \beta_kx_{k,t} + \varepsilon_t.
\]
\end{block}

* $y_t$ is the variable we want to predict: the "response" variable
* Each $x_{j,t}$ is numerical and is called a "predictor".
 They are usually assumed to be known for all past and future times.
* The coefficients $\beta_1,\dots,\beta_k$ measure the effect of each
predictor after taking account of the effect of all other predictors
in the model.

That is, the coefficients measure the **marginal effects**.

* $\varepsilon_t$ is a white noise error term


## Example: US consumption expenditure
\fontsize{10}{11}\sf

```{r ConsInc, echo=TRUE, fig.height=3.5}
us_change %>%
  pivot_longer(c(Consumption, Income), names_to="Series") %>%
  autoplot(value) +
  labs(y="% change")
```

## Example: US consumption expenditure
\fontsize{9}{10}\sf

```{r ConsInc2, echo=TRUE}
us_change %>%
  ggplot(aes(x = Income, y = Consumption)) +
    labs(y = "Consumption (quarterly % change)",
         x = "Income (quarterly % change)") +
    geom_point() + geom_smooth(method = "lm", se = FALSE)
```

## Example: US consumption expenditure

\fontsize{9}{9}\sf

```{r, echo=TRUE}
fit_cons <- us_change %>%
  model(lm = TSLM(Consumption ~ Income))
report(fit_cons)
```

## Example: US consumption expenditure

```{r MultiPredictors, echo=FALSE, fig.height=5}
us_change %>%
  gather("Measure", "Change", Consumption, Income, Production, Savings, Unemployment) %>%
  ggplot(aes(x = Quarter, y = Change, colour = Measure)) +
  geom_line() +
  facet_grid(vars(Measure), scales = "free_y") +
  labs(y="") +
  guides(colour="none")
```

## Example: US consumption expenditure

```{r ScatterMatrix, echo=FALSE, fig.height=5, fig.width=8}
us_change %>%
  as_tibble() %>%
  select(-Quarter) %>%
  GGally::ggpairs()
```

## Example: US consumption expenditure

\fontsize{7.4}{7.4}\sf

```{r usestim, echo=TRUE}
fit_consMR <- us_change %>%
  model(lm = TSLM(Consumption ~ Income + Production + Unemployment + Savings))
report(fit_consMR)
```

## Example: US consumption expenditure

```{r usfitted1, echo=FALSE}
augment(fit_consMR) %>%
  ggplot(aes(x = Quarter)) +
  geom_line(aes(y = Consumption, colour = "Data")) +
  geom_line(aes(y = .fitted, colour = "Fitted")) +
  labs(y = NULL,
    title = "Percent change in US consumption expenditure"
  ) +
  scale_colour_manual(values=c(Data="black",Fitted="#D55E00")) +
  guides(colour = guide_legend(title = NULL))
```

## Example: US consumption expenditure

```{r usfitted2, echo=FALSE, message=FALSE, warning=FALSE}
augment(fit_consMR) %>%
  ggplot(aes(x=.fitted, y=Consumption)) +
    geom_point() +
    labs(y="Fitted (predicted values)",
         x="Data (actual values)",
         title ="Percentage change in US consumption expenditure") +
    geom_abline(intercept=0, slope=1)
```

## Example: US consumption expenditure

```{r, echo=TRUE}
fit_consMR %>% gg_tsresiduals()
```

# Some useful predictors for linear models

## Trend

**Linear trend**
\[
  x_t = t
\]

* $t=1,2,\dots,T$
* Strong assumption that trend will continue.

## Dummy variables

\begin{textblock}{6}(0.4,1.5)
If a categorical variable takes only two values (e.g., `Yes'
or `No'), then an equivalent numerical variable can be constructed
taking value 1 if yes and 0 if no. This is called a \textbf{dummy variable}.
\end{textblock}

\placefig{7.7}{1.}{width=5cm}{dummy2}

## Dummy variables

\begin{textblock}{5}(0.4,1.5)
If there are more than two categories, then the variable can
be coded using several dummy variables (one fewer than the total
number of categories).

\end{textblock}

\placefig{5.5}{1.5}{width=7.3cm}{dummy3}

## Beware of the dummy variable trap!
* Using one dummy for each category gives too many dummy variables!

* The regression will then be singular and inestimable.

* Either omit the constant, or omit the dummy for one category.

* The coefficients of the dummies are relative to the omitted category.

## Uses of dummy variables

\fontsize{13}{15}\sf

**Seasonal dummies**

* For quarterly data: use 3 dummies
* For monthly data: use 11 dummies
* For daily data: use 6 dummies
* What to do with weekly data?

\pause

**Outliers**

* If there is an outlier, you can use a dummy variable to remove its effect.

\pause

**Public holidays**

* For daily data: if it is a public holiday, dummy=1, otherwise dummy=0.

## Beer production revisited

```{r, echo=FALSE}
recent_production <- aus_production %>% filter(year(Quarter) >= 1992)
recent_production %>% autoplot(Beer) +
  labs(y="Megalitres",title ="Australian quarterly beer production")
```

\pause

\begin{block}{Regression model}
\centering
$y_t = \beta_0 + \beta_1 t + \beta_2d_{2,t} + \beta_3 d_{3,t} + \beta_4 d_{4,t} + \varepsilon_t$
\end{block}

* $d_{i,t} = 1$ if $t$ is quarter $i$ and 0 otherwise.

## Beer production revisited
\fontsize{8}{8}\sf

```{r, echo=TRUE}
fit_beer <- recent_production %>% model(TSLM(Beer ~ trend() + season()))
report(fit_beer)
```

## Beer production revisited
\fontsize{8.5}{9}\sf

```{r, fig.height=3.3}
augment(fit_beer) %>%
  ggplot(aes(x = Quarter)) +
  geom_line(aes(y = Beer, colour = "Data")) +
  geom_line(aes(y = .fitted, colour = "Fitted")) +
  labs(y="Megalitres",title ="Australian quarterly beer production") +
  scale_colour_manual(values = c(Data = "black", Fitted = "#D55E00"))
```

## Beer production revisited

```{r, echo=FALSE}
augment(fit_beer) %>%
  ggplot(aes(x=Beer, y=.fitted, colour=factor(quarter(Quarter)))) +
    geom_point() +
    labs(y="Fitted", x="Actual values",
         title = "Quarterly beer production") +
    scale_colour_brewer(palette="Dark2", name="Quarter") +
    geom_abline(intercept=0, slope=1)
```

## Beer production revisited
\fontsize{9}{9}\sf

```{r, echo=TRUE}
fit_beer %>% gg_tsresiduals()
```

## Beer production revisited
\fontsize{9}{9}\sf

```{r, echo=TRUE}
fit_beer %>% forecast %>% autoplot(recent_production)
```

## Fourier series

Periodic seasonality can be handled using pairs of Fourier terms:
$$
s_{k}(t) = \sin\left(\frac{2\pi k t}{m}\right)\qquad c_{k}(t) = \cos\left(\frac{2\pi k t}{m}\right)
$$
$$
y_t = a + bt + \sum_{k=1}^K \left[\alpha_k s_k(t) + \beta_k c_k(t)\right] + \varepsilon_t$$

* Every periodic function can be approximated by sums of sin and cos terms for large enough $K$.
* Choose $K$ by minimizing AICc.
* Called "harmonic regression"

```r
TSLM(y ~ trend() + fourier(K))
```

## Harmonic regression: beer production

\fontsize{8}{8}\sf

```{r fourierbeer, echo=TRUE}
fourier_beer <- recent_production %>% model(TSLM(Beer ~ trend() + fourier(K=2)))
report(fourier_beer)
```

## Harmonic regression: eating-out expenditure

\fontsize{9}{10}\sf

```{r cafe, echo=TRUE, fig.height=3.8, fig.width=8}
aus_cafe <- aus_retail %>% filter(
    Industry == "Cafes, restaurants and takeaway food services",
    year(Month) %in% 2004:2018
  ) %>% summarise(Turnover = sum(Turnover))
aus_cafe %>% autoplot(Turnover)
```

## Harmonic regression: eating-out expenditure

\fontsize{9}{10}\sf

```{r cafefit, dependson='cafe', fig.height=5, echo=TRUE}
fit <- aus_cafe %>% 
  model(K1 = TSLM(log(Turnover) ~ trend() + fourier(K = 1)),
        K2 = TSLM(log(Turnover) ~ trend() + fourier(K = 2)),
        K3 = TSLM(log(Turnover) ~ trend() + fourier(K = 3)),
        K4 = TSLM(log(Turnover) ~ trend() + fourier(K = 4)),
        K5 = TSLM(log(Turnover) ~ trend() + fourier(K = 5)),
        K6 = TSLM(log(Turnover) ~ trend() + fourier(K = 6)))
glance(fit) %>% select(.model, r_squared, adj_r_squared, AICc)
```

## Harmonic regression: eating-out expenditure

```{r, echo=FALSE}
cafe_plot <- function(...){
  fit %>%
    select(...) %>%
    forecast() %>% autoplot(aus_cafe) +
    labs(title = sprintf("Log transformed %s, trend() + fourier(K = %s)", model_sum(select(fit,...)[[1]][[1]]), deparse(..1))) +
    geom_label(
      aes(x = yearmonth("2007 Jan"), y = 4250, label = paste0("AICc = ", format(AICc))),
      data = glance(select(fit,...))
    ) +
    geom_line(aes(y = .fitted), colour = "red", augment(select(fit, ...))) +
    ylim(c(1500, 5100))
}
```

```{r cafe1, dependson='cafe', fig.height=5, echo=FALSE}
cafe_plot(K = 1)
```

## Harmonic regression: eating-out expenditure

```{r cafe2, dependson='cafe', fig.height=5, echo=FALSE}
cafe_plot(K = 2)
```

## Harmonic regression: eating-out expenditure

```{r cafe3, dependson='cafe', fig.height=5, echo=FALSE}
cafe_plot(K = 3)
```

## Harmonic regression: eating-out expenditure

```{r cafe4, dependson='cafe', fig.height=5, echo=FALSE}
cafe_plot(K = 4)
```

## Harmonic regression: eating-out expenditure

```{r cafe5, dependson='cafe', fig.height=5, echo=FALSE}
cafe_plot(K = 5)
```

## Harmonic regression: eating-out expenditure

```{r cafe6, dependson='cafe', fig.height=5, echo=FALSE}
cafe_plot(K = 6)
```

## Intervention variables

**Spikes**

* Equivalent to a dummy variable for handling an outlier.
\pause

**Steps**

* Variable takes value 0 before the intervention and 1 afterwards.
\pause

**Change of slope**

* Variables take values 0 before the intervention and values $\{1,2,3,\dots\}$ afterwards.

## Holidays

**For monthly data**

* Christmas: always in December so part of monthly seasonal effect
* Easter: use a dummy variable $v_t=1$ if any part of Easter is in that month, $v_t=0$ otherwise.
* Ramadan and Chinese new year similar.

## Trading days

With monthly data, if the observations vary depending on how many different types of days in the month, then trading day predictors can be useful.

\begin{align*}
z_1 &= \text{\# Mondays in month;} \\
z_2 &= \text{\# Tuesdays in month;} \\
&\vdots \\
z_7 &= \text{\# Sundays in month.}
\end{align*}

## Distributed lags

Lagged values of a predictor.

Example: $x$ is advertising which has a delayed effect

\begin{align*}
  x_{1} &= \text{advertising for previous month;} \\
  x_{2} &= \text{advertising for two months previously;} \\
        & \vdots \\
  x_{m} &= \text{advertising for $m$ months previously.}
\end{align*}

## Nonlinear trend

**Piecewise linear trend with bend at $\tau$**
\begin{align*}
x_{1,t} &= t \\
x_{2,t} &= \left\{ \begin{array}{ll}
  0 & t <\tau\\
  (t-\tau) & t \ge \tau
\end{array}\right.
\end{align*}
\pause\vspace*{1cm}

**Quadratic or higher order trend**
\[
  x_{1,t} =t,\quad x_{2,t}=t^2,\quad \dots
\]
\pause\vspace*{-0.5cm}

\centerline{\textcolor{orange}{\textbf{NOT RECOMMENDED!}}}

## Example: Boston marathon winning times

\fontsize{9}{9}\sf

```{r, fig.height=3.5, echo=TRUE}
marathon <- boston_marathon %>%
  filter(Event == "Men's open division") %>%
  select(-Event) %>%
  mutate(Minutes = as.numeric(Time)/60)
marathon %>% autoplot(Minutes) +
  labs(y="Winning times in minutes")
```

## Example: Boston marathon winning times

\fontsize{9}{9}\sf

```{r, echo=TRUE}
fit_trends <- marathon %>%
  model(
    # Linear trend
    linear = TSLM(Minutes ~ trend()),
    # Exponential trend
    exponential = TSLM(log(Minutes) ~ trend()),
    # Piecewise linear trend
    piecewise = TSLM(Minutes ~ trend(knots = c(1940, 1980)))
  )
```

```{r}
fit_trends
```

## Example: Boston marathon winning times

\fontsize{10}{10}\sf

```{r, echo=TRUE, eval=FALSE}
fit_trends %>% forecast(h=10) %>% autoplot(marathon)
```

```{r, echo=FALSE, message=TRUE, warning=FALSE}
fc_trends <- fit_trends %>% forecast(h = 10)
marathon %>%
  autoplot(Minutes) +
  geom_line(data = fitted(fit_trends),
            aes(y = .fitted, colour = .model)) +
  autolayer(fc_trends, alpha = 0.5, level = 95) +
  labs(y = "Minutes",
       title = "Boston marathon winning times")
```

## Example: Boston marathon winning times
\fontsize{9}{9}\sf

```{r residPiecewise, message=FALSE, warning=FALSE}
fit_trends %>% select(piecewise) %>%
  gg_tsresiduals()
```

# Residual diagnostics

## Multiple regression and forecasting
For forecasting purposes, we require the following assumptions:

* $\varepsilon_t$ are uncorrelated and zero mean

* $\varepsilon_t$ are uncorrelated with each $x_{j,t}$.
\pause

It is **useful** to also have $\varepsilon_t \sim \text{N}(0,\sigma^2)$ when producing prediction intervals or doing statistical tests.

## Residual plots

Useful for spotting outliers and whether the linear model was
appropriate.

* Scatterplot of residuals $\varepsilon_t$ against each predictor $x_{j,t}$.

* Scatterplot residuals against the fitted values $\hat y_t$

* Expect to see scatterplots resembling a horizontal band with
no values too far from the band and no patterns such as curvature or
increasing spread.

## Residual patterns

* If a plot of the residuals vs any predictor in the model shows a pattern, then the relationship is nonlinear.

* If a plot of the residuals vs any predictor **not** in the model shows a pattern, then the predictor should be added to the model.

* If a plot of the residuals vs fitted values shows a pattern, then there is heteroscedasticity in the errors. (Could try a transformation.)

# Selecting predictors and forecast evaluation

## Comparing regression models

Computer output for regression will always give the $R^2$ value. This is a
useful summary of the model.

* It is equal to the square of the correlation between $y$ and $\hat y$.
* It is often called the ``coefficient of determination''.
* It can also be calculated as follows:
$$R^2 = \frac{\sum(\hat{y}_t - \bar{y})^2}{\sum(y_t-\bar{y})^2}
$$
* It is the proportion of variance accounted for (explained) by the predictors.

## Comparing regression models
\fontsize{13}{15}\sf

However \dots

* $R^2$  does not allow for ``degrees of freedom''.

* Adding *any* variable tends to increase the value of $R^2$, even if that variable is irrelevant.
\pause

To overcome this problem, we can use *adjusted $R^2$*:

\begin{block}{}
\[
\bar{R}^2 = 1-(1-R^2)\frac{T-1}{T-k-1}
\]
where $k=$ no.\ predictors and $T=$ no.\ observations.
\end{block}

\pause

\begin{alertblock}{Maximizing $\bar{R}^2$ is equivalent to minimizing $\hat\sigma^2$.}
\centerline{$\displaystyle
\hat{\sigma}^2 = \frac{1}{T-k-1}\sum_{t=1}^T \varepsilon_t^2$
}
\end{alertblock}

## Akaike's Information Criterion

\begin{block}{}
$$\text{AIC} = -2\log(L) + 2(k+2)$$
\end{block}
where $L$ is the likelihood and $k$ is the number of predictors in the model.\pause

* AIC penalizes terms more heavily than $\bar{R}^2$.
* Minimizing the AIC is asymptotically equivalent to minimizing MSE via **leave-one-out cross-validation** (for any linear regression).

## Corrected AIC

For small values of $T$, the AIC tends to select too many predictors, and so a bias-corrected version of the AIC has been developed.

\begin{block}{}
\[
\text{AIC}_{\text{C}} = \text{AIC} + \frac{2(k+2)(k+3)}{T-k-3}
\]
\end{block}

As with the AIC, the AIC$_{\text{C}}$ should be minimized.

## Bayesian Information Criterion

\begin{block}{}
\[
\text{BIC} = -2\log(L) + (k+2)\log(T)
\]
\end{block}
where $L$ is the likelihood and $k$ is the number of predictors in the model.\pause

* BIC penalizes terms more heavily than AIC

* Also called SBIC and SC.

* Minimizing BIC is asymptotically equivalent to leave-$v$-out cross-validation when $v = T[1-1/(log(T)-1)]$.

## Leave-one-out cross-validation

For regression, leave-one-out cross-validation is faster and more efficient than time-series cross-validation.

* Select one observation for test set, and use *remaining* observations in training set. Compute error on test observation.
* Repeat using each possible observation as the test set.
* Compute accuracy measure over all errors.

```{r tscvplots, echo=FALSE}
tscv_plot <- function(.init, .step, h = 1) {
  expand.grid(
    time = seq(26),
    .id = seq(trunc(20 / .step))
  ) %>%
    group_by(.id) %>%
    mutate(
      observation = case_when(
        time <= ((.id - 1) * .step + .init) ~ "train",
        time %in% c((.id - 1) * .step + .init + h) ~ "test",
        TRUE ~ "unused"
      )
    ) %>%
    ungroup() %>%
    filter(.id <= 26 - .init) %>%
    ggplot(aes(x = time, y = .id)) +
    geom_segment(
      aes(x = 0, xend = 27, y = .id, yend = .id),
      arrow = arrow(length = unit(0.015, "npc")),
      col = "black", size = .25
    ) +
    geom_point(aes(col = observation), size = 2) +
    scale_y_reverse() +
    scale_color_manual(values = c(train = "#0072B2", test = "#D55E00", unused = "gray")) +
    #theme_void() +
    #geom_label(aes(x = 28.5, y = 1, label = "time")) +
    guides(col = FALSE) +
    labs(x="time", y="") +
    theme_void() +
    theme(axis.title = element_text())
}
loocv_plot <- function() {
  expand.grid(time = seq(26), .id = seq(26)) %>%
    group_by(.id) %>%
    mutate(observation = if_else(time == .id, "test", "train")) %>%
    ungroup() %>%
    filter(.id <= 20) %>%
    ggplot(aes(x = time, y = .id)) +
    geom_segment(
      aes(x = 0, xend = 27, y = .id, yend = .id),
      arrow = arrow(length = unit(0.015, "npc")),
      col = "black", size = .25
    ) +
    geom_point(aes(col = observation), size = 2) +
    scale_y_reverse() +
    scale_color_manual(values = c(train = "#0072B2", test = "#D55E00", unused = "gray")) +
    guides(col = FALSE) +
    labs(x="time", y="") +
    theme_void() +
    theme(axis.title = element_text())
}
```

## Cross-validation {-}

**Traditional evaluation**

```{r traintest1, fig.height=1, echo=FALSE, dependson="tscvplots"}
tscv_plot(.init = 18, .step = 20, h = 1:8) +
  geom_text(aes(x=10,y=0.8,label="Training data"), color = "#0072B2") +
  geom_text(aes(x=21,y=0.8,label="Test data"), color = "#D55E00") +
  ylim(1,0)
```

\pause

**Time series cross-validation**

```{r tscvggplot1, echo=FALSE}
tscv_plot(.init = 3, .step = 1, h = 1) +
  geom_text(aes(x=21,y=0,label="h = 1"), color = "#D55E00")
```

## Cross-validation {-}

**Traditional evaluation**

```{r traintest1a, fig.height=1, echo=FALSE, dependson="tscvplots"}
tscv_plot(.init = 18, .step = 20, h = 1:8) +
  geom_text(aes(x=10,y=0.8,label="Training data"), color = "#0072B2") +
  geom_text(aes(x=21,y=0.8,label="Test data"), color = "#D55E00") +
  ylim(1,0)
```

**Leave-one-out cross-validation**

```{r, echo=FALSE}
loocv_plot() +
  geom_text(aes(x=21,y=0,label="h = 1"), color = "#ffffff")
```

\only<2>{\begin{textblock}{4}(6,6)\begin{block}{}\fontsize{13}{15}\sf
CV = MSE on \textcolor[HTML]{D55E00}{test sets}\end{block}\end{textblock}}

## Comparing regression models
\fontsize{11}{14}\sf

```{r, echo=TRUE}
glance(fit_trends) %>%
  select(.model, r_squared, adj_r_squared, AICc, CV)
```

\fontsize{13}{14}\sf

 * Be careful making comparisons when transformations are used.

## Choosing regression variables

**Best subsets regression**

* Fit all possible regression models using one or more of the predictors.

* Choose the best model based on one of the measures of predictive ability (CV, AIC, AICc).
\pause

**Warning!**

* If there are a large number of predictors, this is not possible.

* For example, 44 predictors leads to 18 trillion possible models!

## Choosing regression variables

**Backwards stepwise regression**

* Start with a model containing all variables.
* Try subtracting one variable at a time. Keep the model if it
has lower CV or AICc.
* Iterate until no further improvement.
\pause

**Notes**

* Stepwise regression is not guaranteed to lead to the best possible
model.
* Inference on coefficients of final model will be wrong.

# Forecasting with regression

## Ex-ante versus ex-post forecasts

 * *Ex ante forecasts* are made using only information available in advance.
    - require forecasts of predictors
 * *Ex post forecasts* are made using later information on the predictors.
    - useful for studying behaviour of forecasting models.

 * trend, seasonal and calendar variables are all known in advance, so these don't need to be forecast.

## Scenario based forecasting

 * Assumes possible scenarios for the predictor variables
 * Prediction intervals for scenario based forecasts do not include the uncertainty associated with the future values of the predictor variables.

## Building a predictive regression model {-}

 * If getting forecasts of predictors is difficult, you can use lagged predictors instead.
$$y_{t}=\beta_0+\beta_1x_{1,t-h}+\dots+\beta_kx_{k,t-h}+\varepsilon_{t}$$

 * A different model for each forecast horizon $h$.

## US Consumption
\fontsize{10}{11}\sf

```{r usconsumptionf, echo=TRUE}
fit_consBest <- us_change %>%
  model(
    TSLM(Consumption ~ Income + Savings + Unemployment)
  )

future_scenarios <- scenarios(
  Increase = new_data(us_change, 4) %>%
    mutate(Income=1, Savings=0.5, Unemployment=0),
  Decrease = new_data(us_change, 4) %>%
    mutate(Income=-1, Savings=-0.5, Unemployment=0),
  names_to = "Scenario")

fc <- forecast(fit_consBest, new_data = future_scenarios)
```

## US Consumption
\fontsize{10}{10}\sf

```{r usconsumptionf2, echo=TRUE, fig.height=3.5}
us_change %>% autoplot(Consumption) +
  labs(y="% change in US consumption") +
  autolayer(fc) +
  labs(title = "US consumption", y = "% change")
```

# Matrix formulation

## Matrix formulation

\begin{block}{}
\centerline{$y_t = \beta_0 + \beta_1 x_{1,t} + \beta_2 x_{2,t} + \cdots + \beta_kx_{k,t} + \varepsilon_t.$}
\end{block}
\pause

Let $\bm{y} = (y_1,\dots,y_T)'$, $\bm{\varepsilon} = (\varepsilon_1,\dots,\varepsilon_T)'$, $\bm{\beta} = (\beta_0,\beta_1,\dots,\beta_k)'$ and
\[
\bm{X} = \begin{bmatrix}
1      & x_{1,1} & x_{2,1} & \dots & x_{k,1}\\
1      & x_{1,2} & x_{2,2} & \dots & x_{k,2}\\
\vdots & \vdots  & \vdots  &       & \vdots\\
1      & x_{1,T} & x_{2,T} & \dots & x_{k,T}
  \end{bmatrix}.
\]\pause
Then

###
\centerline{$\bm{y} = \bm{X}\bm{\beta} + \bm{\varepsilon}.$}

## Matrix formulation

**Least squares estimation**

Minimize: $(\bm{y} - \bm{X}\bm{\beta})'(\bm{y} - \bm{X}\bm{\beta})$\pause

Differentiate wrt $\bm{\beta}$ gives

\begin{block}{}
\centerline{$\hat{\bm{\beta}} = (\bm{X}'\bm{X})^{-1}\bm{X}'\bm{y}$}
\end{block}

\pause
(The "normal equation".)\pause

\[
\hat{\sigma}^2 = \frac{1}{T-k-1}(\bm{y} - \bm{X}\hat{\bm{\beta}})'(\bm{y} - \bm{X}\hat{\bm{\beta}})
\]

\structure{Note:} If you fall for the dummy variable trap, $(\bm{X}'\bm{X})$ is a singular matrix.

## Likelihood

If the errors are iid and normally distributed, then
\[
\bm{y} \sim \text{N}(\bm{X}\bm{\beta},\sigma^2\bm{I}).
\]\pause
So the likelihood is
\[
L = \frac{1}{\sigma^T(2\pi)^{T/2}}\exp\left(-\frac1{2\sigma^2}(\bm{y}-\bm{X}\bm{\beta})'(\bm{y}-\bm{X}\bm{\beta})\right)
\]\pause
which is maximized when $(\bm{y}-\bm{X}\bm{\beta})'(\bm{y}-\bm{X}\bm{\beta})$ is minimized.\pause

\centerline{\alert{So \textbf{MLE = OLS}.}}

## Multiple regression forecasts

\begin{block}{Optimal forecasts}\vspace*{-0.2cm}
\[
\hat{y}^* =
\text{E}(y^* | \bm{y},\bm{X},\bm{x}^*) =
\bm{x}^*\hat{\bm{\beta}} = \bm{x}^*(\bm{X}'\bm{X})^{-1}\bm{X}'\bm{y}
\]
\end{block}
where $\bm{x}^*$ is a row vector containing the values of the predictors for the forecasts (in the same format as $\bm{X}$).

\pause

\begin{block}{Forecast variance}\vspace*{-0.2cm}
\[
\text{Var}(y^* | \bm{X},\bm{x}^*) = \sigma^2 \left[1 + \bm{x}^* (\bm{X}'\bm{X})^{-1} (\bm{x}^*)'\right]
\]
\end{block}
\pause

* This ignores any errors in $\bm{x}^*$.
* 95% prediction intervals assuming normal errors:
$$\hat{y}^* \pm 1.96 \sqrt{\text{Var}(y^*| \bm{X},\bm{x}^*)}.$$

<!-- ## Multiple regression forecasts

\begin{block}{Fitted values}
$$
  \hat{\bm{y}} = \bm{X}\hat{\bm{\beta}} = \bm{X}(\bm{X}'\bm{X})^{-1}\bm{X}'\bm{y} = \bm{H}\bm{y}
$$
\end{block}
where $\bm{H} = \bm{X}(\bm{X}'\bm{X})^{-1}\bm{X}'$ is the ``hat matrix''.\pause

**Leave-one-out residuals**

Let $h_1,\dots,h_T$ be the diagonal values of $\bm{H}$, then the cross-validation statistic is

\begin{block}{}
$$
\text{CV} = \frac1T\sum_{t=1}^T[e_t/(1-h_t)]^2,
$$
\end{block}

where $e_t$ is the residual obtained from fitting the model to all $T$ observations. -->

# Correlation, causation and forecasting

## Correlation is not causation
* When $x$ is useful for predicting $y$, it is not necessarily causing $y$.

* e.g., predict number of drownings $y$ using number of ice-creams sold $x$.

* Correlations are useful for forecasting, even when there is no causality.

* Better models usually involve causal relationships (e.g., temperature $x$ and people $z$ to predict drownings $y$).

## Multicollinearity
In regression analysis, multicollinearity occurs when:

*  Two  predictors are highly  correlated (i.e., the correlation between them is close to $\pm1$).
* A linear combination of some of the predictors is highly correlated  with another predictor.
*  A linear combination of one subset of predictors is highly correlated with a linear combination of another
  subset of predictors.

## Multicollinearity

If multicollinearity exists\dots

* the numerical estimates of coefficients may be wrong (worse in Excel than in a statistics package)
* don't rely on the $p$-values to determine significance.
* there is no problem with model *predictions* provided the predictors used for forecasting are within the range used for fitting.
* omitting variables can help.
* combining variables can help.
