---
title: "8. Exponential smoothing"
author: "8.1 Simple exponential smoothing"
date: "OTexts.org/fpp3/"
classoption: aspectratio=169
titlepage: fpp3title.png
titlecolor: fpp3red
toc: false
output:
  binb::monash:
    colortheme: monashwhite
    fig_width: 7.5
    fig_height: 3
    keep_tex: no
    includes:
      in_header: fpp3header.tex
---

```{r setup, include=FALSE}
source("setup.R")
library(gganimate)
library(purrr)
library(rlang)
library(magick)
```

## Simple methods

Time series $y_1,y_2,\dots,y_T$.

\begin{block}{Random walk forecasts}
  \centerline{$\pred{y}{T+h}{T} = y_T$}
\end{block}\pause

\begin{block}{Average forecasts}
  \centerline{$\displaystyle\pred{y}{T+h}{T} = \frac1T\sum_{t=1}^T y_t$}
\end{block}\pause\vspace*{-0.2cm}

* Want something in between these methods.
* Most recent data should have more weight.

<!-- * Simple exponential smoothing uses a weighted moving average with weights that decrease exponentially. -->

## Simple Exponential Smoothing

\begin{block}{Forecast equation}
\centerline{$\pred{y}{T+1}{T} = \alpha y_T + \alpha(1-\alpha) y_{T-1} + \alpha(1-\alpha)^2 y_{T-2}+ \cdots$,}\mbox{}\hfill where $0 \le \alpha \le1$.
\end{block}\pause

\small\centering\begin{tabular}{lllll}
\toprule
& \multicolumn{4}{l}{Weights assigned to observations for:}\\
Observation  &   $\alpha = 0.2$   &   $\alpha = 0.4$  &   $\alpha = 0.6$  & $\alpha = 0.8$ \\
\midrule
$y_{T}$   & 0.2            & 0.4            & 0.6            & 0.8\\
$y_{T-1}$ & 0.16           & 0.24           & 0.24           & 0.16\\
$y_{T-2}$ & 0.128          & 0.144          & 0.096          & 0.032\\
$y_{T-3}$ & 0.1024         & 0.0864         & 0.0384         & 0.0064\\
$y_{T-4}$ & $(0.2)(0.8)^4$ & $(0.4)(0.6)^4$ & $(0.6)(0.4)^4$ & $(0.8)(0.2)^4$\\
$y_{T-5}$ & $(0.2)(0.8)^5$ & $(0.4)(0.6)^5$ & $(0.6)(0.4)^5$ & $(0.8)(0.2)^5$\\
\bottomrule
\end{tabular}

## Simple Exponential Smoothing

```{r alpha-anim, cache=TRUE, echo=FALSE, fig.show='animate', interval=1/5, message=FALSE, fig.height=5, fig.width=8, aniopts='controls,buttonsize=0.3cm,width=11.5cm'}
algeria_economy <- global_economy |>
  filter(Country == "Algeria")
alpha_anim <- purrr::map_dfr(
  purrr::set_names(
    seq(0, 0.99, 0.01),
    seq(0, 0.99, 0.01)
  ),
  function(alpha) {
    algeria_economy |>
      model(ETS(
        Exports ~
          error("A") +
          trend("N", alpha = alpha, alpha_range = c(-1, 1), beta_range = c(-1, 1)) +
          season("N", gamma_range = c(-1, 1)),
        bounds = "admissible"
      )) |>
      augment() |>
      as_tibble()
  },
  .id = "alpha"
) |>
  mutate(alpha = as.numeric(alpha))
alpha_anim |>
  ggplot(aes(x = Year, y = Exports)) +
  geom_line() +
  geom_line(aes(y = .fitted), colour = "blue") +
  transition_manual(alpha) +
  labs(
    y = "% of GDP",
    title = "Algerian exports of goods and services: level (alpha = {format(as.numeric(as.character(current_frame)), nsmall=2)})"
  )
```

## Simple Exponential Smoothing

\begin{block}{Component form}\vspace*{-0.4cm}
\begin{align*}
\text{Forecast equation}&&\pred{y}{t+h}{t} &= \ell_{t}\\
\text{Smoothing equation}&&\ell_{t} &= \alpha y_{t} + (1 - \alpha)\ell_{t-1}
\end{align*}
\end{block}\vspace*{-0.2cm}

* $\ell_t$ is the level (or the smoothed value) of the series at time t.
* $\pred{y}{t+1}{t} = \alpha y_t + (1-\alpha) \pred{y}{t}{t-1}$

\pause

  Iterate to get exponentially weighted moving average form.

\begin{block}{Weighted average form}
$\displaystyle\pred{y}{T+1}{T}=\sum_{j=0}^{T-1} \alpha(1-\alpha)^j y_{T-j}+(1-\alpha)^T \ell_{0}$
\end{block}

## Optimising smoothing parameters

  * Need to choose best values for $\alpha$ and $\ell_0$.
  * Similarly to regression, choose optimal parameters by minimising SSE:
$$
  \text{SSE}=\sum_{t=1}^T(y_t - \pred{y}{t}{t-1})^2.
$$
  * Unlike regression there is no closed form solution --- use numerical optimization.

\pause

```{r ses, echo=FALSE}
fit <- algeria_economy |>
  model(
    ETS(Exports ~ error("A") + trend("N") + season("N"))
  )
```

  * For Algerian Exports example:
    - $\hat\alpha = `r sprintf("%4.4f",tidy(fit)[1,4])`$
    - $\hat\ell_0 = `r sprintf("%4.2f",tidy(fit)[2,4])`$

## Simple Exponential Smoothing

```{r alpha-static, fig.height=4, fig.width=8, echo=FALSE}
alpha_static <- map_dfr(list(0, as.numeric(tidy(fit)[1, 4]), 1), function(alpha) {
  fit <- algeria_economy |>
    model(ETS(Exports ~ error("A") + trend("N",
      alpha = alpha, alpha_range = c(-0.01, 1),
      beta_range = c(-1, 1)
    ) + season("N", gamma_range = c(-1, 1)), bounds = "admissible"))
  fit |>
    augment() |>
    mutate(alpha = tidy(fit)$estimate[tidy(fit)$term == "alpha"]) |>
    as_tibble()
}) |>
  mutate(alpha = factor(format(alpha)))
algeria_economy |>
  ggplot(aes(x = Year, y = Exports)) +
  geom_line() +
  geom_line(aes(y = .fitted, colour = alpha), data = alpha_static) +
  labs(
    y = "% of GDP",
    title = "Algerian exports of goods and services: level"
  )
```


## Example: Algerian Exports

\fontsize{9}{9}\sf

```{r ses-fit, echo=TRUE, cache=TRUE}
algeria_economy <- global_economy |>
  filter(Country == "Algeria")
fit <- algeria_economy |>
  model(ANN = ETS(Exports ~ error("A") + trend("N") + season("N")))
report(fit)
```

## Example: Algerian Exports

```{r ses-cmp0, echo = TRUE, fig.asp=0.7}
components(fit) |> autoplot()
```

## Example: Algerian Exports

\fontsize{9}{9}\sf

```{r ses-cmp, echo = TRUE}
components(fit) |>
  left_join(fitted(fit), by = c("Country", ".model", "Year"))
```

## Example: Algerian Exports

```{r ses-fc, echo=TRUE, cache=TRUE}
fit |>
  forecast(h = 5) |>
  autoplot(algeria_economy) +
  labs(y = "% of GDP", title = "Exports: Algeria")
```
