---
title: "8. Exponential smoothing"
author: "8.2 Methods with trend"
date: "OTexts.org/fpp3/"
classoption: aspectratio=169
titlepage: fpp3title.png
titlecolor: fpp3red
toc: false
output:
  binb::monash:
    colortheme: monashwhite
    fig_width: 7.5
    fig_height: 3
    keep_tex: no
    includes:
      in_header: fpp3header.tex
---

```{r setup, include=FALSE}
source("setup.R")
library(gganimate)
library(purrr)
library(rlang)
library(magick)
```

## Holt's linear trend

\begin{block}{Component form}\vspace*{-.9cm}
\begin{align*}
\text{Forecast }&& \pred{y}{t+h}{t} &= \ell_{t} + hb_{t} \\
\text{Level }&& \ell_{t} &= \alpha y_{t} + (1 - \alpha)(\ell_{t-1} + b_{t-1})\\
\text{Trend }&& b_{t} &= \beta^*(\ell_{t} - \ell_{t-1}) + (1 -\beta^*)b_{t-1},
\end{align*}
\end{block}
\pause\vspace*{-0.2cm}

  * Two smoothing parameters $\alpha$ and $\beta^*$ ($0\le\alpha,\beta^*\le1$).
  * $\ell_t$ level: weighted average between $y_t$ and one-step ahead forecast for time $t$, $(\ell_{t-1} + b_{t-1}=\pred{y}{t}{t-1})$
  * $b_t$ slope: weighted average of $(\ell_{t} - \ell_{t-1})$ and $b_{t-1}$, current and previous estimate of slope.
  * Choose $\alpha, \beta^*, \ell_0, b_0$ to minimise SSE.

## Exponential smoothing: trend/slope

```{r beta-anim, cache=TRUE, echo=FALSE, fig.show='animate', interval=1/5, message=FALSE, fig.height=5, fig.width=8, aniopts='controls,buttonsize=0.3cm,width=11.5cm', eval=TRUE}
aus_economy <- global_economy |>
  filter(Code == "AUS") |>
  mutate(Pop = Population / 1e6)
beta_anim <- map_dfr(
  set_names(
    seq(0, 0.99, by = 0.01),
    seq(0, 0.99, by = 0.01)
  ),
  function(beta) {
    aus_economy |>
      model(ETS(
        Pop ~
          error("A") +
          trend("A", alpha = 0.001, alpha_range = c(-1, 1), beta = beta) +
          season("N"),
        bounds = "admissible"
      )) |>
      augment() |>
      as_tibble()
  },
  .id = "beta"
) |>
  mutate(beta = as.numeric(beta))
beta_anim |>
  left_join(select(aus_economy, Year), by = "Year") |>
  ggplot(aes(x = Year, y = Pop)) +
  geom_line() +
  geom_line(aes(y = .fitted), colour = "blue") +
  transition_manual(beta) +
  labs(
    y = "Millions",
    title = "Australian population: trend (beta = {format(as.numeric(as.character(current_frame)), nsmall=2)})"
  )
```

## Example: Australian population
\fontsize{8}{7}\sf

```{r holt-fit, echo=TRUE}
aus_economy <- global_economy |>
  filter(Code == "AUS") |>
  mutate(Pop = Population / 1e6)
fit <- aus_economy |>
  model(AAN = ETS(Pop ~ error("A") + trend("A") + season("N")))
report(fit)
```

## Example: Australian population

```{r holt-cmp-plot, echo=TRUE, dependson='holt-fit', fig.height=5}
components(fit) |> autoplot()
```

## Example: Australian population

\fontsize{9}{9}\sf

```{r holt-cmp, echo=TRUE, dependson='holt-fit'}
components(fit) |>
  left_join(fitted(fit), by = c("Country", ".model", "Year"))
```

## Example: Australian population

```{r holt-fc, echo=TRUE, cache=TRUE, dependson='holt-fit'}
fit |>
  forecast(h = 10) |>
  autoplot(aus_economy) +
  labs(y = "Millions", title = "Population: Australia")
```

## Damped trend method
\begin{block}{Component form}\vspace*{-0.3cm}
\begin{align*}
\pred{y}{t+h}{t} &= \ell_{t} + (\phi+\phi^2 + \dots + \phi^{h})b_{t} \\
\ell_{t} &= \alpha y_{t} + (1 - \alpha)(\ell_{t-1} + \phi b_{t-1})\\
b_{t} &= \beta^*(\ell_{t} - \ell_{t-1}) + (1 -\beta^*)\phi b_{t-1}.
\end{align*}
\end{block}
\pause

  * Damping parameter $0<\phi<1$.
  * If $\phi=1$, identical to Holt's linear trend.
  * As $h\rightarrow\infty$, $\pred{y}{T+h}{T}\rightarrow \ell_T+\phi b_T/(1-\phi)$.
  * Short-run forecasts trended, long-run forecasts constant.

<!-- ## Your turn -->
<!-- \large -->

<!--  * Write down the model for ETS(A,A\damped,N) -->

## Example: Australian population

```{r, echo=TRUE}
aus_economy |>
  model(holt = ETS(Pop ~ error("A") + trend("Ad") + season("N"))) |>
  forecast(h = 20) |>
  autoplot(aus_economy)
```

## Example: Australian population
\fontsize{10}{10}\sf

```{r, echo=TRUE}
fit <- aus_economy |>
  filter(Year <= 2010) |>
  model(
    ses = ETS(Pop ~ error("A") + trend("N") + season("N")),
    holt = ETS(Pop ~ error("A") + trend("A") + season("N")),
    damped = ETS(Pop ~ error("A") + trend("Ad") + season("N"))
  )
```

```{r, echo = TRUE, results = 'hide'}
tidy(fit)
accuracy(fit)
```

## Example: Australian population
\fontsize{13}{15}\sf

```{r echo=FALSE}
fit_terms <- tidy(fit) |>
  spread(.model, estimate) |>
  mutate(term = factor(term, levels = c("alpha", "beta", "phi", "l", "b"), labels = c("$\\alpha$", "$\\beta^*$", "$\\phi$", "$\\ell_0$", "$b_0$"))) |>
  arrange(term)

fit_accuracy <- accuracy(fit) |>
  bind_rows(
    forecast(fit, h = 9) |>
      accuracy(aus_economy)
  ) |>
  gather(term, estimate, -Country, -.model, -.type) |>
  spread(.model, estimate) |>
  filter(term == "RMSE" | .type == "Test" & term %in% c("RMSE", "MAE", "MAPE", "MASE")) |>
  arrange(desc(.type), desc(term)) |>
  unite("term", .type, term, sep = " ")

bind_rows(fit_terms, fit_accuracy) |>
  select(term, ses, holt, damped) |>
  rename(SES = ses, `Linear trend` = holt, `Damped trend` = damped) |>
  mutate_if(is.numeric, ~ ifelse(is.na(.), "", formatC(., format = "f", 2))) |>
  knitr::kable(booktabs = TRUE, align = "r")
```

<!--
## Your turn

`prices` contains the price of a dozen eggs in the United States from 1900–1993

 1. Use SES and Holt’s method (with and without damping) to forecast “future” data.

     [Hint: use h=100 so you can clearly see the differences between the options when plotting the forecasts.]
 1. Which method gives the best training RMSE?
 1. Are these RMSE values comparable?
 1. Do the residuals from the best fitting method look like white noise?
 -->
