---
title: "8. Exponential smoothing"
author: "8.3 Methods with seasonality"
date: "OTexts.org/fpp3/"
classoption: aspectratio=169
titlepage: fpp3title.png
titlecolor: fpp3red
toc: false
output:
  binb::monash:
    colortheme: monashwhite
    fig_width: 7.5
    fig_height: 3
    keep_tex: no
    includes:
      in_header: fpp3header.tex
---

```{r setup, include=FALSE}
source("setup.R")
library(patchwork)
library(gganimate)
library(purrr)
library(rlang)
library(magick)
```


## Holt-Winters additive method

Holt and Winters extended Holt's method to capture seasonality.
\begin{block}{Component form}\vspace*{-0.9cm}
\begin{align*}
\pred{y}{t+h}{t} &= \ell_{t} + hb _{t} + s_{t+h-m(k+1)} \\
\ell_{t} &= \alpha(y_{t} - s_{t-m}) + (1 - \alpha)(\ell_{t-1} + b_{t-1})\\
b_{t} &= \beta^*(\ell_{t} - \ell_{t-1}) + (1 - \beta^*)b_{t-1}\\
s_{t} &= \gamma (y_{t}-\ell_{t-1}-b_{t-1}) + (1-\gamma)s_{t-m}
\end{align*}
\end{block}\fontsize{12}{14}\sf

  * $k=$ integer part of $(h-1)/m$. Ensures estimates from the final year are used for forecasting.
  * Parameters:&nbsp; $0\le \alpha\le 1$,&nbsp; $0\le \beta^*\le 1$,&nbsp; $0\le \gamma\le 1-\alpha$&nbsp;  and $m=$  period of seasonality (e.g. $m=4$ for quarterly data).

## Holt-Winters additive method

  * Seasonal component is usually expressed as
        $s_{t} = \gamma^* (y_{t}-\ell_{t})+ (1-\gamma^*)s_{t-m}.$
  * Substitute in for $\ell_t$:
        $s_{t} = \gamma^*(1-\alpha) (y_{t}-\ell_{t-1}-b_{t-1})+ [1-\gamma^*(1-\alpha)]s_{t-m}$
  * We set $\gamma=\gamma^*(1-\alpha)$.
  * The usual parameter restriction is $0\le\gamma^*\le1$, which translates to $0\le\gamma\le(1-\alpha)$.

## Exponential smoothing: seasonality

```{r gamma-anim, cache=TRUE, echo=FALSE, fig.show='animate', interval=1/5, message=FALSE, fig.height=5, fig.width=8, aniopts='controls,buttonsize=0.3cm,width=11.5cm'}
j07 <- PBS |>
  filter(ATC2 == "J07") |>
  summarise(Cost = sum(Cost))
gamma_anim <- map_dfr(set_names(seq(0, 0.99, 0.01), seq(0, 0.99, 0.01)), function(gamma) {
  j07 |>
    model(ETS(Cost ~ error("A") + trend("N",
      alpha = 0.002, alpha_range = c(-1, 1), beta = 0.001,
      beta_range = c(-1, 1)
    ) + season("A", gamma = gamma, gamma_range = c(-1, 1)), bounds = "admissible")) |>
    augment() |>
    as_tibble()
}, .id = "gamma") |>
  mutate(gamma = as.numeric(gamma))
gamma_anim |>
  ggplot(aes(x = Month, y = Cost)) +
  geom_line() +
  geom_line(aes(y = .fitted), colour = "blue") +
  transition_manual(gamma) +
  labs(
    y = "$AUD",
    title = "Medicare Australia cost of vaccine scripts: seasonality (gamma = {format(as.numeric(as.character(current_frame)), nsmall=2)})"
  )
```

## Holt-Winters multiplicative method
\fontsize{13}{13.5}\sf\vspace*{-0.1cm}

Seasonal variations change in proportion to the level of the series.

\begin{block}{Component form}\vspace*{-0.7cm}
    \begin{align*}
        \pred{y}{t+h}{t} &= (\ell_{t} + hb_{t})s_{t+h-m(k+1)} \\
        \ell_{t} &= \alpha \frac{y_{t}}{s_{t-m}} + (1 - \alpha)(\ell_{t-1} + b_{t-1})\\
        b_{t} &= \beta^*(\ell_{t}-\ell_{t-1}) + (1 - \beta^*)b_{t-1}        \\
        s_{t} &= \gamma \frac{y_{t}}{(\ell_{t-1} + b_{t-1})} + (1 - \gamma)s_{t-m}
    \end{align*}
\end{block}\vspace*{-0.2cm}\fontsize{13}{14}\sf

  * $k$ is integer part of $(h-1)/m$.
  * Additive method: $s_t$ in absolute terms --- within each year $\sum_i s_i \approx 0$.
  * Multiplicative method: $s_t$ in relative terms --- within each year $\sum_i s_i \approx m$.

## Example: Australian holiday tourism

```{r 7-HW, echo=TRUE}
aus_holidays <- tourism |>
  filter(Purpose == "Holiday") |>
  summarise(Trips = sum(Trips))
fit <- aus_holidays |>
  model(
    additive = ETS(Trips ~ error("A") + trend("A") + season("A")),
    multiplicative = ETS(Trips ~ error("M") + trend("A") + season("M"))
  )
fc <- fit |> forecast()
```

## Example: Australian holiday tourism

```{r}
fc |>
  autoplot(aus_holidays, level = NULL) +
  labs(y = "Thousands", title = "Overnight trips")
```

## Estimated components
\fontsize{10}{11}\sf

```{r, echo = TRUE}
components(fit)
```

## Estimated components

```{r fig-7-LevelTrendSeas, fig.width=8, fig.height=4, out.width="100%", echo=FALSE}
components(fit) |>
  gather("state", "value", -.model, -Quarter, factor_key = TRUE) |>
  group_by(.model) |>
  group_split() |>
  purrr::map(
    ~ ggplot(., aes(x = Quarter, y = value)) +
      geom_line() +
      facet_grid(state ~ ., scales = "free") +
      labs(
        x = "Year", y = "",
        title = stringr::str_to_title(unique(.$.model)) |> paste("states")
      )
  ) |>
  wrap_plots(ncol = 2)
```

## Holt-Winters damped method
Often the single most accurate forecasting method for seasonal data:
\begin{block}{}\vspace*{-0.4cm}
\begin{align*}
\pred{y}{t+h}{t} &= [\ell_{t} + (\phi+\phi^2 + \dots + \phi^{h})b_{t}]s_{t+h-m(k+1)} \\
\ell_{t} &= \alpha(y_{t} / s_{t-m}) + (1 - \alpha)(\ell_{t-1} + \phi b_{t-1})\\
b_{t} &= \beta^*(\ell_{t} - \ell_{t-1}) + (1 - \beta^*)\phi b_{t-1}       \\
s_{t} &= \gamma \frac{y_{t}}{(\ell_{t-1} + \phi b_{t-1})} + (1 - \gamma)s_{t-m}
\end{align*}
\end{block}

<!--
## Your turn

Apply Holt-Winters’ multiplicative method to the Gas data from `aus_production`.

 1. Why is multiplicative seasonality necessary here?
 1. Experiment with making the trend damped.
 1. Check that the residuals from the best method look like white noise.
 -->

## Holt-Winters with daily data

```{r hwdaily, echo=TRUE, eval=FALSE}
sth_cross_ped <- pedestrian |>
  filter(
    Date >= "2016-07-01",
    Sensor == "Southern Cross Station"
  ) |>
  index_by(Date) |>
  summarise(Count = sum(Count) / 1000)
sth_cross_ped |>
  filter(Date <= "2016-07-31") |>
  model(hw = ETS(Count ~ error("M") + trend("Ad") + season("M"))) |>
  forecast(h = "2 weeks") |>
  autoplot(sth_cross_ped |> filter(Date <= "2016-08-14")) +
  labs(
    title = "Daily traffic: Southern Cross",
    y = "Pedestrians ('000)"
  )
```

## Holt-Winters with daily data

```{r hwdaily_repeat, ref.label="hwdaily", echo=FALSE, eval=TRUE}
```
