---
title: "8. Exponential smoothing"
author: "8.5 Innovations state space models"
date: "OTexts.org/fpp3/"
classoption: aspectratio=169
titlepage: fpp3title.png
titlecolor: fpp3red
toc: false
output:
  binb::monash:
    colortheme: monashwhite
    fig_width: 7.5
    fig_height: 3
    keep_tex: no
    includes:
      in_header: fpp3header.tex
---

```{r setup, include=FALSE}
source("setup.R")
```


## Models and methods

### Methods

  * Algorithms that return point forecasts.

### Models

  * Generate same point forecasts but can also generate forecast distributions.
  * A stochastic (or random) data generating process that can generate an entire forecast distribution.
  * Allow for "proper" model selection.

## ETS(A,N,N): SES with additive errors

\begin{block}{Component form}\vspace*{-0.9cm}
\begin{align*}
\text{Forecast equation}&&\pred{y}{t+h}{t} &= \ell_{t}\\
\text{Smoothing equation}&&\ell_{t} &= \alpha y_{t} + (1 - \alpha)\ell_{t-1}
\end{align*}
\end{block}\pause
Forecast error: $e_t = y_t - \pred{y}{t}{t-1} = y_t - \ell_{t-1}$.\pause
\begin{block}{Error correction form}\vspace*{-0.9cm}
\begin{align*}
y_t &= \ell_{t-1} + e_t\\
\ell_{t}
         &= \ell_{t-1}+\alpha( y_{t}-\ell_{t-1})\\
         &= \ell_{t-1}+\alpha e_{t}
\end{align*}
\end{block}\pause\vspace*{-0.2cm}

Specify probability distribution for $e_t$, we assume $e_t = \varepsilon_t\sim\text{NID}(0,\sigma^2)$.

## ETS(A,N,N): SES with additive errors

\begin{block}{}\vspace*{-0.9cm}
\begin{align*}
\text{Measurement equation}&& y_t &= \ell_{t-1} + \varepsilon_t\\
\text{State equation}&& \ell_t&=\ell_{t-1}+\alpha \varepsilon_t
\end{align*}
\end{block}
where $\varepsilon_t\sim\text{NID}(0,\sigma^2)$.

  * "innovations" or "single source of error" because equations have the same error process, $\varepsilon_t$.
  * Measurement equation: relationship between observations and states.
  * State equation(s): evolution of the state(s) through time.

## ETS(M,N,N): SES with multiplicative errors.

  * Specify relative errors  $\varepsilon_t=\frac{y_t-\pred{y}{t}{t-1}}{\pred{y}{t}{t-1}}\sim \text{NID}(0,\sigma^2)$
  * Substituting $\pred{y}{t}{t-1}=\ell_{t-1}$ gives:
    * $y_t = \ell_{t-1}+\ell_{t-1}\varepsilon_t$
    * $e_t = y_t - \pred{y}{t}{t-1} = \ell_{t-1}\varepsilon_t$

 \pause
\begin{block}{}\vspace*{-0.9cm}
\begin{align*}
\text{Measurement equation}&& y_t &= \ell_{t-1}(1 + \varepsilon_t)\\
\text{State equation}&& \ell_t&=\ell_{t-1}(1+\alpha \varepsilon_t)
\end{align*}
\end{block}
\pause\vspace*{-0.2cm}

  * Models with additive and multiplicative errors with the same parameters generate the same point forecasts but different prediction intervals.

## Specifying the model

```{r ann-spec, echo = TRUE, eval=FALSE}
ETS(y ~ error("A") + trend("N") + season("N"))
ETS(y ~ error("M") + trend("N") + season("N"))
```

By default, an optimal value for $\alpha$ and $\ell_0$ is used.

$\alpha$ can be chosen manually in `trend()`.

```{r alpha-spec, echo = TRUE, eval = FALSE}
trend("N", alpha = 0.5)
trend("N", alpha_range = c(0.2, 0.8))
```

## ETS(A,A,N)

Holt's linear method with additive errors.

  * Assume $\varepsilon_t=y_t-\ell_{t-1}-b_{t-1} \sim \text{NID}(0,\sigma^2)$.
  * Substituting into the error correction equations for Holt's linear method\vspace*{-0.2cm}
  \begin{align*}
      y_t&=\ell_{t-1}+b_{t-1}+\varepsilon_t\\
      \ell_t&=\ell_{t-1}+b_{t-1}+\alpha \varepsilon_t\\
      b_t&=b_{t-1}+\alpha\beta^* \varepsilon_t
  \end{align*}
  * For simplicity, set $\beta=\alpha \beta^*$.

## ETS(M,A,N)

Holt's linear method with multiplicative errors.

  * Assume $\varepsilon_t=\frac{y_t-(\ell_{t-1}+b_{t-1})}{(\ell_{t-1}+b_{t-1})}$
  * Following a similar approach as above, the innovations state space model underlying Holt's linear method with multiplicative errors is specified as\vspace*{-0.4cm}
  \begin{align*}
      y_t&=(\ell_{t-1}+b_{t-1})(1+\varepsilon_t)\\
      \ell_t&=(\ell_{t-1}+b_{t-1})(1+\alpha \varepsilon_t)\\
      b_t&=b_{t-1}+\beta(\ell_{t-1}+b_{t-1}) \varepsilon_t
  \end{align*}
  where again  $\beta=\alpha \beta^*$ and $\varepsilon_t \sim \text{NID}(0,\sigma^2)$.

## Specifying the model

```{r aan-spec, echo = TRUE, eval=FALSE}
ETS(y ~ error("A") + trend("A") + season("N"))
ETS(y ~ error("M") + trend("A") + season("N"))
```

By default, optimal values for $\beta$ and $b_0$ are used.

$\beta$ can be chosen manually in `trend()`.

```{r beta-spec, echo = TRUE, eval = FALSE}
trend("A", beta = 0.004)
trend("A", beta_range = c(0, 0.1))
```

## ETS(A,A,A)

Holt-Winters additive method with additive errors.

\begin{block}{}\vspace*{-0.4cm}
\begin{align*}
\text{Forecast equation} && \hat{y}_{t+h|t} &= \ell_{t} + hb_{t} + s_{t+h-m(k+1)}\\
\text{Observation equation}&& y_t&=\ell_{t-1}+b_{t-1}+s_{t-m} + \varepsilon_t\\
\text{State equations}&& \ell_t&=\ell_{t-1}+b_{t-1}+\alpha \varepsilon_t\\
&&        b_t&=b_{t-1}+\beta \varepsilon_t \\
&&s_t &= s_{t-m} + \gamma\varepsilon_t
\end{align*}
\end{block}

* Forecast errors: $\varepsilon_{t} = y_t - \hat{y}_{t|t-1}$
* $k$ is integer part of $(h-1)/m$.

<!-- ## Your turn -->
<!-- \large -->

<!--  * Write down the model for ETS(A,N,A) -->

## ETS(M,A,M)

Holt-Winters multiplicative method with multiplicative errors.

\begin{block}{}\vspace*{-0.4cm}
\begin{align*}
\text{Forecast equation} && \hat{y}_{t+h|t} &= (\ell_{t} + hb_{t}) s_{t+h-m(k+1)}\\
\text{Observation equation}&& y_t&= (\ell_{t-1}+b_{t-1})s_{t-m}(1 + \varepsilon_t)\\
\text{State equations}&& \ell_t&=(\ell_{t-1}+b_{t-1})(1+\alpha \varepsilon_t)\\
&&        b_t&=b_{t-1} +\beta(\ell_{t-1}+b_{t-1}) \varepsilon_t \\
&&s_t &= s_{t-m}(1 + \gamma\varepsilon_t)
\end{align*}
\end{block}

* Forecast errors: $\varepsilon_{t} = (y_t - \hat{y}_{t|t-1})/\hat{y}_{t|t-1}$
* $k$ is integer part of $(h-1)/m$.

## Specifying the model

```{r aaa-spec, echo = TRUE, eval=FALSE}
ETS(y ~ error("A") + trend("A") + season("A"))
ETS(y ~ error("M") + trend("A") + season("M"))
```

By default, optimal values for $\gamma$ and $s_0,s_{-1},\dots,s_{-m+1}$ are used.

$\gamma$ can be chosen manually in `season()`.

```{r gamma-spec, echo = TRUE, eval = FALSE}
season("A", gamma = 0.004)
season("A", gamma_range = c(0, 0.1))
```


## ETS models
\fontsize{11}{12}\sf

\begin{block}{}
\begin{tabular}{ll|ccc}
  \multicolumn{2}{l}{\alert{\bf Additive Error}} &        \multicolumn{3}{c}{\bf Seasonal Component}         \\
          \multicolumn{2}{c|}{\bf Trend}         &         N         &         A         &         M         \\
        \multicolumn{2}{c|}{\bf Component}       &     ~(None)~      &    (Additive)     & (Multiplicative)  \\ \cline{3-5}
           &                                     &                   &                   &  \\[-0.4cm]
  N        & (None)                              &       A,N,N       &       A,N,A       &    A,N,M     \\
           &                                     &                   &                   &  \\[-0.4cm]
  A        & (Additive)                          &       A,A,N       &       A,A,A       &    A,A,M     \\
           &                                     &                   &                   &  \\[-0.4cm]
  A\damped & (Additive damped)                   &   A,A\damped,N    &   A,A\damped,A    & A,A\damped,M
\end{tabular}
\end{block}

\begin{block}{}
\begin{tabular}{ll|ccc}
  \multicolumn{2}{l}{\alert{\bf Multiplicative Error}} &     \multicolumn{3}{c}{\bf Seasonal Component}      \\
             \multicolumn{2}{c|}{\bf Trend}            &      N       &         A         &        M         \\
           \multicolumn{2}{c|}{\bf Component}          &   ~(None)~   &    (Additive)     & (Multiplicative) \\ \cline{3-5}
           &                                           &              &                   &  \\[-0.4cm]
  N        & (None)                                    &    M,N,N     &       M,N,A       &      M,N,M       \\
           &                                           &              &                   &  \\[-0.4cm]
  A        & (Additive)                                &    M,A,N     &       M,A,A       &      M,A,M       \\
           &                                           &              &                   &  \\[-0.4cm]
  A\damped & (Additive damped)                         & M,A\damped,N &   M,A\damped,A    &   M,A\damped,M
\end{tabular}
\end{block}

## Additive error models

\placefig{0}{1.5}{width=16.cm,trim=0 120 0 0,clip=true}{fig_7_ets_add.pdf}

## Multiplicative error models

\placefig{0}{1.5}{width=16.cm,trim=0 120 0 0,clip=true}{fig_7_ets_multi.pdf}
