---
title: "9. ARIMA models"
author: "9.10 ARIMA vs ETS"
date: "OTexts.org/fpp3/"
classoption: aspectratio=169
titlepage: fpp3title.png
titlecolor: fpp3red
toc: false
output:
  binb::monash:
    colortheme: monashwhite
    fig_width: 7.5
    fig_height: 2.8
    keep_tex: no
    includes:
      in_header: fpp3header.tex
---
```{r setup, include=FALSE}
source("setup.R")
library(patchwork)
library(purrr)
```

## ARIMA vs ETS
\fontsize{14}{16}\sf

  * Myth that ARIMA models are more general than exponential smoothing.
  * Linear exponential smoothing models all special cases of ARIMA models.
  * Non-linear exponential smoothing models have no equivalent ARIMA counterparts.
  * Many ARIMA models have no exponential smoothing counterparts.
  * ETS models all non-stationary. Models with seasonality or non-damped trend (or both) have two unit roots; all other models have one unit \rlap{root.}

\vspace*{10cm}

## ARIMA vs ETS

```{r venn, echo=FALSE}
library(latex2exp)
cols <- c(ets = "#D55E00", arima = "#0072b2")
tibble(
  x = c(-1, 1),
  y = c(-0.5, -0.5),
  labels = c("ets", "arima"),
) |>
  ggplot(aes(colour = labels, fill = labels)) +
  ggforce::geom_circle(aes(x0 = x, y0 = y, r = 1.6), alpha = 0.3, size = 1) +
  scale_colour_manual(values = cols) +
  scale_fill_manual(values = cols) +
  coord_fixed() +
  guides(fill = "none") +
  geom_text(aes(label = "ETS models", x = -1.5, y = 1.35), col = cols["ets"], fontface = "bold", size = 6) +
  geom_text(aes(label = "Combination\n of components", x = -1.3, y = 0.5), col = cols["ets"], size = 2.8, fontface = "bold") +
  geom_text(aes(label = "9 ETS models with\n multiplicative errors", x = -1.6, y = -0.5), col = cols["ets"], size = 2.8) +
  geom_text(aes(label = "3 ETS models with\n additive errors and\n multiplicative\n seasonality", x = -1.3, y = -1.4), col = cols["ets"], size = 2.8) +
  geom_text(aes(label = "ARIMA models", x = 1.5, y = 1.35), col = cols["arima"], fontface = "bold", size = 6) +
  geom_text(aes(label = "Modelling\n autocorrelations", x = 1.3, y = 0.5), col = cols["arima"], size = 2.8, fontface = "bold") +
  annotate("text", label = TeX("Potentially $\\infty$ models"), x = 1.6, y = -0.6, col = cols["arima"], size = 2.8) +
  geom_text(aes(label = "All stationary models\n Many large models", x = 1.25, y = -1.5), col = cols["arima"], size = 2.8) +
  geom_text(aes(label = "6 fully additive\n ETS models", x = 0, y = -0.6), col = "#6b6859", size = 2.8) +
  guides(colour = "none", fill = "none") +
  theme_void()
```

## Equivalences

\fontsize{13}{15}\sf

|**ETS model**  | **ARIMA model**             | **Parameters**                       |
| :------------ | :-------------------------- | :----------------------------------- |
| ETS(A,N,N)    | ARIMA(0,1,1)                | $\theta_1 = \alpha-1$                |
| ETS(A,A,N)    | ARIMA(0,2,2)                | $\theta_1 = \alpha+\beta-2$          |
|               |                             | $\theta_2 = 1-\alpha$                |
| ETS(A,A\damped,N)    | ARIMA(1,1,2)                | $\phi_1=\phi$                        |
|               |                             | $\theta_1 = \alpha+\phi\beta-1-\phi$ |
|               |                             | $\theta_2 = (1-\alpha)\phi$          |
| ETS(A,N,A)    | ARIMA(0,0,$m$)(0,1,0)$_m$   |                                      |
| ETS(A,A,A)    | ARIMA(0,1,$m+1$)(0,1,0)$_m$ |                                      |
| ETS(A,A\damped,A)    | ARIMA(1,0,$m+1$)(0,1,0)$_m$ |                                      |

## Example: Australian population
\fontsize{10}{11}\sf

```{r tscvpop, echo=TRUE, warning=FALSE}
aus_economy <- global_economy |>
  filter(Code == "AUS") |>
  mutate(Population = Population / 1e6)
aus_economy |>
  slice(-n()) |>
  stretch_tsibble(.init = 10) |>
  model(ets = ETS(Population), arima = ARIMA(Population)) |>
  forecast(h = 1) |>
  accuracy(aus_economy) |>
  select(.model, ME:RMSSE)
```

## Example: Australian population

```{r popetsplot, echo=TRUE, fig.height=2.6}
aus_economy |>
  model(ETS(Population)) |>
  forecast(h = "5 years") |>
  autoplot(aus_economy) +
  labs(title = "Australian population", y = "People (millions)")
```

## Example: Cement production

```{r qcement1, echo=TRUE}
cement <- aus_production |>
  select(Cement) |>
  filter_index("1988 Q1" ~ .)
train <- cement |> filter_index(. ~ "2007 Q4")
fit <- train |>
  model(
    arima = ARIMA(Cement),
    ets = ETS(Cement)
  )
```

## Example: Cement production
\fontsize{10}{11}\sf

```{r qcement2, dependson="qcement1"}
fit |>
  select(arima) |>
  report()
```

## Example: Cement production
\fontsize{10}{9.5}\sf

```{r qcement3, dependson="qcement1"}
fit |>
  select(ets) |>
  report()
```

## Example: Cement production

```{r qcement4, dependson="qcement1", fig.height=4}
gg_tsresiduals(fit |> select(arima), lag_max = 16)
```

## Example: Cement production

```{r qcement5, dependson="qcement1", fig.height=4}
gg_tsresiduals(fit |> select(ets), lag_max = 16)
```

## Example: Cement production
\fontsize{10}{12}\sf

```{r qcement6, dependson="qcement1"}
fit |>
  select(arima) |>
  augment() |>
  features(.innov, ljung_box, lag = 16, dof = 6)
```

## Example: Cement production
\fontsize{10}{12}\sf

```{r qcement7, dependson="qcement1"}
fit |>
  select(ets) |>
  augment() |>
  features(.innov, ljung_box, lag = 16)
```

## Example: Cement production
\fontsize{10}{12}\sf

```{r qcement8, dependson=c("qcement2","qcement3")}
fit |>
  forecast(h = "2 years 6 months") |>
  accuracy(cement) |>
  select(-ME, -MPE, -ACF1)
```

## Example: Cement production

```{r qcement9, echo=TRUE, dependson="qcement1", fig.height=2.6}
fit |>
  select(arima) |>
  forecast(h = "3 years") |>
  autoplot(cement) +
  labs(title = "Cement production in Australia", y = "Tonnes ('000)")
```
