---
title: "9. ARIMA models"
author: "9.3 Autoregressive models"
date: "OTexts.org/fpp3/"
classoption: aspectratio=169
titlepage: fpp3title.png
titlecolor: fpp3red
toc: false
output:
  binb::monash:
    colortheme: monashwhite
    fig_width: 7.5
    fig_height: 3
    keep_tex: no
    includes:
      in_header: fpp3header.tex
---

```{r setup, include=FALSE}
source("setup.R")
library(patchwork)
```

## Autoregressive models

\begin{block}{Autoregressive (AR) models:}
$$
  y_{t} = c + \phi_{1}y_{t - 1} + \phi_{2}y_{t - 2} + \cdots + \phi_{p}y_{t - p} + \varepsilon_{t},
$$
where $\varepsilon_t$ is white noise. This is a multiple regression with \textbf{lagged values} of $y_t$ as predictors.
\end{block}

```{r arp, echo=FALSE, fig.height=2.4}
set.seed(1)
p1 <- tsibble(idx = seq_len(100), sim = 10 + arima.sim(list(ar = -0.8), n = 100), index = idx) |>
  autoplot(sim) + labs(y = "", title = "AR(1)")
p2 <- tsibble(idx = seq_len(100), sim = 20 + arima.sim(list(ar = c(1.3, -0.7)), n = 100), index = idx) |>
  autoplot(sim) + labs(y = "", title = "AR(2)")
p1 | p2
```

## AR(1) model

\begin{block}{}
\centerline{$y_{t} = 18 -0.8 y_{t - 1} + \varepsilon_{t}$}
\end{block}
\rightline{$\varepsilon_t\sim N(0,1)$,\quad $T=100$.}

```{r, echo=FALSE}
p1
```

## AR(1) model

\begin{block}{}
  \centerline{$y_{t} = c + \phi_1 y_{t - 1} + \varepsilon_{t}$}
\end{block}

* When $\phi_1=0$, $y_t$ is **equivalent to WN**
* When $\phi_1=1$ and $c=0$, $y_t$ is **equivalent to a RW**
* When $\phi_1=1$ and $c\ne0$, $y_t$ is **equivalent to a RW with drift**
* When $\phi_1<0$, $y_t$ tends to **oscillate between positive and negative values**.

## AR(2) model

\begin{block}{}
  \centerline{$y_t = 8 + 1.3y_{t-1} - 0.7 y_{t-2} + \varepsilon_t$}
\end{block}
\rightline{$\varepsilon_t\sim N(0,1)$, \qquad $T=100$.}

```{r, echo=FALSE}
p2
```

## Stationarity conditions

We normally restrict autoregressive models to stationary data, and then some constraints on the values of the parameters are required.

\begin{block}{General condition for stationarity}
  Complex roots of $1-\phi_1 z - \phi_2 z^2 - \dots - \phi_pz^p$ lie outside the unit circle on the complex plane.
\end{block}\pause

* For $p=1$: $-1<\phi_1<1$.
* For $p=2$: $-1<\phi_2<1\qquad \phi_2+\phi_1 < 1 \qquad \phi_2 -\phi_1 < 1$.
* More complicated conditions hold for $p\ge3$.
* Estimation software takes care of this.
