---
title: "9. ARIMA models"
author: "9.4 Moving average models"
date: "OTexts.org/fpp3/"
classoption: aspectratio=169
titlepage: fpp3title.png
titlecolor: fpp3red
toc: false
output:
  binb::monash:
    colortheme: monashwhite
    fig_width: 7.5
    fig_height: 3
    keep_tex: no
    includes:
      in_header: fpp3header.tex
---

```{r setup, include=FALSE}
source("setup.R")
library(patchwork)
library(purrr)
```

## Moving Average (MA) models

\begin{block}{Moving Average (MA) models:}
$$
  y_{t} = c + \varepsilon_t + \theta_{1}\varepsilon_{t - 1} + \theta_{2}\varepsilon_{t - 2} + \cdots + \theta_{q}\varepsilon_{t - q},
$$
where $\varepsilon_t$ is white noise.
This is a multiple regression with \textbf{past \emph{errors}} as predictors. \emph{Don't confuse this with moving average smoothing!}
\end{block}

```{r maq, fig.height=2.4, echo=FALSE}
set.seed(2)
p1 <- tsibble(idx = seq_len(100), sim = 20 + arima.sim(list(ma = 0.8), n = 100), index = idx) |>
  autoplot(sim) + labs(y = "", title = "MA(1)")
p2 <- tsibble(idx = seq_len(100), sim = arima.sim(list(ma = c(-1, +0.8)), n = 100), index = idx) |>
  autoplot(sim) + labs(y = "", title = "MA(2)")

p1 | p2
```

## MA(1) model

\begin{block}{}
  \centerline{$y_t = 20 + \varepsilon_t + 0.8 \varepsilon_{t-1}$}
\end{block}
\rightline{$\varepsilon_t\sim N(0,1)$,\quad $T=100$.}

```{r, echo=FALSE}
p1
```

## MA(2) model

\begin{block}{}
  \centerline{$y_t = \varepsilon_t -\varepsilon_{t-1} + 0.8 \varepsilon_{t-2}$}
\end{block}
\rightline{$\varepsilon_t\sim N(0,1)$,\quad $T=100$.}

```{r, echo=FALSE}
p2
```

## MA($\infty$) models
\fontsize{13.5}{16}\sf

It is possible to write any stationary AR($p$) process as an MA($\infty$) process.

**Example: AR(1)**
\begin{align*}
y_t &= \phi_1y_{t-1} + \varepsilon_t\\
&= \phi_1(\phi_1y_{t-2} + \varepsilon_{t-1}) + \varepsilon_t\\
&= \phi_1^2y_{t-2} + \phi_1 \varepsilon_{t-1} + \varepsilon_t\\
&= \phi_1^3y_{t-3} + \phi_1^2\varepsilon_{t-2} + \phi_1 \varepsilon_{t-1} + \varepsilon_t\\
&\dots
\end{align*}\pause
Provided $-1<\phi_1<1$:
$$y_t = \varepsilon_t + \phi_1\varepsilon_{t-1}+ \phi_1^2\varepsilon_{t-2}+ \phi_1^3\varepsilon_{t-3} + \cdots$$

## Invertibility

* Any MA($q$) process can be written as an AR($\infty$) process if we impose some constraints on the MA parameters.
* Then the MA model is called "invertible".
* Invertible models have some mathematical properties that make them easier to use in practice.
* Invertibility of an ARIMA model is equivalent to forecastability of an ETS model.

## Invertibility

\begin{block}{General condition for invertibility}
  Complex roots of $1+\theta_1 z + \theta_2 z^2 + \dots + \theta_qz^q$ lie outside the unit circle on the complex plane.
\end{block}\pause

* For $q=1$: $-1<\theta_1<1$.
* For $q=2$: $-1<\theta_2<1\qquad \theta_2+\theta_1 >-1 \qquad \theta_1 -\theta_2 < 1$.
* More complicated conditions hold for $q\ge3$.
* Estimation software takes care of this.
