---
title: "9. ARIMA models"
author: "9.5 Non-seasonal ARIMA models"
date: "OTexts.org/fpp3/"
classoption: aspectratio=169
titlepage: fpp3title.png
titlecolor: fpp3red
toc: false
output:
  binb::monash:
    colortheme: monashwhite
    fig_width: 7.5
    fig_height: 2.8
    keep_tex: no
    includes:
      in_header: fpp3header.tex
---

```{r setup, include=FALSE}
source("setup.R")
library(patchwork)
library(purrr)
```

## ARIMA models

\begin{tabular}{@{}rl}
\textbf{AR}: & autoregressive (lagged observations as inputs)\\
\textbf{I}: & integrated (differencing to make series stationary)\\
\textbf{MA}: & moving average (lagged errors as inputs)
\end{tabular}

\pause

###
An ARIMA model is rarely interpretable in terms of visible data structures like trend and seasonality. But it can capture a huge range of time series patterns.


## ARIMA models

\begin{block}{Autoregressive Moving Average models:}\vspace*{-0.7cm}
\begin{align*}
  y_{t} &= c + \phi_{1}y_{t - 1} + \cdots + \phi_{p}y_{t - p} \\
        & \hspace*{2.4cm}\text{} + \theta_{1}\varepsilon_{t - 1} + \cdots + \theta_{q}\varepsilon_{t - q} + \varepsilon_{t}.
\end{align*}
\end{block}\pause

* Predictors include both **lagged values of $y_t$ and lagged errors.**
* Conditions on AR coefficients ensure stationarity.
* Conditions on MA coefficients ensure invertibility.
\pause

### Autoregressive Integrated Moving Average models
* Combine ARMA model with **differencing**.
* $(1-B)^d y_t$ follows an ARMA model.

## ARIMA models
\fontsize{13}{15}\sf
\alert{Autoregressive Integrated Moving Average models}
\begin{block}{ARIMA($p, d, q$) model}
\begin{tabular}{rl}
AR:& $p =$ order of the autoregressive part\\
I: & $d =$ degree of first differencing involved\\
MA:& $q =$ order of the moving average part.
\end{tabular}
\end{block}

* White noise model: ARIMA(0,0,0)
* Random walk: ARIMA(0,1,0) with no constant
* Random walk with drift: ARIMA(0,1,0) with \rlap{const.}
* AR($p$): ARIMA($p$,0,0)
* MA($q$): ARIMA(0,0,$q$)

## Backshift notation for ARIMA

* ARMA model:\vspace*{-1cm}\newline
\parbox{12cm}{\small\begin{align*}
\hspace*{-1cm}
y_{t} &= c + \phi_{1}By_{t} + \cdots + \phi_pB^py_{t}
           + \varepsilon_{t} + \theta_{1}B\varepsilon_{t} + \cdots + \theta_qB^q\varepsilon_{t} \\
\hspace*{-1cm} \text{or}\quad
      & (1-\phi_1B - \cdots - \phi_p B^p) y_t = c + (1 + \theta_1 B + \cdots + \theta_q B^q)\varepsilon_t
\end{align*}}

* ARIMA(1,1,1) model:\vspace*{-0.5cm}

\[
\begin{array}{c c c c}
(1 - \phi_{1} B) & (1 - B) y_{t} &= &c + (1 + \theta_{1} B) \varepsilon_{t}\\
{\uparrow} & {\uparrow} & &{\uparrow}\\
{\text{AR(1)}} & {\text{First}} &  &{\text{MA(1)}}\\
& {\hbox to 0cm{\hss\text{difference}\hss}}\\
\end{array}
\]\pause\vspace*{-0.2cm}
$$\text{Expand:}\qquad
  y_t = c + y_{t-1} + \phi_1 y_{t-1}- \phi_1 y_{t-2} + \theta_1\varepsilon_{t-1} + \varepsilon_t
$$

## R model

\begin{block}{Intercept form}
  \centerline{$(1-\phi_1B - \cdots - \phi_p B^p) y_t' = c + (1 + \theta_1 B + \cdots + \theta_q B^q)\varepsilon_t$}
\end{block}

\begin{block}{Mean form}
  \centerline{$(1-\phi_1B - \cdots - \phi_p B^p)(y_t' - \mu) = (1 + \theta_1 B + \cdots + \theta_q B^q)\varepsilon_t$}
\end{block}

 * $y_t' = (1-B)^d y_t$
 * $\mu$ is the mean of $y_t'$.
 * $c = \mu(1-\phi_1 - \cdots - \phi_p )$.
 * fable uses intercept form

## Egyptian exports

```{r egyptexportsauto, echo=TRUE}
global_economy |>
  filter(Code == "EGY") |>
  autoplot(Exports) +
  labs(y = "% of GDP", title = "Egyptian Exports")
```

## Egyptian exports
\fontsize{10}{9}\sf

```{r, echo=TRUE, dependson="egyptexportsauto"}
fit <- global_economy |>
  filter(Code == "EGY") |>
  model(ARIMA(Exports))
report(fit)
```

\pause\vspace*{-0.6cm}

```{r egyptexportsmodel, include=FALSE, warning=FALSE, dependson="egyptexportsauto"}
stopifnot(identical(
  unlist(fit[1, 2][[1]][[1]]$fit$spec),
  c(p = 2L, d = 0L, q = 1L, P = 0, D = 0, Q = 0, constant = TRUE, period = 1)
))
coef <- rlang::set_names(tidy(fit)$estimate, tidy(fit)$term)
```

### ARIMA(2,0,1) model:
$$
  y_t = `r sprintf("%.2f", coef['constant'])`
         + `r sprintf("%.2f", coef['ar1'])` y_{t-1}
          `r sprintf("%.2f", coef['ar2'])` y_{t-2}
          `r sprintf("%.2f", coef['ma1'])` \varepsilon_{t-1}
          + \varepsilon_{t},
$$
where $\varepsilon_t$ is white noise with a standard deviation of $`r sprintf("%.3f", sqrt(glance(fit)$sigma2))` = \sqrt{`r sprintf("%.3f", glance(fit)$sigma2)`}$.

## Egyptian exports

```{r, echo=TRUE, fig.height=4}
gg_tsresiduals(fit)
```

## Egyptian exports

```{r egyptexportsf, dependson="egyptexportsauto"}
fit |>
  forecast(h = 10) |>
  autoplot(global_economy) +
  labs(y = "% of GDP", title = "Egyptian Exports")
```

## Understanding ARIMA models
\fontsize{14}{16}\sf

* If $c=0$ and $d=0$, the long-term forecasts will go to zero.
* If $c=0$ and $d=1$, the long-term forecasts will go to a non-zero constant.
* If $c=0$ and $d=2$, the long-term forecasts will follow a straight line.

* If $c\ne0$ and $d=0$, the long-term forecasts will go to the mean of the data.
* If $c\ne0$ and $d=1$, the long-term forecasts will follow a straight line.
* If $c\ne0$ and $d=2$, the long-term forecasts will follow a quadratic trend.

## Understanding ARIMA models
\fontsize{14}{16}\sf

### Forecast variance and $d$
  * The higher the value of $d$, the more rapidly the prediction intervals increase in size.
  * For $d=0$, the long-term forecast standard deviation will go to the standard deviation of the historical data.

### Cyclic behaviour
  * For cyclic forecasts, $p\ge2$ and some restrictions on coefficients are required.
  * If $p=2$, we need $\phi_1^2+4\phi_2<0$. Then average cycle of length
\[
  (2\pi)/\left[\text{arc cos}(-\phi_1(1-\phi_2)/(4\phi_2))\right].
\]

## Partial autocorrelations

\fontsize{13}{14}\sf

\alert{Partial autocorrelations} measure relationship\newline between $y_{t}$ and $y_{t - k}$, when the effects of other time lags --- $1, 2, 3, \dots, k - 1$ --- are removed.\pause
\begin{block}{}
\begin{align*}
\alpha_k &= \text{$k$th partial autocorrelation coefficient}\\
         &= \text{equal to the estimate of $\phi_k$ in regression:}\\
         & \hspace*{0.8cm} y_t = c + \phi_1 y_{t-1} + \phi_2 y_{t-2} + \dots + \phi_k y_{t-k} +\varepsilon_t.
\end{align*}
\end{block}\pause

* Varying number of terms on RHS gives $\alpha_k$ for different values of $k$.
* $\alpha_1=\rho_1$
* same critical values of $\pm 1.96/\sqrt{T}$ as for ACF.
* Last significant $\alpha_k$ indicates the order of an AR model.

## Egyptian exports

```{r, eval=FALSE}
egypt <- global_economy |> filter(Code == "EGY")
egypt |> ACF(Exports) |> autoplot()
egypt |> PACF(Exports) |> autoplot()
```

```{r, echo=FALSE}
p1 <- global_economy |>
  filter(Code == "EGY") |>
  ACF(Exports) |>
  autoplot()
p2 <- global_economy |>
  filter(Code == "EGY") |>
  PACF(Exports) |>
  autoplot()
p1 | p2
```

## Egyptian exports

```{r, echo=TRUE, fig.height=3.25}
global_economy |>
  filter(Code == "EGY") |>
  gg_tsdisplay(Exports, plot_type = "partial")
```

## ACF and PACF interpretation

**AR(1)**
\begin{align*}
\hspace*{1cm}
  \rho_k &= \phi_1^k\qquad\text{for $k=1,2,\dots$};\\
\alpha_1 &= \phi_1 \qquad\alpha_k = 0\qquad\text{for $k=2,3,\dots$}.
\end{align*}

So we have an AR(1) model when

  * autocorrelations exponentially decay
  * there is a single significant partial autocorrelation.

## ACF and PACF interpretation

**AR($p$)**

  * ACF dies out in an exponential or damped sine-wave manner
  * PACF has all zero spikes beyond the $p$th spike

So we have an AR($p$) model when

  * the ACF is exponentially decaying or sinusoidal
  * there is a significant spike at lag $p$ in PACF, but none beyond $p$

## ACF and PACF interpretation

**MA(1)**
\begin{align*}
\hspace*{1cm}\rho_1 &= \theta_1/(1 + \theta_1^2)\qquad \rho_k = 0\qquad\text{for $k=2,3,\dots$};\\
\alpha_k &= -(-\theta_1)^k/(1 + \theta_1^2 + \dots + \theta_1^{2k})
\end{align*}

So we have an MA(1) model when

 * the PACF is exponentially decaying and
 * there is a single significant spike in ACF

## ACF and PACF interpretation

**MA($q$)**

 * PACF dies out in an exponential or damped sine-wave manner
 * ACF has all zero spikes beyond the $q$th spike

So we have an MA($q$) model when

  * the PACF is exponentially decaying or sinusoidal
  * there is a significant spike at lag $q$ in ACF, but none beyond $q$
