---
title: "9. ARIMA models"
author: "9.9 Seasonal ARIMA models"
date: "OTexts.org/fpp3/"
classoption: aspectratio=169
titlepage: fpp3title.png
titlecolor: fpp3red
toc: false
output:
  binb::monash:
    colortheme: monashwhite
    fig_width: 7.5
    fig_height: 2.8
    keep_tex: no
    includes:
      in_header: fpp3header.tex
---
```{r setup, include=FALSE}
source("setup.R")
library(patchwork)
library(purrr)
```

## Seasonal ARIMA models

| ARIMA | $~\underbrace{(p, d, q)}$ | $\underbrace{(P, D, Q)_{m}}$ |
| ----: | :-----------------------: | :--------------------------: |
|       | ${\uparrow}$              | ${\uparrow}$                 |
|       | Non-seasonal part         | Seasonal part of             |
|       | of the model              | of the model                 |

where $m =$ number of observations per year.

## Seasonal ARIMA models

E.g., ARIMA$(1, 1, 1)(1, 1, 1)_{4}$ model (without constant)\pause
$$(1 - \phi_{1}B)(1 - \Phi_{1}B^{4}) (1 - B) (1 - B^{4})y_{t} ~= ~
(1 + \theta_{1}B) (1 + \Theta_{1}B^{4})\varepsilon_{t}.
$$\pause\vspace*{-1cm}

\setlength{\unitlength}{1mm}
\begin{footnotesize}
\begin{picture}(100,25)(-5,0)
\thinlines
{\put(15,22){\vector(0,1){6}}}
{\put(32,10){\vector(0,1){18}}}
{\put(54,22){\vector(0,1){6}}}
{\put(67,10){\vector(0,1){18}}}
{\put(97,22){\vector(0,1){6}}}
{\put(115,10){\vector(0,1){18}}}
{\put(0,17){$\left(\begin{array}{@{}c@{}} \text{Non-seasonal} \\ \text{AR(1)}
                    \end{array}\right)$}}
{\put(22,5){$\left(\begin{array}{@{}c@{}} \text{Seasonal} \\ \text{AR(1)}
                    \end{array}\right)$}}
{\put(40,17){$\left(\begin{array}{@{}c@{}} \text{Non-seasonal} \\ \text{difference}
                    \end{array}\right)$}}
{\put(60,5){$\left(\begin{array}{@{}c@{}} \text{Seasonal} \\ \text{difference}
                    \end{array}\right)$}}
{\put(85,17){$\left(\begin{array}{@{}c@{}} \text{Non-seasonal} \\ \text{MA(1)}
                    \end{array}\right)$}}
{\put(108,5){$\left(\begin{array}{@{}c@{}} \text{Seasonal} \\ \text{MA(1)}
                    \end{array}\right)$}}
\end{picture}
\end{footnotesize}

\vspace*{10cm}

## Seasonal ARIMA models

E.g., ARIMA$(1, 1, 1)(1, 1, 1)_{4}$ model (without constant)
$$(1 - \phi_{1}B)(1 - \Phi_{1}B^{4}) (1 - B) (1 - B^{4})y_{t} ~= ~
(1 + \theta_{1}B) (1 + \Theta_{1}B^{4})\varepsilon_{t}.
$$\vspace*{-0.8cm}

All the factors can be multiplied out and the general model
written as follows:\vspace*{-0.7cm}
\begin{align*}
y_{t} &= (1 + \phi_{1})y_{t - 1} - \phi_1y_{t-2} + (1 + \Phi_{1})y_{t - 4}\\
&\text{}
 - (1 + \phi_{1} + \Phi_{1} + \phi_{1}\Phi_{1})y_{t - 5}
 + (\phi_{1} + \phi_{1} \Phi_{1}) y_{t - 6} \\
& \text{} - \Phi_{1} y_{t - 8} + (\Phi_{1} + \phi_{1} \Phi_{1}) y_{t - 9}
  - \phi_{1} \Phi_{1} y_{t - 10}\\
  &\text{}
  + \varepsilon_{t} + \theta_{1}\varepsilon_{t - 1} + \Theta_{1}\varepsilon_{t - 4} + \theta_{1}\Theta_{1}\varepsilon_{t - 5}.
\end{align*}
\vspace*{10cm}

## Seasonal ARIMA models
The seasonal part of an AR or MA model will be seen in the seasonal lags of
the PACF and ACF.

\alert{ARIMA(0,0,0)(0,0,1)$_{12}$ will show:}

  * a spike at lag 12 in the ACF but no other significant spikes.
  * The PACF will show exponential decay in the seasonal lags; that is, at lags 12, 24, 36, \dots.

\alert{ARIMA(0,0,0)(1,0,0)$_{12}$ will show:}

  * exponential decay in the seasonal lags of the ACF
  * a single significant spike at lag 12 in the PACF.

## US leisure employment

```{r, fig.height=2.45}
leisure <- us_employment |>
  filter(Title == "Leisure and Hospitality", year(Month) > 2000) |>
  mutate(Employed = Employed / 1000) |>
  select(Month, Employed)
autoplot(leisure, Employed) +
  labs(title = "US employment: leisure & hospitality", y = "People (millions)")
```

## US leisure employment

```{r, fig.height=3}
leisure |>
  gg_tsdisplay(difference(Employed, 12), plot_type = "partial", lag = 36) +
  labs(title = "Seasonally differenced", y = "")
```

## US leisure employment

```{r, fig.height=2.8}
leisure |>
  gg_tsdisplay(difference(Employed, 12) |> difference(),
    plot_type = "partial", lag = 36) +
  labs(title = "Double differenced", y = "")
```

## US leisure employment
\fontsize{10}{11}\sf

```{r}
fit <- leisure |>
  model(arima012011 = ARIMA(Employed ~ pdq(0, 1, 2) + PDQ(0, 1, 1)),
        arima210011 = ARIMA(Employed ~ pdq(2, 1, 0) + PDQ(0, 1, 1)),
        auto = ARIMA(Employed, stepwise = FALSE, approx = FALSE))
fit |>
  pivot_longer(everything(),
    names_to = "Model name",
    values_to = "Orders")
```

## US leisure employment
\fontsize{10}{11}\sf

```{r}
glance(fit) |>
  arrange(AICc) |>
  select(.model:BIC)
```

## US leisure employment

```{r}
fit |>
  select(auto) |>
  gg_tsresiduals(lag = 36)
```

## US leisure employment
\fontsize{10}{11}\sf

```{r}
augment(fit) |> features(.innov, ljung_box, lag = 24, dof = 4)
```

## US leisure employment

```{r}
forecast(fit, h = 36) |>
  filter(.model == "auto") |>
  autoplot(leisure) +
  labs(title = "US employment: leisure & hospitality", y = "People (millions)")
```

## Cortecosteroid drug sales

```{r}
h02 <- PBS |>
  filter(ATC2 == "H02") |>
  summarise(Cost = sum(Cost) / 1e6)
```

## Cortecosteroid drug sales

```{r}
h02 |> autoplot(
  Cost
)
```

## Cortecosteroid drug sales

```{r}
h02 |> autoplot(
  log(Cost)
)
```

## Cortecosteroid drug sales

```{r}
h02 |> autoplot(
  log(Cost) |> difference(12)
)
```

## Cortecosteroid drug sales

```{r h02b}
h02 |> gg_tsdisplay(difference(log(Cost), 12),
  lag_max = 36, plot_type = "partial"
)
```

## Cortecosteroid drug sales

  * Choose $D=1$ and $d=0$.
  * Spikes in PACF at lags 12 and 24 suggest seasonal AR(2) term.
  * Spikes in PACF suggests possible non-seasonal AR(3) term.
  * Initial candidate model: ARIMA(3,0,0)(2,1,0)$_{12}$.

## Cortecosteroid drug sales

```{r h02aicc, echo=FALSE}
models <- list(
  c(3, 0, 0, 2, 1, 0),
  c(3, 0, 1, 2, 1, 0),
  c(3, 0, 2, 2, 1, 0),
  c(3, 0, 1, 1, 1, 0),
  c(3, 0, 1, 0, 1, 1),
  c(3, 0, 1, 0, 1, 2),
  c(3, 0, 1, 1, 1, 1)
)
model_defs <- map(models, ~ ARIMA(log(Cost) ~ 0 + pdq(!!.[1], !!.[2], !!.[3]) + PDQ(!!.[4], !!.[5], !!.[6])))
model_defs <- set_names(model_defs, map_chr(
  models,
  ~ sprintf("ARIMA(%i,%i,%i)(%i,%i,%i)[12]", .[1], .[2], .[3], .[4], .[5], .[6])
))

fit <- h02 |>
  model(!!!model_defs)

fit |>
  glance() |>
  arrange(AICc) |>
  select(.model, AICc) |>
  knitr::kable(digits = 2, row.names = FALSE, align = "cc", booktabs = TRUE)
```

## Cortecosteroid drug sales
\fontsize{10}{12}\sf

```{r arimah02, echo=TRUE}
fit <- h02 |>
  model(best = ARIMA(log(Cost) ~ 0 + pdq(3, 0, 1) + PDQ(0, 1, 2)))
report(fit)
```

## Cortecosteroid drug sales

```{r h02res, echo=TRUE, fig.height=4, dependson='arimah02'}
gg_tsresiduals(fit)
```

## Cortecosteroid drug sales
\fontsize{10}{13}\sf

```{r h02resb, echo = TRUE, fig.height=4, dependson='arimah02'}
augment(fit) |>
  features(.innov, ljung_box, lag = 36, dof = 6)
```

## Cortecosteroid drug sales
\fontsize{10}{13}\sf

```{r h02auto, echo=TRUE, fig.height=3.6}
fit <- h02 |> model(auto = ARIMA(log(Cost)))
report(fit)
```

## Cortecosteroid drug sales

```{r, echo=TRUE, fig.height=4, dependson='h02auto'}
gg_tsresiduals(fit)
```

## Cortecosteroid drug sales
\fontsize{10}{13}\sf

```{r, echo = TRUE, dependson='h02auto'}
augment(fit) |>
  features(.innov, ljung_box, lag = 36, dof = 3)
```

## Cortecosteroid drug sales
\fontsize{10}{10}\sf

```{r h02tryharder, echo=TRUE, fig.height=3.6}
fit <- h02 |>
  model(best = ARIMA(log(Cost),
    stepwise = FALSE, approximation = FALSE,
    order_constraint = p + q + P + Q <= 9))
report(fit)
```

## Cortecosteroid drug sales

```{r, echo=TRUE, fig.height=4, dependson='h02tryharder'}
gg_tsresiduals(fit)
```

## Cortecosteroid drug sales
\fontsize{10}{12}\sf

```{r, echo = TRUE, dependson='h02tryharder'}
augment(fit) |>
  features(.innov, ljung_box, lag = 36, dof = 9)
```

## Cortecosteroid drug sales
\fontsize{10}{12}\sf

Training data: July 1991 to June 2006

Test data: July 2006--June 2008

```r
fit <- h02 |>
  filter_index(~ "2006 Jun") |>
  model(
    ARIMA(log(Cost) ~ 0 + pdq(3, 0, 0) + PDQ(2, 1, 0)),
    ARIMA(log(Cost) ~ 0 + pdq(3, 0, 1) + PDQ(2, 1, 0)),
    ARIMA(log(Cost) ~ 0 + pdq(3, 0, 2) + PDQ(2, 1, 0)),
    ARIMA(log(Cost) ~ 0 + pdq(3, 0, 1) + PDQ(1, 1, 0))
    # ... #
  )

fit |>
  forecast(h = "2 years") |>
  accuracy(h02)
```

## Cortecosteroid drug sales
\fontsize{13}{15}\sf

```{r h02-rmse, cache=TRUE, echo=FALSE}
models <- list(
  c(3, 0, 1, 0, 1, 2),
  c(3, 0, 1, 1, 1, 1),
  c(3, 0, 1, 0, 1, 1),
  c(3, 0, 1, 2, 1, 0),
  c(3, 0, 0, 2, 1, 0),
  c(3, 0, 2, 2, 1, 0),
  c(3, 0, 1, 1, 1, 0),
  c(2, 1, 0, 0, 1, 1),
  c(4, 1, 1, 2, 1, 2)
)

model_defs <- map(models, ~ ARIMA(log(Cost) ~ 0 + pdq(!!.[1], !!.[2], !!.[3]) + PDQ(!!.[4], !!.[5], !!.[6])))
model_defs <- set_names(model_defs, map_chr(
  models,
  ~ sprintf("ARIMA(%i,%i,%i)(%i,%i,%i)[12]", .[1], .[2], .[3], .[4], .[5], .[6])
))

fit <- h02 |>
  filter_index(~"2006 Jun") |>
  model(!!!model_defs)

fit |>
  forecast(h = "2 years") |>
  accuracy(h02) |>
  arrange(RMSE) |>
  select(.model, RMSE) |>
  knitr::kable(digits = 4)
```

## Cortecosteroid drug sales

  * Models with lowest AICc values tend to give slightly better results than the other models.
  * AICc comparisons must have the same orders of differencing. But RMSE test set comparisons can involve any models.
  * Use the best model available, even if it does not pass all tests.

## Cortecosteroid drug sales

```{r h02f, echo=TRUE}
fit <- h02 |>
  model(ARIMA(Cost ~ 0 + pdq(3, 0, 1) + PDQ(0, 1, 2)))
fit |>
  forecast() |>
  autoplot(h02) + labs(y = "H02 Expenditure ($AUD)")
```
