---
title: "ETC3550/ETC5550 Applied&nbsp;forecasting"
author: "Ch9. ARIMA models"
institute: "OTexts.org/fpp3/"
pdf-engine: pdflatex
fig-width: 7.5
fig-height: 3.5
format:
  beamer:
    theme: monash
    aspectratio: 169
    fontsize: 14pt
    section-titles: false
    knitr:
      opts_chunk:
        dev: "cairo_pdf"
include-in-header: header.tex
execute:
  echo: false
  message: false
  warning: false
---


```{r setup, include=FALSE}
source("setup.R")
library(patchwork)
library(purrr)
```


## Point forecasts

1. Rearrange ARIMA equation so $y_t$ is on LHS.
2. Rewrite equation by replacing $t$ by $T+h$.
3. On RHS, replace future observations by their forecasts, future errors by zero, and past errors by corresponding residuals.

Start with $h=1$. Repeat for $h=2,3,\dots$.

## Prediction intervals
\vspace*{0.2cm}\fontsize{14}{15}\sf

\begin{block}{95\% prediction interval}
$$\hat{y}_{T+h|T} \pm 1.96\sqrt{v_{T+h|T}}$$
where $v_{T+h|T}$ is estimated forecast variance.
\end{block}\pause\vspace*{-0.3cm}

* $v_{T+1|T}=\hat{\sigma}^2$ for all ARIMA models regardless of parameters and orders.\pause
* Multi-step prediction intervals for ARIMA(0,0,$q$):
\centerline{$\displaystyle y_t = \varepsilon_t + \sum_{i=1}^q \theta_i \varepsilon_{t-i}.$}
\centerline{$\displaystyle
v_{T|T+h} = \hat{\sigma}^2 \left[ 1 + \sum_{i=1}^{h-1} \theta_i^2\right], \qquad\text{for~} h=2,3,\dots.$}

## Prediction intervals

* Prediction intervals **increase in size with forecast horizon**.
* Prediction intervals can be difficult to calculate by hand
* Calculations assume residuals are **uncorrelated** and **normally distributed**.
* Prediction intervals tend to be too narrow.
    * the uncertainty in the parameter estimates has not been accounted for.
    * the ARIMA model assumes historical patterns will not change during the forecast period.
    * the ARIMA model assumes uncorrelated future \rlap{errors}


## Seasonal ARIMA models

| ARIMA | $~\underbrace{(p, d, q)}$ | $\underbrace{(P, D, Q)_{m}}$ |
| ----: | :-----------------------: | :--------------------------: |
|       | ${\uparrow}$              | ${\uparrow}$                 |
|       | Non-seasonal part         | Seasonal part of             |
|       | of the model              | of the model                 |

where $m =$ number of observations per year.

## Seasonal ARIMA models

E.g., ARIMA$(1, 1, 1)(1, 1, 1)_{4}$ model (without constant)\pause
$$(1 - \phi_{1}B)(1 - \Phi_{1}B^{4}) (1 - B) (1 - B^{4})y_{t} ~= ~
(1 + \theta_{1}B) (1 + \Theta_{1}B^{4})\varepsilon_{t}.
$$\pause\vspace*{-1cm}

\setlength{\unitlength}{1mm}
\begin{footnotesize}
\begin{picture}(100,25)(-5,0)
\thinlines
{\put(15,22){\vector(0,1){6}}}
{\put(32,10){\vector(0,1){18}}}
{\put(48,22){\vector(0,1){6}}}
{\put(66,10){\vector(0,1){18}}}
{\put(97,22){\vector(0,1){6}}}
{\put(115,10){\vector(0,1){18}}}
{\put(0,17){$\left(\begin{array}{@{}c@{}} \text{Non-seasonal} \\ \text{AR(1)}
                    \end{array}\right)$}}
{\put(22,5){$\left(\begin{array}{@{}c@{}} \text{Seasonal} \\ \text{AR(1)}
                    \end{array}\right)$}}
{\put(35,17){$\left(\begin{array}{@{}c@{}} \text{Non-seasonal} \\ \text{difference}
                    \end{array}\right)$}}
{\put(55,5){$\left(\begin{array}{@{}c@{}} \text{Seasonal} \\ \text{difference}
                    \end{array}\right)$}}
{\put(85,17){$\left(\begin{array}{@{}c@{}} \text{Non-seasonal} \\ \text{MA(1)}
                    \end{array}\right)$}}
{\put(105,5){$\left(\begin{array}{@{}c@{}} \text{Seasonal} \\ \text{MA(1)}
                    \end{array}\right)$}}
\end{picture}
\end{footnotesize}

\vspace*{10cm}

## Seasonal ARIMA models

E.g., ARIMA$(1, 1, 1)(1, 1, 1)_{4}$ model (without constant)
$$(1 - \phi_{1}B)(1 - \Phi_{1}B^{4}) (1 - B) (1 - B^{4})y_{t} ~= ~
(1 + \theta_{1}B) (1 + \Theta_{1}B^{4})\varepsilon_{t}.
$$\vspace*{-1cm}

All the factors can be multiplied out and the general model
written as follows:\vspace*{-0.7cm}
\begin{align*}
y_{t} &= (1 + \phi_{1})y_{t - 1} - \phi_1y_{t-2} + (1 + \Phi_{1})y_{t - 4}\\
&\text{}
 - (1 + \phi_{1} + \Phi_{1} + \phi_{1}\Phi_{1})y_{t - 5}
 + (\phi_{1} + \phi_{1} \Phi_{1}) y_{t - 6} \\
& \text{} - \Phi_{1} y_{t - 8} + (\Phi_{1} + \phi_{1} \Phi_{1}) y_{t - 9}
  - \phi_{1} \Phi_{1} y_{t - 10}\\
  &\text{}
  + \varepsilon_{t} + \theta_{1}\varepsilon_{t - 1} + \Theta_{1}\varepsilon_{t - 4} + \theta_{1}\Theta_{1}\varepsilon_{t - 5}.
\end{align*}
\vspace*{10cm}

## Seasonal ARIMA models
The seasonal part of an AR or MA model will be seen in the seasonal lags of
the PACF and ACF.

\alert{ARIMA(0,0,0)(0,0,1)$_{12}$ will show:}\vspace*{-0.2cm}

  * a spike at lag 12 in the ACF but no other significant spikes.
  * The PACF will show exponential decay in the seasonal lags; that is, at lags 12, 24, 36, \dots.

\alert{ARIMA(0,0,0)(1,0,0)$_{12}$ will show:}\vspace*{-0.2cm}

  * exponential decay in the seasonal lags of the ACF
  * a single significant spike at lag 12 in the PACF.


## ARIMA vs ETS
\fontsize{14}{16}\sf

  * Myth that ARIMA models are more general than exponential smoothing.
  * Linear exponential smoothing models all special cases of ARIMA models.
  * Non-linear exponential smoothing models have no equivalent ARIMA counterparts.
  * Many ARIMA models have no exponential smoothing counterparts.
  * ETS models all non-stationary. Models with seasonality or non-damped trend (or both) have two unit roots; all other models have one unit \rlap{root.}

\vspace*{10cm}

## ARIMA vs ETS

```{r venn, echo=FALSE}
#| fig-height: 3
#| fig-width: 5
library(latex2exp)
cols <- c(ets = "#D55E00", arima = "#0072b2")
tibble(
  x = c(-1, 1),
  y = c(-0.5, -0.5),
  labels = c("ets", "arima"),
) |>
  ggplot(aes(colour = labels, fill = labels)) +
  ggforce::geom_circle(aes(x0 = x, y0 = y, r = 1.6), alpha = 0.3, linewidth = 1) +
  scale_colour_manual(values = cols) +
  scale_fill_manual(values = cols) +
  coord_fixed() +
  guides(fill = "none") +
  annotate("text", label = "ETS models", x = -1.5, y = 1.35, col = cols["ets"], fontface = "bold", size = 6, family = "Fira Sans") +
  annotate("text", label = "Combination\n of components", x = -1.3, y = 0.5, col = cols["ets"], size = 2.8, fontface = "bold", family = "Fira Sans") +
  annotate("text", label = "9 ETS models with\n multiplicative errors", x = -1.6, y = -0.5, col = cols["ets"], size = 2.8, family = "Fira Sans") +
  annotate("text", label = "3 ETS models with\n additive errors and\n multiplicative\n seasonality", x = -1.3, y = -1.4, col = cols["ets"], size = 2.8, family = "Fira Sans") +
  annotate("text", label = "ARIMA models", x = 1.5, y = 1.35, col = cols["arima"], fontface = "bold", size = 6, family = "Fira Sans") +
  annotate("text", label = "Modelling\n autocorrelations", x = 1.3, y = 0.5, col = cols["arima"], size = 2.8, fontface = "bold", family = "Fira Sans") +
  annotate("text", label = TeX("Potentially $\\infty$ models"), x = 1.6, y = -0.6, col = cols["arima"], size = 2.8, family = "Fira Sans") +
  annotate("text", label = "All stationary models\n Many large models", x = 1.25, y = -1.5, col = cols["arima"], size = 2.8, family = "Fira Sans") +
  annotate("text", label = "6 fully additive\n ETS models", x = 0, y = -0.6, col = "#6b6859", size = 2.8, family = "Fira Sans") +
  guides(colour = "none", fill = "none") +
  theme_void()
```

## Equivalences

\fontsize{13}{15}\sf

|**ETS model**  | **ARIMA model**             | **Parameters**                       |
| :------------ | :-------------------------- | :----------------------------------- |
| ETS(A,N,N)    | ARIMA(0,1,1)                | $\theta_1 = \alpha-1$                |
| ETS(A,A,N)    | ARIMA(0,2,2)                | $\theta_1 = \alpha+\beta-2$          |
|               |                             | $\theta_2 = 1-\alpha$                |
| ETS(A,A\damped,N)    | ARIMA(1,1,2)                | $\phi_1=\phi$                        |
|               |                             | $\theta_1 = \alpha+\phi\beta-1-\phi$ |
|               |                             | $\theta_2 = (1-\alpha)\phi$          |
| ETS(A,N,A)    | ARIMA(0,0,$m$)(0,1,0)$_m$   |                                      |
| ETS(A,A,A)    | ARIMA(0,1,$m+1$)(0,1,0)$_m$ |                                      |
| ETS(A,A\damped,A)    | ARIMA(1,0,$m+1$)(0,1,0)$_m$ |                                      |
