---
title: "ETC3550/ETC5550 Applied&nbsp;forecasting"
author: "Revision"
date: "OTexts.org/fpp3/"
classoption: aspectratio=169
titlepage: title16x9.png
titlecolor: burntorange
toc: true
output:
  binb::monash:
    colortheme: monashwhite
    fig_width: 7.5
    fig_height: 3.5
    keep_tex: no
    includes:
      in_header: header.tex
---

```{r setup, include=FALSE}
source("setup.R")
melsyd <- tsibbledata::ansett %>%
  filter(Airports == "MEL-SYD")
```

# Some case studies

## CASE STUDY 1: Paperware company

\fontsize{11.5}{13}\sf

\begin{textblock}{9.2}(0.2,1.5)
\textbf{Problem:} Want forecasts of each of hundreds of
items. Series can be stationary, trended or seasonal. They currently
have a large forecasting program written in-house but it doesn't seem
to produce sensible forecasts. They want me to fix it.

\textbf{Additional information}\vspace*{-0.2cm}\fontsize{11.5}{13}\sf
\begin{itemize}\itemsep=0cm\parskip=0cm
\item  Program  written in COBOL making numerical calculations limited. It is not possible to do any optimisation.
\item Their programmer has little experience in numerical computing.
\item They employ no statisticians and want the program to produce forecasts automatically.
\end{itemize}
\end{textblock}

\placefig{10.2}{1.4}{width=5.8cm}{tableware2}

## CASE STUDY 1: Paperware company

### Methods currently used

A
: 12 month average

C
: 6 month average

E
: straight line regression over last 12 months

G
: straight line regression over last 6 months

H
: average slope between last year's and this year's values.
  (Equivalent to differencing at lag 12 and taking mean.)

I
: Same as H except over 6 months.

K
: I couldn't understand the explanation.

## CASE STUDY 2: PBS

\fullwidth{pills}

## CASE STUDY 2: PBS

### The Pharmaceutical Benefits Scheme (PBS) is the Australian government drugs subsidy scheme.

  * Many drugs bought from pharmacies are subsidised to allow more equitable access to modern drugs.
  * The cost to government is determined by the number and types of drugs purchased. Currently nearly 1\% of GDP.
  * The total cost is budgeted based on forecasts of drug usage.

## CASE STUDY 2: PBS

\fullheight{pbs2}

## CASE STUDY 2: PBS

  * In 2001: \$4.5 billion budget, under-forecasted by \$800 million.
  * Thousands of products. Seasonal demand.
  * Subject to covert marketing, volatile products, uncontrollable expenditure.
  * Although monthly data available for 10 years, data are aggregated to annual values, and only the first three years are used in estimating the forecasts.
  * All forecasts being done with the \texttt{FORECAST} function in MS-Excel!

## CASE STUDY 3: Car fleet company

**Client:** One of Australia's largest car fleet companies

**Problem:** how to forecast resale value of vehicles? How
should this affect leasing and sales policies?

\pause

### Additional information
 - They can provide a large amount of data on previous vehicles and their eventual resale values.
 - The resale values are currently estimated by a group of specialists. They see me as a threat and do not cooperate.

## CASE STUDY 4: Airline

\fullheight{ansettlogo}

## CASE STUDY 4: Airline

```{r, echo=FALSE}
melsyd %>%
  filter(Class == "Economy") %>%
  autoplot(Passengers) +
  labs(
    title = "Economy class passengers",
    subtitle = "Melbourne-Sydney",
    x = "Year",
    y = "Thousands"
  )
```

\only<2>{\begin{textblock}{4.2}(11,5.5)
\begin{alertblock}{}
Not the real data! Or is it?
\end{alertblock}
\end{textblock}}

## CASE STUDY 4: Airline

**Problem:** how to forecast passenger traffic on major routes?

### Additional information

  * They can provide a large amount of data on previous routes.
  * Traffic is affected by school holidays, special events such as
the Grand Prix, advertising campaigns, competition behaviour, etc.
  * They have a highly capable team of people who are able to do
most of the computing.

# Review of topics covered

## 1. Introduction to forecasting and R

 - time series data and `tsibble` objects.
 - what makes things hard/easy to forecast.
 - Reading and writing simple R code. Interpreting R output.

## 2. Time series graphics

 - time plot.
 - season plot.
 - subseries plot.
 - lag plot.
 - ACF.
 - white noise.

## 3: Time series decomposition

  - Describing a time series: seasonality, trend, cycles, changing variance, unusual features.
  - Difference between seasonality and cyclicity.
  - Interpreting a decomposition.
  - Seasonal adjustment.
  - Forecasting and decomposition.

## 5. The forecasters' toolbox
\fontsize{14}{15}\sf

  - Four benchmark methods: naive, seasonal naive, drift, mean.
  - Transformations to stabilize changing variance.
  - Forecasting involves distributions of future observations.
  - Residual diagnostics: ACF, white noise, LB test.
  - Training/test sets. Out-of-sample accuracy.
  - Problem of over-fitting.
  - Measures of forecast accuracy: MAE, MSE, RMSE, MAPE, MASE.
  - One-step prediction intervals based on RMSE from residuals.
  - Time series cross-validation.

## 8: Exponential smoothing

  - Simple exponential smoothing.
  - Holt's local trend method
  - Damped trend methods
  - Holt-Winters seasonal method (additive and multiplicative versions)
  - ETS state space formulation
  - Interpretation of output in R.
  - Computing forecasts by setting future $\varepsilon_t$ to 0.
  - Assumptions for prediction intervals

## 9: ARIMA models
\fontsize{14}{16}\sf

  - Stationarity.
  - Differencing: first-differences and seasonal-differences. What to use when.
  - White noise, random walk, AR($p$), MA($q$), ARMA($p$,$q$),
        ARIMA($p,d,q$), ARIMA($p,d,q$)($P,D,Q$)$_s$.
  - Backshift operator notation
  - ACF, PACF.  Model identification.
  - ARIMA models, Seasonal ARIMA models
  - Order selection and goodness of fit (AICc)
  - Interpretation of output in R.

## 9: ARIMA models (cont'd)

  - Expanding out an ARIMA model for forecasting
  - Finding point forecasts for given ARIMA process
  - Assumptions for prediction intervals
  - One-step prediction intervals based on RMSE
  - Effect of differencing on forecasts.
  - Effect of a constant on forecasts.
  - ARIMA vs ETS


## 6: Multiple regression

  - Dummy variables, seasonal dummies, piecewise linear trends, interventions
  - Harmonic regression
  - AIC, AICc, BIC, $R^2$, adjusted $R^2$
  - variable selection
  - Interpretation of R output.
  - ex ante vs ex post forecasts
  - scenario forecasting

## 10: Dynamic regression models
\fontsize{14}{16}\sf

  - Regression with ARMA errors
  - Using lagged predictors
  - Difference between regression residuals and ARIMA residuals.
  - Problems with OLS and autocorrelated errors
  - Forecasting for regression models with ARMA errors
  - Stochastic vs deterministic trends
  - Dynamic harmonic regression

# Exam

## Exam: 9.30am (AEST) 2 July

Five questions, all to be attempted.

A
: Short answers/explanations. Write about 1/4 page on four topics (out of six possible topics). Nuanced answers required.

B
: Describing a time series, choosing a forecasting method

C
: ETS models

D
: ARIMA models

E
: (Dynamic) regression models\newline (with extra part for PG students)

###
  $\bullet$ Moodle Quiz\qquad $\bullet$ Open book\qquad $\bullet$ 3.5 hours

## Exam and R

 * Parts \alert{B}--\alert{C} require interpretation of R output, but no coding.
 * Parts \alert{D}--\alert{E} require some coding (part of the code will be provided) and interpretation of R output.
 * All R coding will be very similar to examples you have done before.
 * Every student will have different data sets.
 * Submitted answers will be automatically checked for close matches.
 * Enter answers on Moodle as you go, to avoid internet issues at the end.

## Preparing for the exam

  - Exams from 2015--2019 on Moodle. Solutions to follow soon.
  - Exercises.  Make sure you have done them all!
  - Identify your weak points and practice them.
  - Write your own summary of the material.
  - Practice explaining the material to a class-mate.

\pause

### Help available
 * Ask on Moodle forum
 * See your tutor during the consultation times.

## Useful resources for forecasters

**Organization:**

 * International Institute of Forecasters.

**Annual Conference:**

 * 2020 International Symposium on Forecasting,\newline \st{Rio de Janeiro, Brazil}. \alert{Online}

**Journals:**

 * International Journal of Forecasting
 * Foresight

###
Links to all of the above at **forecasters.org**


## Happy forecasting

\begin{block}{}
Good forecasters are not smarter than everyone else, they merely have their ignorance better organised.
\end{block}

\vspace*{2cm}\pause

\begin{alertblock}{}\centering
Please fill in your SETU
\end{alertblock}
